---
title: "MACseq Plate 1 Analysis AUG2025"
output: html_document
date: "2025-08-27"
---
```{r setup, echo = FALSE, message =FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.height=8, fig.width = 10)
```

```{r}
### loading required packages
library(tidyverse)
library(Seurat)
library(gtools)
library(tidyseurat)
library(DESeq2)
library(edgeR)
library(DT)
library(dplyr)
library(pheatmap)
library(EnhancedVolcano)
library(RColorBrewer)
library(ggvenn)
library(gplots)
library(WGCNA)
library(tidyr)
library(ggplot2)
```

```{r}
### Functions 

source('/Users/dalvinikita/Desktop/macpie/R/filter_genes_by_expression.R')

### buttons for datatables 
create_gene_dt <- function(x){
  colnames <- c('logFC','AveExpr','t','B')
  df <- x %>%
    mutate(across(all_of(colnames),round, digits = 2)) %>%
    mutate(adj.P.Val = format(adj.P.Val, format='e', digits=4)) %>%
    mutate(P.Value = format(P.Value, format='e', digits=4))
  DT::datatable(df,
                extensions = 'Buttons',
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All"))))
}

create_dt <- function(x){
  DT::datatable(x,
                extensions = 'Buttons',
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All"))))
}

plotRLE <- function(data, main = "", ylab = "Relative Log Expression") {
  # Calculate medians
  col <- brewer.pal(8, 'Set3')
  medians <- apply(data, 1, median, na.rm = TRUE)
  rle_data <- data - medians
  
  boxplot(as.data.frame(rle_data),
          outline = FALSE,
          las = 2,
          ylab = ylab,
          main = main,
          col = col)
  abline(h = 0, col = "red", lty = 2)
}

red_blue_palette <- colorRampPalette(brewer.pal(11, "RdBu"))(100)
```

## Notes

Thresholds used for DE analysis:

• pvalue threshold = 0.05

• log FC threshold = log2(1.2) = 20% change in expression compared to control

Differences from previous analyses of MACseq plate 1:

• Utilising new filtering method: count of 10 in at least 80/96 of samples within an ORG line. 

```{r,echo=TRUE}
### Threshold Constants

p_value_threshold <- 0.05
```

## Data
```{r}
### Importing MACseq data
### Importing metadata
### Joining data in a Seurat object 

### import metadata
metadata<-read.csv("/Volumes/bioinf/home/ndalvi/MACseq_PMC228/metadata/PMC228.csv",header=T) %>%
  arrange("Barcode") %>%
  mutate(id=gsub("Plate","",Plate_ID)) %>%
  mutate(label=paste0(Treatment_1," ","(",Well_ID,")")) %>%
  mutate(interaction_condition=paste0(Drug,".",SPLINTR)) %>%
  mutate(analysis_condition=paste0(Organoid,".",SPLINTR,".",Time))

### load in count data 
raw_counts <- Read10X(data.dir = '/Users/dalvinikita/Documents/3_MACseq_PMC228/raw/PMC228/starsolo/SA_PMC228-null.Solo.out/Gene/raw')

row.names(metadata)<-metadata$label

mac <- CreateSeuratObject(counts = raw_counts,
                          project = 'PMC228',
                          min.cells = 1,
                          min.features = 1)

mac[["percent.mt"]] <- PercentageFeatureSet(mac, pattern = "^mt-|^MT-")
mac[["percent.ribo"]] <- PercentageFeatureSet(mac, pattern = "^Rp[slp][[:digit:]]|^Rpsa|^RP[SLP][[:digit:]]|^RPSA")

### add location metadata
mac<- mac %>%
  inner_join(metadata,by=c(".cell"="Barcode"))

mac <- filter_genes_by_expression(mac,
                                       group_by = "Organoid",
                                       min_counts = 30,
                                       min_samples = 91)

### exclude mito and ribo genes from the counts
mac<-mac[!grepl("^MT-",rownames(mac))]
mac<-mac[!grepl("^RP[SL][[:digit:]]|^RPLP[[:digit:]]|^RPSA",rownames(mac))]

colnames(mac)<- mac$label
```

A quick look at the metadata:
```{r}
create_dt(metadata[metadata$Organoid=="ORG38"| metadata$Organoid=='ORG49' | metadata$Organoid=="ORG66",c('Barcode','Well_ID','Organoid','SPLINTR','Drug','Concentration','Time_2')])
```

## Basic Visualisation {.tabset}

### Violin Plot

```{r, fig.height=5, fig.width=8}
vlnplot <- VlnPlot(mac, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.ribo"),
    ncol = 4, group.by = "Organoid")

vlnplot

ggsave(vlnplot, filename = '../output/P1/violin_plot.png', width=8, height=5)
```

### Pre and post normalisation {.tabset}

#### Entire Plate

DMSO raw VS normalised wells across entire plate:

```{r, fig.width=15, fig.height=6}
###   Select wells from mac that are not empty 
rawdata <- mac
sampleinfo <- metadata[is.na(metadata$Cell_type)!=TRUE | metadata$Treatment_1!="Staurosporine",]
samples <- rownames(sampleinfo)
rawdata <- mac[,samples]

### Keep only DMSO treated wells
rawdata <- rawdata %>%
  filter(Treatment_1=="DMSO")

### Ensure counts data and sampleinfo are in the same order
order <- colnames(rawdata)
sampleinfo <- sampleinfo[order,]
order_check <- rownames(sampleinfo)==colnames(rawdata)
cat('Instances where metadata sample matches counts matric sample (checking correct order):',"\n", order_check)

dge <- DGEList(counts = rawdata@assays$RNA$counts)
dge <- calcNormFactors(dge, methods = "TMMwsp")
group <- as.factor(sampleinfo$interaction_condition)
organoid <- as.factor(sampleinfo$Organoid)
design <- model.matrix(~0+organoid+group)
v <- voom(dge, design)
```

```{r, fig.width=10, fig.height=5}
par(mar = c(8, 4, 4, 2) + 0.1)

par(mfrow=c(1,2))
plotRLE(as.data.frame(rawdata@assays$RNA$counts), main='RLE of all DMSO treated wells - raw', )
plotRLE(v$E, main='RLE of all DMSO treated wells - normalised')
```

## Differential Expression Analysis 

### ORG38 

#### Design Matrix and Contrasts 

Factors to include in design matrix: 

• SPLINTR/treatment group (drug_label, drug_unlabelled)

• Concentration (IC12.5, IC25, IC50)

• Time (48H VS DMSO)

• Not including ORG as only 1 ORG line 

Design matrix:

```{r}
### selecting metadata and countdata relevant for ORG 38
### subset metadata to include only ORG38 cells at 48H
columns <- metadata[metadata$Organoid=='ORG38'&metadata$Time=='48H' & metadata$Treatment_1!="STAUR",]
sampleinfo <- columns
columns <- c(columns[,c("label")])

### subset counts to include only relevant ORG38 cells 
counts_matrix <- mac[,columns]

### factors to include in deisgn matrix: treatment/splintr, concentration, time
group <- sampleinfo$interaction_condition
concentration <- sampleinfo$Concentration
time <- sampleinfo$Time_2

### set up design matrix
### without intercept (means model)
design <- model.matrix(~0+group+concentration+time)
rownames(design) <- rownames(sampleinfo)
datatable(design)

### Specifying contrasts for analysis
pacVSdmso <- makeContrasts(((groupPACLITAXEL.SPL1+groupPACLITAXEL.SPL2+groupPACLITAXEL.UNLABELLED)/3)-((groupDMSO.SPL1+groupDMSO.SPL2+groupDMSO.UNLABELLED)/3), levels=colnames(design))

sn38VSdmso <- makeContrasts(((groupSN38.SPL1+groupSN38.SPL2+groupSN38.UNLABELLED)/3)-((groupDMSO.SPL1+groupDMSO.SPL2+groupDMSO.UNLABELLED)/3), levels=colnames(design))
```

#### Dfferential Expression {.tabset}

Voom plot: 

```{r, fig.height=5, fig.width=5}
### Normalising data
counts<-counts_matrix@assays$RNA$counts
counts <- DGEList(counts = counts) %>%
  calcNormFactors(method = 'TMMwsp')
v <- voom(counts, design, plot = T)
org38_voom_matrix <- v
```

RLE of all ORG38 wells before and after normalisation:

```{r, fig.width=16}
par(mfrow=c(1,2))
plotRLE(counts$counts, main='RLE of all wells - raw')
plotRLE(v$E, main='RLE of all wells - normalised')
```

##### Paclitaxel - DMSO

Differentially expressed genes between Paclitaxel and DMSO in ORG38:

```{r}
### following limma trend workflow
### difference between PAC and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, pacVSdmso)
fit <- treat(fit, lfc= log2(1.2))
saveRDS(fit, file='../output/P1/org38_pac_fit.rds')

all_genes <- topTreat(fit, number=Inf)
org38_pac_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')
write.csv(all_genes, file = '../data/DEgenes/P1/ORG38_ncPACvsDMSO.csv')

datatable(org38_pac_DEgenes)

### Print summary of DE genes table
up <- paste('• UPREGULATED:',sum(org38_pac_DEgenes$logFC > 0))
down <- paste('• DOWNREGULATED:',sum(org38_pac_DEgenes$logFC< 0))
total <- paste('• TOTAL DE genes:', length(rownames(org38_pac_DEgenes)))
summary=paste (total, up, down, sep="\n")
cat(summary[1])

```

###### Heatmap of DEGs

```{r, fig.height=10, fig.width=10}
### Assigning genes of interest from DEgenes table
### Formatting counts to log(CPM) for heatmap
### subsetting wells of interest
genes <- rownames(org38_pac_DEgenes)
samples_of_interest <- sampleinfo[sampleinfo$Drug=='PACLITAXEL'|sampleinfo$Drug=='DMSO',c('label')]
heatmap_counts <- as.data.frame(v$E[,samples_of_interest])


### Setting up annotation data frame
type_annotation <- data.frame(Drug = as.factor(sampleinfo[,c(17)]),
                              Concentration = as.factor(sampleinfo[,c(11)]),
                              SPLINTR = as.factor(sampleinfo[,c(16)]))
rownames(type_annotation) <- rownames(sampleinfo)

pheatmap(heatmap_counts[genes[1:50],],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         cellheight = 8,
         color=rev(red_blue_palette))
     
```

###### Volcano plot

```{r, fig.height=8, fig.width=6}
### Assigning different colours for upregulated/downregulates genes
keyvals <- ifelse(
    all_genes$logFC < -1, 'darkred',
      ifelse(all_genes$logFC > 1, 'darkgreen',
        'black'))
  keyvals[is.na(keyvals)] <- 'black'
  names(keyvals)[keyvals == 'darkgreen'] <- 'UP'
  names(keyvals)[keyvals == 'black'] <- 'NO'
  names(keyvals)[keyvals == 'darkred'] <- 'DOWN'
  
### labelling 
volcano_plot <- EnhancedVolcano(all_genes,
                                title="Paclitaxel VS DMSO",
                                subtitle="ORG38",
                                lab=rownames(all_genes),
                                selectLab = rownames(org38_pac_DEgenes)[1:10],
                                boxedLabels = TRUE,
                                drawConnectors = TRUE,
                                arrowheads = FALSE,
                                pointSize = 3,
                                colAlpha = 1, 
                                x='logFC', 
                                y='adj.P.Val', 
                                pCutoff=0.05,
                                FCcutoff = 1,
                                border="full",
                                legendPosition = "bottom",
                                colCustom = keyvals,
                                ylim = c(0,4))
volcano_plot

ggsave(plot= volcano_plot, filename = '../output/P1/VP_org38_pac.png', height=8, width=6)

```

##### SN38 - DMSO

Differentially expressed genes between SN38 and DMSO in ORG38:

```{r}
### following limma trend workflow
### difference between SN38 and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, sn38VSdmso)
fit <- treat(fit, lfc = log2(1.2))
saveRDS(fit, file='../output/P1/org38_sn38_fit.rds')
all_genes <- topTreat(fit, number=Inf)
org38_sn38_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

write.csv(all_genes, file = '../data/DEgenes/P1/ORG38_ncSN38vsDMSO.csv')

datatable(org38_sn38_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org38_sn38_DEgenes)))
up <- paste('• UPREGULATED:',sum(org38_sn38_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org38_sn38_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

###### Heatmap of DEGs

```{r, fig.height=10, fig.width=10}
### Assigning genes of interest from DEgenes table
### Formatting counts to log(CPM) for heatmap
### subsetting wells of interest
genes <- rownames(org38_sn38_DEgenes)
samples_of_interest <- sampleinfo[sampleinfo$Drug=='SN38'|sampleinfo$Drug=='DMSO',c('label')]
heatmap_counts <- as.data.frame(v$E[,samples_of_interest])


### Setting up annotation data frame
type_annotation <- data.frame(Drug = as.factor(sampleinfo[,c(17)]),
                              Concentration = as.factor(sampleinfo[,c(11)]),
                              SPLINTR = as.factor(sampleinfo[,c(16)]))
rownames(type_annotation) <- rownames(sampleinfo)

pheatmap(heatmap_counts[genes,],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         cellheight = 8,
         color=rev(red_blue_palette))
     
```

###### Volcano plot
```{r, fig.height=8, fig.width=6}
### Assigning different colours for upregulated/downregulates genes
keyvals <- ifelse(
    all_genes$logFC < -1, 'darkred',
      ifelse(all_genes$logFC > 1, 'darkgreen',
        'black'))
  keyvals[is.na(keyvals)] <- 'black'
  names(keyvals)[keyvals == 'darkgreen'] <- 'UP'
  names(keyvals)[keyvals == 'black'] <- 'NO'
  names(keyvals)[keyvals == 'darkred'] <- 'DOWN'
  
### labelling 
volcano_plot <- EnhancedVolcano(all_genes,
                                title="SN38 VS DMSO",
                                subtitle="ORG38",
                                lab=rownames(all_genes),
                                selectLab = rownames(org38_sn38_DEgenes)[1:10],
                                boxedLabels = TRUE,
                                drawConnectors = TRUE,
                                arrowheads = FALSE,
                                pointSize = 3,
                                colAlpha = 1, 
                                x='logFC', 
                                y='adj.P.Val', 
                                pCutoff=0.05,
                                FCcutoff = 1,
                                border="full",
                                legendPosition = "bottom",
                                colCustom = keyvals,
                                ylim = c(0,2))
volcano_plot

ggsave(plot= volcano_plot, filename = '../output/P1/VP_org38_SN38.png', height=8, width=6)

```

##### Gene list comparison

```{r, fig.height=6, fig.width=6}
### Venn Diagram tests
x <- list(`PAC-DMSO`=rownames(org38_pac_DEgenes),
          `SN38-DMSO`=rownames(org38_sn38_DEgenes))

x.table <- venn(x[c("PAC-DMSO","SN38-DMSO")], show.plot = FALSE)
x_attributes <- attributes(x.table)

ggvenn(x, 
       show_percentage = FALSE,
       text_size = 8)

common_genes <- x_attributes$intersections$`PAC-DMSO:SN38-DMSO`

```

`r length(common_genes)` genes were found to be shared between nc(PAC-DMSO) and nc(SN38-DMSO) : 

  • `r common_genes`
  
### ORG49 

#### Design Matrix and Contrasts 

Factors to include in design matrix: 

• SPLINTR/treatment group (drug_label, drug_unlabelled)

• Concentration (IC12.5, IC25, IC50)

• Time (48H VS DMSO)

• Not including ORG as only 1 ORG line 

Design matrix:

```{r}
### selecting metadata and countdata relevant for ORG49
### subset metadata to include only ORG49 cells at 48H
columns <- metadata[metadata$Organoid=='ORG49'&metadata$Time=='48H' & metadata$Treatment_1!="STAUR",]
sampleinfo <- columns
columns <- c(columns[,c("label")])

### subset counts to include only relevant ORG49 cells 
counts_matrix <- mac[,columns]

### factors to include in deisgn matrix: splintr/treatment, concentration, time
group <- sampleinfo$interaction_condition
concentration <- sampleinfo$Concentration
time <- sampleinfo$Time_2

### set up design matrix
### without intercept (means model)
design <- model.matrix(~0+group+concentration+time)
rownames(design) <- rownames(sampleinfo)
datatable(design)

### Specifying contrasts for analysis
pacVSdmso <- makeContrasts(((groupPACLITAXEL.SPL1+groupPACLITAXEL.SPL2+groupPACLITAXEL.UNLABELLED)/3)-((groupDMSO.SPL1+groupDMSO.SPL2+groupDMSO.UNLABELLED)/3), levels=colnames(design))

sn38VSdmso <- makeContrasts(((groupSN38.SPL1+groupSN38.SPL2+groupSN38.UNLABELLED)/3)-((groupDMSO.SPL1+groupDMSO.SPL2+groupDMSO.UNLABELLED)/3), levels=colnames(design))
```

#### Dfferential Expression {.tabset}

Voom plot: 

```{r, fig.height=5, fig.width=5}
### Normalising data
counts<-counts_matrix@assays$RNA$counts
counts <- DGEList(counts = counts) %>%
  calcNormFactors(method = 'TMMwsp')
v <- voom(counts, design, plot = T)
```

RLE of all ORG49 wells before and after normalisation:

```{r, fig.width=16}
par(mfrow=c(1,2))
plotRLE(counts$counts, main='RLE of all wells - raw')
plotRLE(v$E, main='RLE of all wells - normalised')
```

##### Paclitaxel - DMSO

Differentially expressed genes between Paclitaxel and DMSO in ORG49:

```{r}
### following limma trend workflow
### difference between PAC and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, pacVSdmso)
fit <- treat(fit, lfc=log2(1.2))
saveRDS(fit, file='../output/P1/org49_pac_fit.rds')
all_genes <- topTreat(fit, number=Inf)
org49_pac_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')
write.csv(all_genes, file = '../data/DEgenes/P1/ORG49_ncPACvsDMSO.csv')

datatable(org49_pac_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org49_pac_DEgenes)))
up <- paste('• UPREGULATED:',sum(org49_pac_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org49_pac_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])

```

###### Heatmap of DEGs

```{r, fig.height=10, fig.width=10}
### Assigning genes of interest from DEgenes table
### Formatting counts to log(CPM) for heatmap
### subsetting wells of interest
genes <- rownames(org49_pac_DEgenes)
samples_of_interest <- sampleinfo[sampleinfo$Drug=='PACLITAXEL'|sampleinfo$Drug=='DMSO',c('label')]
heatmap_counts <- as.data.frame(v$E[,samples_of_interest])


### Setting up annotation data frame
type_annotation <- data.frame(Drug = as.factor(sampleinfo[,c(17)]),
                              Concentration = as.factor(sampleinfo[,c(11)]),
                              SPLINTR = as.factor(sampleinfo[,c(16)]))
rownames(type_annotation) <- rownames(sampleinfo)

pheatmap(heatmap_counts[genes,],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         cellheight = 8,
         color=rev(red_blue_palette))
     
```

###### Volcano Plot

```{r, fig.height=8, fig.width=6}
### Assigning different colours for upregulated/downregulates genes
keyvals <- ifelse(
    all_genes$logFC < -1, 'darkred',
      ifelse(all_genes$logFC > 1, 'darkgreen',
        'black'))
  keyvals[is.na(keyvals)] <- 'black'
  names(keyvals)[keyvals == 'darkgreen'] <- 'UP'
  names(keyvals)[keyvals == 'black'] <- 'NO'
  names(keyvals)[keyvals == 'darkred'] <- 'DOWN'
  
### labelling 
volcano_plot <- EnhancedVolcano(all_genes,
                                title="PAC VS DMSO",
                                subtitle="ORG49",
                                lab=rownames(all_genes),
                                selectLab = rownames(org49_pac_DEgenes)[1:10],
                                boxedLabels = TRUE,
                                drawConnectors = TRUE,
                                arrowheads = FALSE,
                                pointSize = 3,
                                colAlpha = 1, 
                                x='logFC', 
                                y='adj.P.Val', 
                                pCutoff=0.05,
                                FCcutoff = 1,
                                border="full",
                                legendPosition = "bottom",
                                colCustom = keyvals,
                                ylim = c(0,6.5))
volcano_plot

ggsave(plot= volcano_plot, filename = '../output/P1/VP_org49_pac.png', height=8, width=6)

```

##### SN38 - DMSO

Differentially expressed genes between SN38 and DMSO in ORG49:

```{r}
### following limma trend workflow
### difference between SN38 and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, sn38VSdmso)
fit <- treat(fit, lfc=log2(1.2))
saveRDS(fit, file='../output/P1/org49_sn38_fit.rds')
all_genes <- topTreat(fit, number=Inf)
org49_sn38_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

write.csv(all_genes, file = '../data/DEgenes/P1/ORG49_ncSN38vsDMSO.csv')

datatable(org49_sn38_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org49_sn38_DEgenes)))
up <- paste('• UPREGULATED:',sum(org49_sn38_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org49_sn38_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

###### Heatmap of DEGs

```{r, fig.height=10, fig.width=10}
# ### Assigning genes of interest from DEgenes table
# ### Formatting counts to log(CPM) for heatmap
# ### subsetting wells of interest
# genes <- rownames(org49_sn38_DEgenes)
# samples_of_interest <- sampleinfo[sampleinfo$Drug=='SN38'|sampleinfo$Drug=='DMSO',c('label')]
# heatmap_counts <- as.data.frame(v$E[,samples_of_interest])
# 
# 
# ### Setting up annotation data frame
# type_annotation <- data.frame(Drug = as.factor(sampleinfo[,c(17)]),
#                               Concentration = as.factor(sampleinfo[,c(11)]),
#                               SPLINTR = as.factor(sampleinfo[,c(16)]))
# rownames(type_annotation) <- rownames(sampleinfo)
# 
# pheatmap(heatmap_counts[genes,],
#          border_color = 'NA',
#          annotation_col = type_annotation,
#          treeheight_row = 0,
#          cellheight = 8,
#          color=rev(red_blue_palette))
     
```

###### Volcano Plot 

```{r, fig.height=8, fig.width=6}
### Assigning different colours for upregulated/downregulates genes
keyvals <- ifelse(
    all_genes$logFC < -1, 'darkred',
      ifelse(all_genes$logFC > 1, 'darkgreen',
        'black'))
  keyvals[is.na(keyvals)] <- 'black'
  names(keyvals)[keyvals == 'darkgreen'] <- 'UP'
  names(keyvals)[keyvals == 'black'] <- 'NO'
  names(keyvals)[keyvals == 'darkred'] <- 'DOWN'
  
### labelling 
volcano_plot <- EnhancedVolcano(all_genes,
                                title="SN38 VS DMSO",
                                subtitle="ORG49",
                                lab=rownames(all_genes),
                                #selectLab = rownames(org49_sn38_DEgenes)[1:10],
                                boxedLabels = TRUE,
                                drawConnectors = TRUE,
                                arrowheads = FALSE,
                                pointSize = 3,
                                colAlpha = 1, 
                                x='logFC', 
                                y='adj.P.Val', 
                                pCutoff=0.05,
                                FCcutoff = 1,
                                border="full",
                                legendPosition = "bottom",
                                colCustom = keyvals)
volcano_plot

ggsave(plot= volcano_plot, filename = '../output/P1/VP_org49_SN38.png', height=8, width=6)

```

##### Gene list comparison

```{r, fig.height=6, fig.width=6}
### Venn Diagram tests
x <- list(`PAC-DMSO`=rownames(org49_pac_DEgenes),
          `SN38-DMSO`=rownames(org49_sn38_DEgenes))

x.table <- venn(x[c("PAC-DMSO","SN38-DMSO")], show.plot = FALSE)
x_attributes <- attributes(x.table)

ggvenn(x, 
       show_percentage = FALSE,
       text_size = 8)

common_genes <- x_attributes$intersections$`PAC-DMSO:SN38-DMSO`

```

`r length(common_genes)` genes were found to be shared between nc(PAC-DMSO) and nc(SN38-DMSO) : 

  • `r common_genes`
  
### ORG66 

#### Design Matrix and Contrasts 

Factors to include in design matrix: 

• SPLINTR/treatment group (drug_label, drug_unlabelled)

• Concentration (IC12.5, IC25, IC50)

• Time (48H VS DMSO)

• Not including ORG as only 1 ORG line 

Design matrix:

```{r}
### selecting metadata and countdata relevant for ORG66
### subset metadata to include only ORG66 cells at 48H
columns <- metadata[metadata$Organoid=='ORG66'&metadata$Time=='48H' & metadata$Treatment_1!="STAUR",]
sampleinfo <- columns
columns <- c(columns[,c("label")])

### subset counts to include only relevant ORG66 cells 
counts_matrix <- mac[,columns]

### factors to include in deisgn matrix: treatment/SPLINTR, concentration, time
group <- sampleinfo$interaction_condition
concentration <- sampleinfo$Concentration
time <- sampleinfo$Time_2

### set up design matrix
### without intercept (means model)
design <- model.matrix(~0+group+concentration+time)
rownames(design) <- rownames(sampleinfo)
datatable(design)

### Specifying contrasts for analysis
pacVSdmso <- makeContrasts(((groupPACLITAXEL.SPL1+groupPACLITAXEL.SPL2)/2)-((groupDMSO.SPL1+groupDMSO.SPL2)/2), levels=colnames(design))

sn38VSdmso <- makeContrasts(((groupSN38.SPL1+groupSN38.SPL2)/2)-((groupDMSO.SPL1+groupDMSO.SPL2)/2), levels=colnames(design))
```

#### Dfferential Expression {.tabset}

Voom plot: 

```{r, fig.height=5, fig.width=5}
### Normalising data
counts<-counts_matrix@assays$RNA$counts
counts <- DGEList(counts = counts) %>%
  calcNormFactors(method = 'TMMwsp')
v <- voom(counts, design, plot = T)
```

RLE of all ORG66 wells before and after normalisation:

```{r, fig.width=16}
par(mfrow=c(1,2))
plotRLE(counts$counts, main='RLE of all wells - raw')
plotRLE(v$E, main='RLE of all wells - normalised')
```

##### Paclitaxel - DMSO

Differentially expressed genes between Paclitaxel and DMSO in ORG66:

```{r}
### following limma trend workflow
### difference between PAC and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, pacVSdmso)
fit <- treat(fit, lfc=log2(1.2))
saveRDS(fit, file='../output/P1/org66_pac_fit.rds')
all_genes <- topTreat(fit, number=Inf)
org66_pac_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')
write.csv(all_genes, file = '../data/DEgenes/P1/ORG66_ncPACvsDMSO.csv')

datatable(org66_pac_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org66_pac_DEgenes)))
up <- paste('• UPREGULATED:',sum(org66_pac_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org66_pac_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

###### Heatmap of DEGs

```{r, fig.height=10, fig.width=10}
### Assigning genes of interest from DEgenes table
### Formatting counts to log(CPM) for heatmap
### subsetting wells of interest
genes <- rownames(org66_pac_DEgenes)
samples_of_interest <- sampleinfo[sampleinfo$Drug=='PACLITAXEL'|sampleinfo$Drug=='DMSO',c('label')]
heatmap_counts <- as.data.frame(v$E[,samples_of_interest])


### Setting up annotation data frame
type_annotation <- data.frame(Drug = as.factor(sampleinfo[,c(17)]),
                              Concentration = as.factor(sampleinfo[,c(11)]),
                              SPLINTR = as.factor(sampleinfo[,c(16)]))
rownames(type_annotation) <- rownames(sampleinfo)

pheatmap(heatmap_counts[genes[1:50],],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         cellheight = 8,
         color=rev(red_blue_palette))
     
```

###### Volcano Plot
```{r, fig.height=8, fig.width=6}
### Assigning different colours for upregulated/downregulates genes
keyvals <- ifelse(
    all_genes$logFC < -1, 'darkred',
      ifelse(all_genes$logFC > 1, 'darkgreen',
        'black'))
  keyvals[is.na(keyvals)] <- 'black'
  names(keyvals)[keyvals == 'darkgreen'] <- 'UP'
  names(keyvals)[keyvals == 'black'] <- 'NO'
  names(keyvals)[keyvals == 'darkred'] <- 'DOWN'
  
### labelling 
volcano_plot <- EnhancedVolcano(all_genes,
                                title="PAC VS DMSO",
                                subtitle="ORG66",
                                lab=rownames(all_genes),
                                selectLab = rownames(org66_pac_DEgenes)[1:10],
                                boxedLabels = TRUE,
                                drawConnectors = TRUE,
                                arrowheads = FALSE,
                                pointSize = 3,
                                colAlpha = 1, 
                                x='logFC', 
                                y='adj.P.Val', 
                                pCutoff=0.05,
                                FCcutoff = 1,
                                border="full",
                                legendPosition = "bottom",
                                colCustom = keyvals)
                                #ylim = c(0,2))
volcano_plot

ggsave(plot= volcano_plot, filename = '../output/P1/VP_org66_pac.png', height=8, width=6)

```

##### SN38 - DMSO

Differentially expressed genes between SN38 and DMSO in ORG66:

```{r}
### following limma trend workflow
### difference between SN38 and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, sn38VSdmso)
fit <- treat(fit, lfc=log2(1.2))
saveRDS(fit, file='../output/P1/org66_sn38_fit.rds')
all_genes <- topTreat(fit, number=Inf)
org66_sn38_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

write.csv(all_genes, file = '../data/DEgenes/P1/ORG66_ncSN38vsDMSO.csv')

datatable(org66_sn38_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org66_sn38_DEgenes)))
up <- paste('• UPREGULATED:',sum(org66_sn38_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org66_sn38_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

###### Heatmap of DEGs

```{r, fig.height=10, fig.width=10}
### Assigning genes of interest from DEgenes table
### Formatting counts to log(CPM) for heatmap
### subsetting wells of interest
genes <- rownames(org66_sn38_DEgenes)
samples_of_interest <- sampleinfo[sampleinfo$Drug=='SN38'|sampleinfo$Drug=='DMSO',c('label')]
heatmap_counts <- as.data.frame(v$E[,samples_of_interest])


### Setting up annotation data frame
type_annotation <- data.frame(Drug = as.factor(sampleinfo[,c(17)]),
                              Concentration = as.factor(sampleinfo[,c(11)]),
                              SPLINTR = as.factor(sampleinfo[,c(16)]))
rownames(type_annotation) <- rownames(sampleinfo)

pheatmap(heatmap_counts[genes[1:50],],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         cellheight = 8,
         color=rev(red_blue_palette))
     
```

###### Volcano Plot

```{r, fig.height=8, fig.width=6}
### Assigning different colours for upregulated/downregulates genes
keyvals <- ifelse(
    all_genes$logFC < -1, 'darkred',
      ifelse(all_genes$logFC > 1, 'darkgreen',
        'black'))
  keyvals[is.na(keyvals)] <- 'black'
  names(keyvals)[keyvals == 'darkgreen'] <- 'UP'
  names(keyvals)[keyvals == 'black'] <- 'NO'
  names(keyvals)[keyvals == 'darkred'] <- 'DOWN'
  
### labelling 
volcano_plot <- EnhancedVolcano(all_genes,
                                title="SN38 VS DMSO",
                                subtitle="ORG66",
                                lab=rownames(all_genes),
                                selectLab = rownames(org66_sn38_DEgenes)[1:10],
                                boxedLabels = TRUE,
                                drawConnectors = TRUE,
                                arrowheads = FALSE,
                                pointSize = 3,
                                colAlpha = 1, 
                                x='logFC', 
                                y='adj.P.Val', 
                                pCutoff=0.05,
                                FCcutoff = 1,
                                border="full",
                                legendPosition = "bottom",
                                colCustom = keyvals,
                                ylim = c(0,4))
volcano_plot

ggsave(plot= volcano_plot, filename = '../output/P1/VP_org66_SN38.png', height=8, width=6)

```

##### Gene list comparison

```{r, fig.height=6, fig.width=6}
### Venn Diagram tests
x <- list(`PAC-DMSO`=rownames(org66_pac_DEgenes),
          `SN38-DMSO`=rownames(org66_sn38_DEgenes))

x.table <- venn(x[c("PAC-DMSO","SN38-DMSO")], show.plot = FALSE)
x_attributes <- attributes(x.table)

ggvenn(x, 
       show_percentage = FALSE,
       text_size = 8)

common_genes <- x_attributes$intersections$`PAC-DMSO:SN38-DMSO`

```

`r length(common_genes)` genes were found to be shared between nc(PAC-DMSO) and nc(SN38-DMSO) : 

  • `r common_genes`

## Pathway Enrichment Analysis 

### Comparing gene lists between organoids {.tabset}

#### Paclitaxel - DMSO

```{r, fig.height=6, fig.width=6}
### Venn Diagram df
x <- list(ORG38=rownames(org38_pac_DEgenes),
          ORG49=rownames(org49_pac_DEgenes),
          ORG66=rownames(org66_pac_DEgenes))

x.table <- venn(x, show.plot = FALSE)
x_attributes <- attributes(x.table)

venn <- ggvenn(x, 
       show_percentage = FALSE,
       text_size = 8)

ggsave(plot = venn, filename = '../output/P1/VD_all_pac.png', height=6, width=6)

common_genes <- paste(x_attributes$intersections$`ORG38:ORG49:ORG66`)
common_genes_38_49 <- paste(x_attributes$intersections$`ORG38:ORG49`)
common_genes_38_66 <- paste(x_attributes$intersections$`ORG38:ORG66`)
```

`r length(common_genes)` genes were shared across all ORGs. 

  • `r common_genes`
  
`r length(common_genes_38_49)` genes were shared between ORG38 and ORG49. 

  • `r common_genes_38_49`
  
`r length(common_genes_38_66)` genes were shared between ORG38 and ORG66. 

  • `r common_genes_38_66`
  

#### SN38 - DMSO

```{r, fig.height=6, fig.width=6}
### Venn Diagram df
x <- list(ORG38=rownames(org38_sn38_DEgenes),
          ORG49=rownames(org49_sn38_DEgenes),
          ORG66=rownames(org66_sn38_DEgenes))

x.table <- venn(x, show.plot = FALSE)
x_attributes <- attributes(x.table)

venn <- ggvenn(x, 
       show_percentage = FALSE,
       text_size = 8)

ggsave(plot = venn, filename = '../output/P1/VD_all_SN38.png', height=6, width=6)

common_genes <- paste(x_attributes$intersections$`ORG38:ORG49:ORG66`)
```

`r length(common_genes)` genes were shared across all ORGs. 

  • `r common_genes`
  
### Barcode plots + enrichment testing (limma::camera)

#### Between treatments, within an organoid {.tabset}

##### ORG38

DEGs identified from SN38-DMSO comparison in ORG38, ranked against t-statistic generated from PAC-DMSO comparison in ORG38, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org38_pac_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org38_sn38_DEgenes[org38_sn38_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org38_sn38_DEgenes[org38_sn38_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in SN38 - DMSO","downregulated in SN38 - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```


DEGs identified from PAC-DMSO comparison in ORG38, ranked against t-statistic generated from SN38-DMSO comparison in ORG38, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org38_sn38_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org38_pac_DEgenes[org38_pac_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org38_pac_DEgenes[org38_pac_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in PAC - DMSO","downregulated in PAC - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```

##### ORG49

DEGs identified from SN38-DMSO comparison in ORG49, ranked against t-statistic generated from PAC-DMSO comparison in ORG49, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org49_pac_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org49_sn38_DEgenes[org49_sn38_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org49_sn38_DEgenes[org49_sn38_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in SN38 - DMSO","downregulated in SN38 - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```


DEGs identified from PAC-DMSO comparison in ORG49, ranked against t-statistic generated from SN38-DMSO comparison in ORG49, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org49_sn38_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org49_pac_DEgenes[org49_pac_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org49_pac_DEgenes[org49_pac_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in PAC - DMSO","downregulated in PAC - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```

##### ORG66

DEGs identified from SN38-DMSO comparison in ORG66, ranked against t-statistic generated from PAC-DMSO comparison in ORG66, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org66_pac_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org66_sn38_DEgenes[org66_sn38_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org66_sn38_DEgenes[org66_sn38_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in SN38 - DMSO","downregulated in SN38 - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```


DEGs identified from PAC-DMSO comparison in ORG66, ranked against t-statistic generated from SN38-DMSO comparison in ORG66, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org66_sn38_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org66_pac_DEgenes[org66_pac_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org66_pac_DEgenes[org66_pac_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in PAC - DMSO","downregulated in PAC - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```

#### Between orgnaoids, within a treatment {.tabset}

##### Paclitaxel - DMSO {.tabset}

###### ORG38 vs ORG49

DEGs identified from PAC-DMSO comparison in ORG49, ranked against t-statistic generated from PAC-DMSO comparison in ORG38, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org38_pac_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org49_pac_DEgenes[org49_pac_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org49_pac_DEgenes[org49_pac_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in PAC - DMSO","downregulated in PAC - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```

###### ORG49 vs ORG38

DEGs identified from PAC-DMSO comparison in ORG38, ranked against t-statistic generated from PAC-DMSO comparison in ORG49, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org49_pac_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org38_pac_DEgenes[org38_pac_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org38_pac_DEgenes[org38_pac_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in PAC - DMSO","downregulated in PAC - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```

###### ORG38 vs ORG66

DEGs identified from PAC-DMSO comparison in ORG66, ranked against t-statistic generated from PAC-DMSO comparison in ORG38, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org38_pac_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org66_pac_DEgenes[org66_pac_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org66_pac_DEgenes[org66_pac_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in PAC - DMSO","downregulated in PAC - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```

###### ORG66 vs ORG38

DEGs identified from PAC-DMSO comparison in ORG38, ranked against t-statistic generated from PAC-DMSO comparison in ORG66, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org66_pac_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org38_pac_DEgenes[org38_pac_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org38_pac_DEgenes[org38_pac_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in PAC - DMSO","downregulated in PAC - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```

###### ORG49 vs ORG66

DEGs identified from PAC-DMSO comparison in ORG66, ranked against t-statistic generated from PAC-DMSO comparison in ORG49, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org49_pac_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org66_pac_DEgenes[org66_pac_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org66_pac_DEgenes[org66_pac_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in PAC - DMSO","downregulated in PAC - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```

###### ORG66 vs ORG49

DEGs identified from PAC-DMSO comparison in ORG49, ranked against t-statistic generated from PAC-DMSO comparison in ORG66, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org66_pac_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org49_pac_DEgenes[org49_pac_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org49_pac_DEgenes[org49_pac_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in PAC - DMSO","downregulated in PAC - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```


##### SN38-DMSO {.tabset}

###### ORG38 vs ORG49

DEGs identified from SN38-DMSO comparison in ORG49, ranked against t-statistic generated from SN38-DMSO comparison in ORG38, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org38_sn38_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org49_sn38_DEgenes[org49_sn38_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org49_sn38_DEgenes[org49_sn38_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in SN38 - DMSO","downregulated in SN38 - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```

###### ORG49 vs ORG38

DEGs identified from SN38-DMSO comparison in ORG38, ranked against t-statistic generated from SN38-DMSO comparison in ORG49, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org49_sn38_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org38_sn38_DEgenes[org38_sn38_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org38_sn38_DEgenes[org38_sn38_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in SN38 - DMSO","downregulated in SN38 - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```

###### ORG38 vs ORG66

DEGs identified from SN38-DMSO comparison in ORG66, ranked against t-statistic generated from SN38-DMSO comparison in ORG38, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org38_sn38_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org66_sn38_DEgenes[org66_sn38_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org66_sn38_DEgenes[org66_sn38_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in SN38 - DMSO","downregulated in SN38 - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```

###### ORG66 vs ORG38

DEGs identified from SN38-DMSO comparison in ORG38, ranked against t-statistic generated from SN38-DMSO comparison in ORG66, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org66_sn38_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org38_sn38_DEgenes[org38_sn38_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org38_sn38_DEgenes[org38_sn38_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in SN38 - DMSO","downregulated in SN38 - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```

###### ORG49 vs ORG66

DEGs identified from SN38-DMSO comparison in ORG66, ranked against t-statistic generated from SN38-DMSO comparison in ORG49, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org49_sn38_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org66_sn38_DEgenes[org66_sn38_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org66_sn38_DEgenes[org66_sn38_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in SN38 - DMSO","downregulated in SN38 - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```

###### ORG66 vs ORG49

DEGs identified from SN38-DMSO comparison in ORG49, ranked against t-statistic generated from SN38-DMSO comparison in ORG66, and corresponding limma::camera enrichment p-values:

```{r, fig.height=5, fig.width=6}
### first set 
statistic <- readRDS('output/P1/org66_sn38_fit.rds')
statistic <- statistic$t[,1]

### second (comparison set)
set2_upregulated <- rownames(org49_sn38_DEgenes[org49_sn38_DEgenes$logFC > 0,])
set2_downregulated <- rownames(org49_sn38_DEgenes[org49_sn38_DEgenes$logFC < 0,])

### upregulated barcode plot
limma::barcodeplot(statistic = statistic,index=set2_upregulated,index2=set2_downregulated,xlab='t-statistic',labels=c("downregulated", "upregulated")) 
legend("top",legend=c("upregulated in SN38 - DMSO","downregulated in SN38 - DMSO"),lwd=2,box.lty=0,col=c("red","blue"), cex=0.8, border=NULL, inset=c(0.3,-0.1), xpd=NA)

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = statistic,
       index = list(upregulated=set2_upregulated, downregulated=set2_downregulated),
       use.ranks = TRUE)
```


### Geneset testing

#### Paclitaxel Resistance {.tabset}

Papers used for comparison:

[Sherman-Baust et al., 2011](https://pmc.ncbi.nlm.nih.gov/articles/PMC3259089/#sec20)

• Generated drug resistant OvCa cell lines (cisplatin, paclitaxel, or doxorubicin)

• Identified DEGs between resistant and sensitive phenotypes 

[McCorkle et al., 2021](https://pmc.ncbi.nlm.nih.gov/articles/PMC8336885/)

• Established in vitro paclitaxel-resistant ovarian cancer cell line and organoid models

• DEGs identified between resistant and sensitive lines and organoids 

[Yang et al., 2024](https://pmc.ncbi.nlm.nih.gov/articles/PMC10996034/#SD2-ol-27-5-14365)

• A2780 cell line and the A2780-Taxol-resistant cell line were purchased and cultured 

• RNAseq between resistant and parental cell lines to identify DEGs

```{r}
### loading in genesets

### Sherman-Baust, 2011
sherman_baust <- read_excel('../data/Literature_genesets/SHERMAN_BAUST_2011.XLSX', sheet='geneset_testing')
sherman_baust_up <- as.data.frame(sherman_baust[sherman_baust$OV90P > 1,c('Symbol')])
sherman_baust_dn <- as.data.frame(sherman_baust[sherman_baust$OV90P < -1,c('Symbol')])

### McCorkle, 2021
mccorkle <- read_excel('../data/Literature_genesets/MCCORKLE_2021.xlsx', sheet='geneset_testing')
mccorkle_up <- as.data.frame(mccorkle[mccorkle$log2FoldChange > 1, c('hgnc_symbol')])
mccorkle_dn <- as.data.frame(mccorkle[mccorkle$log2FoldChange < -1, c('hgnc_symbol')])

### Yang, 2024
yang <- read_excel('../data/Literature_genesets/YANGetal2024.xlsx', sheet='geneset_testing')
yang_up <- as.data.frame(yang[yang$A.128PTX_logFC > 1, c('Gene_name')])
yang_dn <- as.data.frame(yang[yang$A.128PTX_logFC < -1, c('Gene_name')])

### creating geneset list 
paclitaxel_resistance_genesets <- list(sherman_baust_up = sherman_baust_up$Symbol,
                                       sherman_baust_dn = sherman_baust_dn$Symbol,
                                       mccorkle_up = mccorkle_up$hgnc_symbol,
                                       mccorkle_dn = mccorkle_dn$hgnc_symbol,
                                       yang_up = yang_up$Gene_name,
                                       yang_dn = yang_dn$Gene_name)


### loading "fit"
org38_pac_fit <- readRDS('../output/P1/org38_pac_fit.rds') 
org38_sn38_fit <- readRDS('../output/P1/org38_sn38_fit.rds') 
org49_pac_fit <- readRDS('../output/P1/org49_pac_fit.rds') 
org49_sn38_fit <- readRDS('../output/P1/org49_sn38_fit.rds') 
org66_pac_fit <- readRDS('../output/P1/org66_pac_fit.rds')
org66_sn38_fit <- readRDS('../output/P1/org66_sn38_fit.rds') 
```

##### ORG38

```{r}
### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = org38_pac_fit$t[,1],
       index = paclitaxel_resistance_genesets,
       use.ranks = TRUE)
```

##### ORG49

```{r}
### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = org49_pac_fit$t[,1],
       index = paclitaxel_resistance_genesets,
       use.ranks = TRUE)
```

##### ORG66

```{r}
### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = org66_pac_fit$t[,1],
       index = paclitaxel_resistance_genesets,
       use.ranks = TRUE)
```

#### SN38 Resistance {.tabset}

```{r}
### Jensen, 2015 - HCT116 cell line
jensen_HCT116 <- read_excel('../data/Literature_genesets/JENSEN_2015.xlsx', sheet='HCT116-SN38')
jensen_HCT116_up <- as.data.frame(jensen_HCT116[jensen_HCT116$`Fold-change` > 1,c('GeneSymbol')])
jensen_HCT116_dn <- as.data.frame(jensen_HCT116[jensen_HCT116$`Fold-change` < -1,c('GeneSymbol')])

### Jensen, 2015 - HT29 cell line
jensen_HT29 <- read_excel('../data/Literature_genesets/JENSEN_2015.xlsx', sheet='HT29-SN38')
jensen_HT29_up <- as.data.frame(jensen_HT29[jensen_HT29$`Fold-change` > 1,c('GeneSymbol')])
jensen_HT29_dn <- as.data.frame(jensen_HT29[jensen_HT29$`Fold-change` < -1,c('GeneSymbol')])

### Jensen, 2015 - LoVo cell line
jensen_LoVo <- read_excel('../data/Literature_genesets/JENSEN_2015.xlsx', sheet='LoVo-SN38')
jensen_LoVo_up <- as.data.frame(jensen_LoVo[jensen_LoVo$`Fold-change` > 1,c('GeneSymbol')])
jensen_LoVo_dn <- as.data.frame(jensen_LoVo[jensen_LoVo$`Fold-change` < -1,c('GeneSymbol')])

### creating SN38 resistance gene lists
sn38_resistance_genesets <- list(jensen_HCT116_up = jensen_HCT116_up$GeneSymbol,
                                       jensen_HCT116_dn = jensen_HCT116_dn$GeneSymbol,
                                       jensen_HT29_up = jensen_HT29_up$GeneSymbol,
                                       jensen_HT29_dn = jensen_HT29_dn$GeneSymbol,
                                       jensen_LoVo_up = jensen_LoVo_up$GeneSymbol,
                                       jensen_LoVo_dn = jensen_LoVo_dn$GeneSymbol)
```

##### ORG38

```{r}
### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = org38_sn38_fit$t[,1],
       index = sn38_resistance_genesets,
       use.ranks = TRUE)
```

##### ORG49

```{r}
### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = org49_sn38_fit$t[,1],
       index = sn38_resistance_genesets,
       use.ranks = TRUE)
```

##### ORG66

```{r}
### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = org66_sn38_fit$t[,1],
       index = sn38_resistance_genesets,
       use.ranks = TRUE)
```

### Alicia meeting
```{r, fig.height=3.5, fig.width=6}
data <- v

### subset genes of interest 
genes <- c('MUCL3','SDCBP2','PLAUR','TPI1')
data <- data[genes,]

### identify samples of interest by treatment group
samples <- sampleinfo[sampleinfo$Drug == 'PACLITAXEL'|
                        sampleinfo$Drug == 'DMSO',]
samples <- samples[samples$Time=='48H',]
samples <- samples$label

### subset data to include only samples of interest
data <- data[, samples]
expr_data <- as.data.frame(t(data$E))

### get treatment information
samples <- rownames(expr_data)
drug <- sampleinfo[samples, c('Drug')]
concentration <- sampleinfo[samples, c('Concentration')]

expr_data <- expr_data %>%
  mutate(drug = drug) %>%
  mutate(concentration = concentration) %>%
  mutate(plate= "Plate 1")

### calculate means of each group 
mean_MUCL3 <- expr_data %>%
  group_by(drug) %>%
  summarise(mean_val = mean(MUCL3))

mean_SDCBP2 <- expr_data %>%
  group_by(drug) %>%
  summarise(mean_val = mean(SDCBP2))

mean_TPI1 <- expr_data %>%
  group_by(drug) %>%
  summarise(mean_val = mean(TPI1))

mean_PLAUR <- expr_data %>%
  group_by(drug) %>%
  summarise(mean_val = mean(PLAUR))

### add mean to expr_data for each treatment group
expr_data <- expr_data %>%
  mutate(MUCL3_mean = case_when(
    drug == 'DMSO' ~ '8.49',
    drug == 'PACLITAXEL' ~ '9.75'
  )) %>%
  mutate(SDCBP2_mean = case_when(
    drug == 'DMSO' ~ '7.03',
    drug == 'PACLITAXEL' ~ '9.19'
  )) %>%
  mutate(TPI1_mean = case_when(
    drug == 'DMSO' ~ '11.1',
    drug == 'PACLITAXEL' ~ '10.4'
  )) %>%
  mutate(PLAUR_mean = case_when(
    drug == 'DMSO' ~ '7.28',
    drug == 'PACLITAXEL' ~ '9.25'
  ))

### plotting
ggplot(data=expr_data, aes(y=MUCL3, x=plate, colour = drug)) +
  geom_point() +
  geom_hline(data = expr_data, aes(yintercept = as.numeric(MUCL3_mean), col=drug))

ggplot(data=expr_data, aes(y=SDCBP2, x=plate, colour = drug)) +
  geom_point() +
  geom_hline(data = expr_data, aes(yintercept = as.numeric(SDCBP2_mean), col=drug))

ggplot(data=expr_data, aes(y=TPI1, x=plate, colour = drug)) +
  geom_point() +
  geom_hline(data = expr_data, aes(yintercept = as.numeric(TPI1_mean), col=drug))

ggplot(data=expr_data, aes(y=PLAUR, x=plate, colour = drug)) +
  geom_point() +
  geom_hline(data = expr_data, aes(yintercept = as.numeric(PLAUR_mean), col=drug))
```
