---
title: "MACseq Plate 2 Analysis APR2025"
output: html_document
date: "2025-04-28"
---

```{r setup, echo = FALSE, message =FALSE, warning=FALSE, fig.height=6, fig.width=8}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.height = 8, fig.width = 6)
```

```{r}
### loading required packages
library(tidyverse)
library(Seurat)
library(gtools)
library(tidyseurat)
library(DESeq2)
library(edgeR)
library(RUVSeq)
library(variancePartition)
library(reshape2)
library(gridExtra)
library(ggrepel)
library(ggpubr)
library(fgsea)
library(org.Hs.eg.db)
library(DT)
library(DOSE)
library(enrichplot)
library(clusterProfiler)
library(dplyr)
library(pheatmap)
library(EnhancedVolcano)
library(msigdbr)
library(ReactomePA)
library(RColorBrewer)
```

```{r}
### Functions 

source('/Users/dalvinikita/Desktop/macpie/R/filter_genes_by_expression.R')

### buttons for datatables 
create_dt <- function(x){
  colnames <- c('logFC','AveExpr','t','B')
  df <- x %>%
    mutate(across(all_of(colnames),round, digits = 2)) %>%
    mutate(adj.P.Val = format(adj.P.Val, format='e', digits=4)) %>%
    mutate(P.Value = format(P.Value, format='e', digits=4))
  DT::datatable(df,
                extensions = 'Buttons',
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All"))))
}

plotRLE <- function(data, main = "", ylab = "Relative Log Expression") {
  # Calculate medians
  col <- brewer.pal(8, 'Set3')
  medians <- apply(data, 1, median, na.rm = TRUE)
  rle_data <- data - medians
  
  boxplot(as.data.frame(rle_data),
          outline = FALSE,
          las = 2,
          ylab = ylab,
          main = main,
          col = col)
  abline(h = 0, col = "red", lty = 2)
}

red_blue_palette <- colorRampPalette(brewer.pal(11, "RdBu"))(100)
```

## Notes

Thresholds used for DE analysis:

• pvalue threshold = 0.05

• log FC threshold = lfc > 1 | lfc < -1

## Data

```{r}
### Importing MACseq data
### Importing metadata
### Joining data in a Seurat object 

### import metadata
metadata<-read.csv("/Users/dalvinikita/Documents/11_MACseq_PMMSq042/PMMSq042.csv",header=T) %>%
  arrange("Barcode") %>%
  mutate(id=gsub("Plate","",Plate_ID)) %>%
  mutate(Treatment_Combination=paste0(substr(Cell_type,1,2),'_',substr(Treatment_1,1,4))) %>%
  mutate(label=paste0(Treatment_Combination,"_",Well_ID)) %>%
  mutate(Concentration_1=as.factor(Concentration_1)) %>%
  mutate(SPLINTR = str_sub(Cell_type,-1,-1)) %>%
  mutate(group = paste0(Treatment_Combination,'.',SPLINTR)) %>%
  mutate(combined_id = paste0(Cell_type, '_', Treatment_1))


### load in count data 
raw_counts <- Read10X(data.dir = '/Users/dalvinikita/Documents/11_MACseq_PMMSq042/raw')

### filter for the minimum number of reads per gene per sample, as per edgeR
row.names(metadata)<-metadata$label

mac <- CreateSeuratObject(counts = raw_counts,
                          project = 'PMMSq042',
                          min.cells = 1,
                          min.features = 0)

mac[["percent.mt"]] <- PercentageFeatureSet(mac, pattern = "^mt-|^MT-")
mac[["percent.ribo"]] <- PercentageFeatureSet(mac, pattern = "^Rp[slp][[:digit:]]|^Rpsa|^RP[SLP][[:digit:]]|^RPSA")

### add location metadata
mac<- mac %>%
  inner_join(metadata,by=c(".cell"="Barcode"))

mac <- filter_genes_by_expression(mac,
                                       group_by = "Organoid",
                                       min_counts = 10,
                                       min_samples = 48)

### exclude mito and ribo genes from the counts
mac<-mac[!grepl("^MT-",rownames(mac))]
mac<-mac[!grepl("^RP[SL][[:digit:]]|^RPLP[[:digit:]]|^RPSA",rownames(mac))]

colnames(mac)<- mac$label
```
A quick look at the metadata:

```{r}
datatable(metadata[metadata$Sample_type!="EMPTY",c("Well_ID","Barcode","Organoid", "Treatment_Combination", "SPLINTR", "Treatment_1")])
```
## Basic Visualisation {.tabset}

### Violin Plot

```{r, fig.height=5, fig.width=8}
VlnPlot(mac, features = c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.ribo"),
    ncol = 4, group.by = "Sample_type")
```

### Pre and post normalisation {.tabset}

#### Entire Plate

DMSO raw VS normalised wells across entire plate:

```{r, fig.width=18, fig.height=6}
###   Select wells from mac that are not empty 
rawdata <- mac
sampleinfo <- metadata[is.na(metadata$Cell_type)!=TRUE | metadata$Treatment_1!="Staurosporine",]
samples <- rownames(sampleinfo)
rawdata <- mac[,samples]

### Keep only DMSO treated wells
rawdata <- rawdata %>%
  filter(Treatment_1=="DMSO")

### Ensure counts data and sampleinfo are in the same order
order <- colnames(rawdata)
sampleinfo <- sampleinfo[order,]
order_check <- rownames(sampleinfo)==colnames(rawdata)
cat('Instances where metadata sample matches counts matric sample (checking correct order):',"\n", order_check)

dge <- DGEList(counts = rawdata@assays$RNA$counts)
dge <- calcNormFactors(dge, methods = "TMMwsp")
treatment_combination <- as.factor(sampleinfo$Treatment_Combination)
splintr <- as.factor(sampleinfo$SPLINTR)
concentration <- as.factor(sampleinfo$Concentration_1)
organoid <- as.factor(sampleinfo$Organoid)
design <- model.matrix(~0+organoid+treatment_combination+splintr+concentration)
v <- voom(dge, design)

par(mfrow=c(1,2))
plotRLE(as.data.frame(rawdata@assays$RNA$counts), main='RLE of all DMSO treated wells - raw')
plotRLE(v$E, main='RLE of all DMSO treated wells - normalised')

```

## Analysis

### ORG38 

#### Design Matrix and Contrasts 

Factors to include in design matrix: 

• SPLINTR/treatment group (splintrA/ splintrB)

• Treatment combination

• Concentration (High/Low for each treatment combination)

Design matrix:

```{r}
### selecting metadata and countdata relevant for ORG 38
### subset metadata to include only ORG38 cells 
columns <- metadata[metadata$Organoid=='ORG38' & metadata$Treatment_1!="Staurosporine",]
sampleinfo <- columns
samples <- c(rownames(sampleinfo))
counts_matrix <- mac[,samples]
order <- colnames(counts_matrix)
sampleinfo <- sampleinfo[colnames(counts_matrix),]

### factors to include in deisgn matrix: SPLINTR group and treatment, concentration
treatment <- as.factor(sampleinfo$Treatment_Combination)
concentration <- sampleinfo$Concentration_1
splintr <- as.factor(sampleinfo$SPLINTR)

### set up design matrix
### without intercept (means model)
design <- model.matrix(~0+treatment+splintr+concentration)
rownames(design) <- rownames(sampleinfo)
datatable(design)

NC_pacvsdmso <- makeContrasts(treatmentNC_Pacl-treatmentNC_DMSO,levels=colnames(design))
NC_sn38vsdmso <- makeContrasts(treatmentNC_SN38-treatmentNC_DMSO,levels=colnames(design))
PAC_pacvsdmso <- makeContrasts(treatmentPa_Pacl-treatmentPa_DMSO,levels=colnames(design))
SN38_sn38vsdmso <- makeContrasts(treatmentSN_SN38-treatmentSN_DMSO,levels=colnames(design))
long_PAC <- makeContrasts((treatmentPa_Pacl-treatmentPa_DMSO)-(treatmentNC_Pacl-treatmentNC_DMSO),levels=colnames(design))
long_SN38 <- makeContrasts((treatmentSN_SN38-treatmentSN_DMSO)-(treatmentNC_SN38-treatmentNC_DMSO),levels=colnames(design))

```

#### Differential Expression {.tabset}

Voom plot: 

```{r}
### Normalising data
counts<-counts_matrix@assays$RNA$counts
counts <- DGEList(counts = counts) %>%
  calcNormFactors(method = 'TMMwsp')
v <- voom(counts, design, plot = T, save.plot = TRUE)
```

RLE of all ORG38 wells before and after normalisation:

```{r, fig.width=16}
par(mfrow=c(1,2))
plotRLE(counts$counts, main='RLE of all wells - raw')
plotRLE(v$E, main='RLE of all wells - normalised')
```

##### NC_PAC-NC_DMSO

Differentially expressed genes between NC Pac and NC DMSO in ORG38:

```{r}
### following limma workflow
### difference between PAC and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_pacvsdmso)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

datatable(DE_genes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

```{r, fig.height=6, fig.width=8}
### Assign Significance 
all_genes$significance <- 'nonsignificant'
all_genes[all_genes$logFC < -1 & all_genes$adj.P.Val < 0.05,c('significance')] <- 'downregulated'
all_genes[all_genes$logFC >1 & all_genes$adj.P.Val < 0.05,c('significance')] <- 'upregulated'

### MA plot
MA <- new("MAList")
MA$A <- all_genes$AveExpr
MA$M <- all_genes$logFC
status <- all_genes$significance
values <- c("downregulated","upregulated","nonsignificant")
col <- c("blue", "red", "black")
plotMA(MA,main="MA-Plot with significant genes highlighted",
       status=status, values=values, hl.col=col)
```

Next, we can look the differentially expressed genes by visualising the voom-transformed counts of the treated and DMSO samples in a heatmap, to make sure that the DEgenes we've identified actually look "different" in terms of their expression. 

```{r}
### Assigning genes of interest from DEgenes table
### Formatting counts to log(CPM) for heatmap
### subsetting wells of interest
genes <- rownames(DE_genes)
samples_of_interest <- sampleinfo[sampleinfo$Treatment_Combination=='NC_Pacl'|sampleinfo$Treatment_Combination=='NC_DMSO',c('label')]
heatmap_counts <- as.data.frame(v$E[,samples_of_interest])


### Setting up annotation data frame
type_annotation <- data.frame(Treatment = as.factor(sampleinfo[,c(18)]),
                              Concentration = as.factor(sampleinfo[,c(11)]),
                              SPLINTR = as.factor(sampleinfo[,c(16)]))
rownames(type_annotation) <- rownames(sampleinfo)
```

Heatmap of differentially expressed genes between NC_PAC and NC_DMSO in PAC/DMSO samples in ORG38:

```{r, fig.height=8, fig.width=8}
pheatmap(heatmap_counts[genes,],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         cellheight = 20,
         color=rev(red_blue_palette))
```


Volcano plot of differentially expressed genes with top 10 differentially expressed genes labelled:

```{r, fig.height=6, fig.width=6}
### Assigning different colours for upregulated/downregulates genes
keyvals <- ifelse(
    all_genes$logFC < -1, 'darkred',
      ifelse(all_genes$logFC > 1, 'darkgreen',
        'black'))
  keyvals[is.na(keyvals)] <- 'black'
  names(keyvals)[keyvals == 'darkgreen'] <- 'UP'
  names(keyvals)[keyvals == 'black'] <- 'NO'
  names(keyvals)[keyvals == 'darkred'] <- 'DOWN'
  
### Volcano plot 
volcano_plot <- EnhancedVolcano(all_genes,
                                title="NC Paclitaxel VS NC DMSO",
                                subtitle="ORG38",
                                lab=rownames(all_genes),
                                selectLab = rownames(DE_genes)[1:10],
                                boxedLabels = TRUE,
                                drawConnectors = TRUE,
                                arrowheads = FALSE,
                                pointSize = 3,
                                colAlpha = 1, 
                                x='logFC', 
                                y='adj.P.Val', 
                                pCutoff=0.05,
                                FCcutoff = 1,
                                border="full",
                                legendPosition = "bottom",
                                colCustom = keyvals)
volcano_plot
```

Pathway analysis results of differentially expressed genes:

```{r, fig.height=8, fig.width=6}
### Pathway Analysis
### returning a list of corresponding entrez IDs for gene names in DE genes
entrezids <- bitr(rownames(DE_genes), fromType=c("SYMBOL"), toType="ENTREZID", OrgDb="org.Hs.eg.db")

### add entrezid data
DE_genes$gene <- rownames(DE_genes)
pathway_data<- left_join(DE_genes, entrezids, by=join_by(gene==SYMBOL))

### removing genes that dont map to an entrezID (~19.85%)
pathway_data <- drop_na(pathway_data)

### setting up gene list
geneList <- pathway_data$logFC
names(geneList)<- pathway_data$ENTREZID

#performing reactome enrichment analysis 
x <- ReactomePA::enrichPathway(gene=names(geneList),
                               organism='human',
                               pvalueCutoff =1,
                               qvalueCutoff = 1,
                               readable=TRUE)

### Functional enrichment result using PA:

datatable(x@result)

### Pathway visualisation:

p1<- barplot(x) 
p1 +scale_fill_gradient(low="blue", high="red")
```

##### NC_SN38-NC_DMSO

Differentially expressed genes between NC SN38 and NC DMSO in ORG38:

```{r}
### following limma workflow
### difference between SN38 and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_sn38vsdmso)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

datatable(DE_genes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

```{r, fig.height=6, fig.width=8}
### Assign Significance 
all_genes$significance <- 'nonsignificant'
all_genes[all_genes$logFC < -1 & all_genes$adj.P.Val < 0.05,c('significance')] <- 'downregulated'
all_genes[all_genes$logFC >1 & all_genes$adj.P.Val < 0.05,c('significance')] <- 'upregulated'

### MA plot
MA <- new("MAList")
MA$A <- all_genes$AveExpr
MA$M <- all_genes$logFC
status <- all_genes$significance
values <- c("downregulated","upregulated","nonsignificant")
col <- c("blue", "red", "black")
plotMA(MA,main="MA-Plot with significant genes highlighted",
       status=status, values=values, hl.col=col)
```

Heatmap of top 50 differentially expressed genes between NC_SN38 and NC_DMSO in SN38/DMSO samples in ORG38:

```{r, fig.height=8, fig.width=8}
### Assigning genes of interest from DEgenes table
### Formatting counts to log(CPM) for heatmap
### subsetting wells of interest
genes <- rownames(DE_genes)
samples_of_interest <- sampleinfo[sampleinfo$Treatment_Combination=='NC_SN38'|sampleinfo$Treatment_Combination=='NC_DMSO',c('label')]
heatmap_counts <- as.data.frame(v$E[,samples_of_interest])


### Setting up annotation data frame
type_annotation <- data.frame(Treatment = as.factor(sampleinfo[,c(18)]),
                              Concentration = as.factor(sampleinfo[,c(11)]),
                              SPLINTR = as.factor(sampleinfo[,c(16)]))
rownames(type_annotation) <- rownames(sampleinfo)

pheatmap(heatmap_counts[genes[1:50],],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         cellheight = 8,
         color=rev(red_blue_palette))
```

##### PAC_PAC - PAC_DMSO

Differentially expressed genes between PAC_PAC and PAC_DMSO in ORG38:

```{r}
### following limma workflow
### difference between PAC and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, PAC_pacvsdmso)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

datatable(DE_genes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

##### SN38_SN38 - SN38_DMSO)

Differentially expressed genes between SN38_SN38 and SN38_DMSO in ORG38:

```{r}
### following limma workflow
### difference between SN38 and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, SN38_sn38vsdmso)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

datatable(DE_genes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

Heatmap of differentially expressed genes between SN_SN38 and SN_DMSO in SN38/DMSO samples in ORG38:

```{r, fig.height=8, fig.width=8}
### Assigning genes of interest from DEgenes table
### Formatting counts to log(CPM) for heatmap
### subsetting wells of interest
genes <- rownames(DE_genes)
samples_of_interest <- sampleinfo[sampleinfo$Treatment_Combination=='SN_SN38'|sampleinfo$Treatment_Combination=='SN_DMSO',c('label')]
heatmap_counts <- as.data.frame(v$E[,samples_of_interest])


### Setting up annotation data frame
type_annotation <- data.frame(Treatment = as.factor(sampleinfo[,c(18)]),
                              Concentration = as.factor(sampleinfo[,c(11)]),
                              SPLINTR = as.factor(sampleinfo[,c(16)]))
rownames(type_annotation) <- rownames(sampleinfo)

pheatmap(heatmap_counts[genes,],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         cellheight = 8,
         color=rev(red_blue_palette))
```

##### (PAC_PAC - PAC_DMSO) - (NC_PAC-NC_DMSO)

Differentially expressed genes between (PAC_PAC - PAC_DMSO) and (NC_PAC-NC_DMSO) in ORG38:

```{r}
### following limma workflow
### difference between PAC and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, long_PAC)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

datatable(DE_genes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

Heatmap of differentially expressed genes between (PAC_PAC - PAC_DMSO) and (NC_PAC-NC_DMSO) in PAC/DMSO samples in ORG38:

```{r, fig.height=6, fig.width=10}
### Assigning genes of interest from DEgenes table
### Formatting counts to log(CPM) for heatmap
### subsetting wells of interest
genes <- rownames(DE_genes)
samples_of_interest <- sampleinfo[sampleinfo$Treatment_Combination=='Pa_Pacl'|
                                    sampleinfo$Treatment_Combination=='Pa_DMSO'|
                                    sampleinfo$Treatment_Combination=='NC_DMSO'|
                                    sampleinfo$Treatment_Combination=='NC_Pacl',c('label')]
heatmap_counts <- as.data.frame(v$E[,samples_of_interest])


### Setting up annotation data frame
type_annotation <- data.frame(Initial_Treatment = as.factor(sampleinfo[,c(10)]),
                              Treatment_Combination = as.factor(sampleinfo[,c(18)]),
                              Concentration = as.factor(sampleinfo[,c(11)]),
                              SPLINTR = as.factor(sampleinfo[,c(16)]))
rownames(type_annotation) <- rownames(sampleinfo)

pheatmap(heatmap_counts[genes,],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         cellheight = 20)
```


##### (SN38_SN38 - SN38_DMSO) - (NC_SN38-NC_DMSO)

Differentially expressed genes between (SN38_SN38 - SN38_DMSO) and (NC_SN38-NC_DMSO) in ORG38:

```{r}
### following limma workflow
### difference between SN38 and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, long_SN38)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

datatable(DE_genes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

### ORG49 

#### Design Matrix and Contrasts 

Factors to include in design matrix: 

• SPLINTR/treatment group (splintrA/ splintrB)

• Treatment combination

• Concentration

Design matrix:

```{r, fig.height=8, fig.width=6}
### selecting metadata and countdata relevant for ORG 38
### subset metadata to include only ORG38 cells 
columns <- metadata[metadata$Organoid=='ORG49' & metadata$Treatment_1!="Staurosporine",]
sampleinfo <- columns
samples <- c(rownames(sampleinfo))
counts_matrix <- mac[,samples]
order <- colnames(counts_matrix)
sampleinfo <- sampleinfo[colnames(counts_matrix),]

### factors to include in deisgn matrix: SPLINTR group and treatment, concentration
treatment <- as.factor(sampleinfo$Treatment_Combination)
concentration <- sampleinfo$Concentration_1
splintr <- as.factor(sampleinfo$SPLINTR)

### set up design matrix
### without intercept (means model)
design <- model.matrix(~0+treatment+concentration)
rownames(design) <- rownames(sampleinfo)
datatable(design)

NC_pacvsdmso <- makeContrasts(treatmentNC_Pacl-treatmentNC_DMSO,levels=colnames(design))
NC_sn38vsdmso <- makeContrasts(treatmentNC_SN38-treatmentNC_DMSO,levels=colnames(design))
PAC_pacvsdmso <- makeContrasts(treatmentPa_Pacl-treatmentPa_DMSO,levels=colnames(design))
SN38_sn38vsdmso <- makeContrasts(treatmentSN_SN38-treatmentSN_DMSO,levels=colnames(design))
long_PAC <- makeContrasts((treatmentPa_Pacl-treatmentPa_DMSO)-(treatmentNC_Pacl-treatmentNC_DMSO),levels=colnames(design))
long_SN38 <- makeContrasts((treatmentSN_SN38-treatmentSN_DMSO)-(treatmentNC_SN38-treatmentNC_DMSO),levels=colnames(design))

```

#### Differential Expression {.tabset}

Voom plot: 

```{r}
### Normalising data
counts<-counts_matrix@assays$RNA$counts
counts <- DGEList(counts = counts) %>%
  calcNormFactors(method = 'TMMwsp')
v <- voom(counts, design, plot = T, save.plot = TRUE)
```

RLE of all ORG49 wells before and after normalisation:

```{r, fig.width=16}
par(mfrow=c(1,2))
plotRLE(counts$counts, main='RLE of all wells - raw')
plotRLE(v$E, main='RLE of all wells - normalised')
```

##### NC_PAC-NC_DMSO

Differentially expressed genes between NC Pac and NC DMSO in ORG49:

```{r}
### following limma workflow
### difference between PAC and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_pacvsdmso)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```
##### NC_SN38-NC_DMSO

Differentially expressed genes between NC SN38 and NC DMSO in ORG38:

```{r}
### following limma workflow
### difference between SN38 and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_sn38vsdmso)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

datatable(DE_genes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

Heatmap of differentially expressed genes between NC_SN38 and NC_DMSO in SN38/DMSO samples in ORG49:

```{r, fig.height=8, fig.width=8}
### Assigning genes of interest from DEgenes table
### Formatting counts to log(CPM) for heatmap
### subsetting wells of interest
genes <- rownames(DE_genes)
samples_of_interest <- sampleinfo[sampleinfo$Treatment_Combination=='NC_SN38'|sampleinfo$Treatment_Combination=='NC_DMSO',c('label')]
heatmap_counts <- as.data.frame(v$E[,samples_of_interest])


### Setting up annotation data frame
type_annotation <- data.frame(Treatment = as.factor(sampleinfo[,c(18)]),
                              Concentration = as.factor(sampleinfo[,c(11)]),
                              SPLINTR = as.factor(sampleinfo[,c(16)]))
rownames(type_annotation) <- rownames(sampleinfo)

pheatmap(heatmap_counts[genes,],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         cellheight = 50,
         color=rev(red_blue_palette))
```

##### PAC_PAC - PAC_DMSO

Differentially expressed genes between PAC_PAC and PAC_DMSO in ORG49:

```{r}
### following limma workflow
### difference between PAC and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, PAC_pacvsdmso)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

datatable(DE_genes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

##### SN38_SN38 - SN38_DMSO)

Differentially expressed genes between SN38_SN38 and SN38_DMSO in ORG49:

```{r}
### following limma workflow
### difference between SN38 and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, SN38_sn38vsdmso)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

datatable(DE_genes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

Heatmap of top 50 differentially expressed genes between SN_SN38 and SN_DMSO in SN38/DMSO samples in ORG49:

```{r, fig.height=8, fig.width=8}
### Assigning genes of interest from DEgenes table
### Formatting counts to log(CPM) for heatmap
### subsetting wells of interest
genes <- rownames(DE_genes)
samples_of_interest <- sampleinfo[sampleinfo$Treatment_Combination=='SN_SN38'|sampleinfo$Treatment_Combination=='SN_DMSO',c('label')]
heatmap_counts <- as.data.frame(v$E[,samples_of_interest])


### Setting up annotation data frame
type_annotation <- data.frame(Treatment = as.factor(sampleinfo[,c(18)]),
                              Concentration = as.factor(sampleinfo[,c(11)]),
                              SPLINTR = as.factor(sampleinfo[,c(16)]))
rownames(type_annotation) <- rownames(sampleinfo)

pheatmap(heatmap_counts[genes[1:50],],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         cellheight = 8,
         color=rev(red_blue_palette))
```

##### (PAC_PAC - PAC_DMSO) - (NC_PAC-NC_DMSO)

Differentially expressed genes between (PAC_PAC - PAC_DMSO) and (NC_PAC-NC_DMSO) in ORG49:

```{r}
### following limma workflow
### difference between PAC and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, long_PAC)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

datatable(DE_genes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

##### (SN38_SN38 - SN38_DMSO) - (NC_SN38-NC_DMSO)

Differentially expressed genes between (SN38_SN38 - SN38_DMSO) and (NC_SN38-NC_DMSO) in ORG49:

```{r}
### following limma workflow
### difference between SN38 and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, long_SN38)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

datatable(DE_genes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

Heatmap of  differentially expressed genes between SN38 and DMSO in SN38/DMSO samples in ORG49:

```{r, fig.height=8, fig.width=8}
### Assigning genes of interest from DEgenes table
### Formatting counts to log(CPM) for heatmap
### subsetting wells of interest
genes <- rownames(DE_genes)
samples_of_interest <- sampleinfo[sampleinfo$Treatment_Combination=='NC_SN38'|
                                    sampleinfo$Treatment_Combination=='NC_DMSO'|
                                    sampleinfo$Treatment_Combination=='SN_DMSO'|
                                    sampleinfo$Treatment_Combination=='SN_DMSO',c('label')]
heatmap_counts <- as.data.frame(v$E[,samples_of_interest])


### Setting up annotation data frame
type_annotation <- data.frame(Treatment = as.factor(sampleinfo[,c(18)]),
                              Concentration = as.factor(sampleinfo[,c(11)]),
                              SPLINTR = as.factor(sampleinfo[,c(16)]))
rownames(type_annotation) <- rownames(sampleinfo)

pheatmap(heatmap_counts[genes,],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         cellheight = 40,
         color=rev(red_blue_palette))
```

### ORG66 

#### Design Matrix and Contrasts 

Factors to include in design matrix: 

• SPLINTR/treatment group (splintrA/ splintrB)

• Treatment combination

• Concentration

Design matrix:

```{r, fig.height=8, fig.width=6}
### selecting metadata and countdata relevant for ORG 38
### subset metadata to include only ORG38 cells 
columns <- metadata[metadata$Organoid=='ORG66' & metadata$Treatment_1!="Staurosporine",]
sampleinfo <- columns
samples <- c(rownames(sampleinfo))
counts_matrix <- mac[,samples]
order <- colnames(counts_matrix)
sampleinfo <- sampleinfo[colnames(counts_matrix),]

### factors to include in deisgn matrix: SPLINTR group and treatment, concentration
treatment <- as.factor(sampleinfo$Treatment_Combination)
concentration <- sampleinfo$Concentration_1
splintr <- as.factor(sampleinfo$SPLINTR)

### set up design matrix
### without intercept (means model)
design <- model.matrix(~0+treatment+splintr+concentration)
rownames(design) <- rownames(sampleinfo)
datatable(design)

NC_pacvsdmso <- makeContrasts(treatmentNC_Pacl-treatmentNC_DMSO,levels=colnames(design))
NC_sn38vsdmso <- makeContrasts(treatmentNC_SN38-treatmentNC_DMSO,levels=colnames(design))
PAC_pacvsdmso <- makeContrasts(treatmentPa_Pacl-treatmentPa_DMSO,levels=colnames(design))
SN38_sn38vsdmso <- makeContrasts(treatmentSN_SN38-treatmentSN_DMSO,levels=colnames(design))
long_PAC <- makeContrasts((treatmentPa_Pacl-treatmentPa_DMSO)-(treatmentNC_Pacl-treatmentNC_DMSO),levels=colnames(design))
long_SN38 <- makeContrasts((treatmentSN_SN38-treatmentSN_DMSO)-(treatmentNC_SN38-treatmentNC_DMSO),levels=colnames(design))
```

#### Differential Expression {.tabset}

Voom plot: 

```{r}
### Normalising data
counts<-counts_matrix@assays$RNA$counts
counts <- DGEList(counts = counts) %>%
  calcNormFactors(method = 'TMMwsp')
v <- voom(counts, design, plot = T, save.plot = TRUE)
```

RLE of all ORG38 wells before and after normalisation:

```{r, fig.width=16}
par(mfrow=c(1,2))
plotRLE(counts$counts, main='RLE of all wells - raw')
plotRLE(v$E, main='RLE of all wells - normalised')
```

##### NC_PAC-NC_DMSO

Differentially expressed genes between NC Pac and NC DMSO in ORG66:

```{r}
### following limma workflow
### difference between PAC and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_pacvsdmso)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

datatable(DE_genes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```


##### NC_SN38-NC_DMSO

Differentially expressed genes between NC SN38 and NC DMSO in ORG38:

```{r}
### following limma workflow
### difference between SN38 and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_sn38vsdmso)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

datatable(DE_genes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

```{r, fig.height=6, fig.width=8}
### Assign Significance 
all_genes$significance <- 'nonsignificant'
all_genes[all_genes$logFC < -1 & all_genes$adj.P.Val < 0.05,c('significance')] <- 'downregulated'
all_genes[all_genes$logFC >1 & all_genes$adj.P.Val < 0.05,c('significance')] <- 'upregulated'

### MA plot
MA <- new("MAList")
MA$A <- all_genes$AveExpr
MA$M <- all_genes$logFC
status <- all_genes$significance
values <- c("downregulated","upregulated","nonsignificant")
col <- c("blue", "red", "black")
plotMA(MA,main="MA-Plot with significant genes highlighted",
       status=status, values=values, hl.col=col)
```

Heatmap of differentially expressed genes between NC_SN38 and NC_DMSO in SN38/DMSO samples in ORG66:

```{r, fig.height=8, fig.width=8}
### Assigning genes of interest from DEgenes table
### Formatting counts to log(CPM) for heatmap
### subsetting wells of interest
genes <- rownames(DE_genes)
samples_of_interest <- sampleinfo[sampleinfo$Treatment_Combination=='NC_SN38'|sampleinfo$Treatment_Combination=='NC_DMSO',c('label')]
heatmap_counts <- as.data.frame(v$E[,samples_of_interest])


### Setting up annotation data frame
type_annotation <- data.frame(Treatment = as.factor(sampleinfo[,c(18)]),
                              Concentration = as.factor(sampleinfo[,c(11)]),
                              SPLINTR = as.factor(sampleinfo[,c(16)]))
rownames(type_annotation) <- rownames(sampleinfo)

pheatmap(heatmap_counts[genes,],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         cellheight = 20,
         color=rev(red_blue_palette))
```

##### PAC_PAC - PAC_DMSO

Differentially expressed genes between PAC_PAC and PAC_DMSO in ORG66:

```{r}
### following limma workflow
### difference between PAC and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, PAC_pacvsdmso)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

datatable(DE_genes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

##### SN38_SN38 - SN38_DMSO)

Differentially expressed genes between SN38_SN38 and SN38_DMSO in ORG66:

```{r}
### following limma workflow
### difference between SN38 and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, SN38_sn38vsdmso)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

datatable(DE_genes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

Heatmap of differentially expressed genes between SN_SN38 and SN_DMSO in SN38/DMSO samples in ORG38:

```{r, fig.height=8, fig.width=8}
### Assigning genes of interest from DEgenes table
### Formatting counts to log(CPM) for heatmap
### subsetting wells of interest
genes <- rownames(DE_genes)
samples_of_interest <- sampleinfo[sampleinfo$Treatment_Combination=='SN_SN38'|sampleinfo$Treatment_Combination=='SN_DMSO',c('label')]
heatmap_counts <- as.data.frame(v$E[,samples_of_interest])


### Setting up annotation data frame
type_annotation <- data.frame(Treatment = as.factor(sampleinfo[,c(18)]),
                              Concentration = as.factor(sampleinfo[,c(11)]),
                              SPLINTR = as.factor(sampleinfo[,c(16)]))
rownames(type_annotation) <- rownames(sampleinfo)

pheatmap(heatmap_counts[genes,],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         cellheight = 40,
         color=rev(red_blue_palette),
         cluster_rows = FALSE,
         cluster_cols=FALSE)
```

##### (PAC_PAC - PAC_DMSO) - (NC_PAC-NC_DMSO)

Differentially expressed genes between (PAC_PAC - PAC_DMSO) and (NC_PAC-NC_DMSO) in ORG66:

```{r}
### following limma workflow
### difference between PAC and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, long_PAC)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

datatable(DE_genes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```


##### (SN38_SN38 - SN38_DMSO) - (NC_SN38-NC_DMSO)

Differentially expressed genes between (SN38_SN38 - SN38_DMSO) and (NC_SN38-NC_DMSO) in ORG66:

```{r}
### following limma workflow
### difference between SN38 and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, long_SN38)
fit <- treat(fit, lfc=1)
all_genes <- topTreat(fit, number=Inf)
DE_genes <- topTreat(fit, p.value = 0.05, number=Inf, sort.by='p')

datatable(DE_genes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(DE_genes)))
up <- paste('• UPREGULATED:',sum(DE_genes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(DE_genes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

