---
title: "MACseq Plate 3 Analysis v2 - FEB2026"
output: html_document
date: "2026-02-24"
---

```{r setup}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.height=8, fig.width = 10)
```

```{r}
### loading required packages
library(tidyverse)
library(Seurat)
library(gtools)
library(tidyseurat)
library(DESeq2)
library(edgeR)
library(DT)
library(dplyr)
library(pheatmap)
library(EnhancedVolcano)
library(RColorBrewer)
library(ggvenn)
library(gplots)
library(clusterProfiler)
library(enrichplot)
```


```{r}
### Functions 

source('/Users/dalvinikita/Desktop/macpie/R/filter_genes_by_expression.R')

### buttons for datatables 
create_dt_genes <- function(x){
  colnames <- c('logFC','AveExpr','t','adj.P.Val')
  df <- x %>%
    mutate(across(all_of(colnames),round, digits = 2)) %>%
    mutate(adj.P.Val = format(adj.P.Val, format='e', digits=4)) %>%
    mutate(P.Value = format(P.Value, format='e', digits=4))
  DT::datatable(df,
                extensions = 'Buttons',
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All"))))
}

create_dt <-  function(x){
  datatable(x,
            extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
}


plotRLE <- function(data, main = "", ylab = "Relative Log Expression") {
  # Calculate medians
  col <- brewer.pal(8, 'Set3')
  medians <- apply(data, 1, median, na.rm = TRUE)
  rle_data <- data - medians
  
  boxplot(as.data.frame(rle_data),
          outline = FALSE,
          las = 2,
          ylab = ylab,
          main = main,
          col = col)
  abline(h = 0, col = "red", lty = 2)
}

red_blue_palette <- colorRampPalette(brewer.pal(11, "RdBu"))(100)
```

## Notes

Thresholds used for DE analysis:

• pvalue threshold = 0.05

• log FC threshold = log2(1.2) = 20% change in expression compared to control

```{r, echo=TRUE}
### Threshold Constants
p_value_threshold <- 0.05
```

```{r}
### Importing MACseq data
### Importing metadata
### Joining data in a Seurat object 

### import metadata
metadata<-read.csv("/Users/dalvinikita/Documents/13_MACseq_PMMSq057/metadata/PMMSq057_metadata.csv",header=T) %>%
  arrange("Barcode") %>%
  mutate(id=gsub("Plate","",Plate_ID)) %>%
  mutate(Treatment_Combination=paste0(PreTreatment,'_',substr(Treatment_1,1,4),"_",Concentration_1)) %>%
  mutate(label=paste0(Treatment_Combination,"_",Well_ID)) %>%
  mutate(Concentration_1=as.factor(Concentration_1)) %>%
  mutate(group = paste0(Treatment_Combination,'.',SPLINTR))

row.names(metadata)<-metadata$label

### load in count data 
raw_counts <- Read10X(data.dir = '/Users/dalvinikita/Documents/13_MACseq_PMMSq057/raw/starsolo/')

### filter for the minimum number of reads per gene per sample, as per edgeR
mac <- CreateSeuratObject(counts = raw_counts,
                          project = 'PMMSq057',
                          min.cells = 1,
                          min.features = 0)

mac[["percent.mt"]] <- PercentageFeatureSet(mac, pattern = "^mt-|^MT-")
mac[["percent.ribo"]] <- PercentageFeatureSet(mac, pattern = "^Rp[slp][[:digit:]]|^Rpsa|^RP[SLP][[:digit:]]|^RPSA")

### add location metadata
mac<- mac %>%
  inner_join(metadata,by=c(".cell"="Barcode"))

colnames(mac)<- mac$label

mac <- filter_genes_by_expression(mac,
                                       group_by = "Organoid",
                                       min_counts = 10,
                                       min_samples = 80)

### exclude mito and ribo genes from the counts
mac<-mac[!grepl("^MT-",rownames(mac))]
mac<-mac[!grepl("^RP[SL][[:digit:]]|^RPLP[[:digit:]]|^RPSA",rownames(mac))]

### removing ORG60 as not part of chemoresistance experiment
mac_filtered <- mac[,mac$Organoid!='Org60']
#mac_filtered <- mac_filtered[,mac_filtered$Treatment_1!='STAUR']

### removing Org60 from metadata 
samples <- colnames(mac_filtered)
metadata_filtered <- metadata[samples,]
```

### Pre and post normalisation 

#### Entire Plate

DMSO raw VS normalised wells across entire plate:

```{r, fig.width=18, fig.height=6}
###   Select wells from mac that are not empty 
rawdata <- mac_filtered
sampleinfo <- metadata_filtered[is.na(metadata_filtered$Organoid)!=TRUE | metadata$Treatment_1!="Staurosporine",]
samples <- rownames(sampleinfo)
rawdata <- mac_filtered[,samples]

### Keep only DMSO treated wells
rawdata <- rawdata %>%
  filter(Treatment_1=="DMSO")

### Ensure counts data and sampleinfo are in the same order
order <- colnames(rawdata)
sampleinfo <- sampleinfo[order,]
order_check <- rownames(sampleinfo)==colnames(rawdata)
cat('Instances where metadata sample matches counts matric sample (checking correct order):',"\n", order_check)

dge <- DGEList(counts = rawdata@assays$RNA$counts)
dge <- calcNormFactors(dge, methods = "TMMwsp")
treatment_combination <- as.factor(sampleinfo$Treatment_Combination)
splintr <- as.factor(sampleinfo$SPLINTR)
concentration <- as.factor(sampleinfo$Concentration_1)
organoid <- as.factor(sampleinfo$Organoid)
design <- model.matrix(~0+organoid+treatment_combination+splintr+concentration)
v <- voom(dge, design)

par(mfrow=c(1,2))
plotRLE(as.data.frame(rawdata@assays$RNA$counts), main='RLE of all DMSO treated wells - raw')
plotRLE(v$E, main='RLE of all DMSO treated wells - normalised')

```

## Analysis

### ORG38

#### Design matrix and contrasts

Factors to include in design matrix: 

• SPLINTR/treatment group (splintrA/ splintrB)

• Treatment combination

• Concentration (High/Low for each treatment combination)

Design matrix:


```{r}
### selecting metadata and countdata relevant for ORG 38
### subset metadata to include only ORG38 cells 
columns <- metadata_filtered[metadata_filtered$Organoid=='Org38' & metadata_filtered$Treatment_1!="STAUR",]
sampleinfo <- columns
samples <- c(rownames(sampleinfo))
counts_matrix <- mac_filtered[,samples]
order <- colnames(counts_matrix)
sampleinfo <- sampleinfo[colnames(counts_matrix),]

### factors to include in deisgn matrix: SPLINTR group and treatment, concentration
treatment <- as.factor(sampleinfo$Treatment_Combination)
splintr <- as.factor(sampleinfo$SPLINTR)

### set up design matrix
### without intercept (means model)
design <- model.matrix(~0+treatment+splintr)
rownames(design) <- rownames(sampleinfo)
create_dt(design)
org38_design <- design

NC_pacvsdmso_low <- makeContrasts(treatmentNC_PACL_1.5303-treatmentNC_DMSO_0.2,levels=colnames(design))
NC_pacvsdmso_high <- makeContrasts(treatmentNC_PACL_3.1271-treatmentNC_DMSO_0.2,levels=colnames(design))

NC_sn38vsdmso_low <- makeContrasts(treatmentNC_SN38_0.499-treatmentNC_DMSO_0.2,levels=colnames(design))
NC_sn38vsdmso_high <- makeContrasts(treatmentNC_SN38_0.998-treatmentNC_DMSO_0.2,levels=colnames(design))

PAC_pacvsdmso_low <- makeContrasts(treatmentPAC_PACL_1.5303-treatmentPAC_DMSO_0.2,levels=colnames(design))
PAC_pacvsdmso_high <- makeContrasts(treatmentPAC_PACL_3.1271-treatmentPAC_DMSO_0.2,levels=colnames(design))

SN38_sn38vsdmso_low <- makeContrasts(treatmentSN38_SN38_0.499-treatmentSN38_DMSO_0.2,levels=colnames(design))
SN38_sn38vsdmso_high <- makeContrasts(treatmentSN38_SN38_0.998-treatmentSN38_DMSO_0.2,levels=colnames(design))

```

#### Differential expression {.tabset}

Voom plot: 

```{r, fig.height=5, fig.width=5}
### Normalising data
counts<-counts_matrix@assays$RNA$counts
counts <- DGEList(counts = counts) %>%
  calcNormFactors(method = 'TMMwsp')
v <- voom(counts, design, plot = T)
org38_voom_mat <- v
```

RLE of all ORG38 wells before and after normalisation:

```{r, fig.width=16}
par(mfrow=c(1,2))
plotRLE(counts$counts, main='RLE of all wells - raw')
plotRLE(v$E, main='RLE of all wells - normalised')
```

##### NC_PAC-NC_DMSO {.tabset}

###### low_conc

Differentially expressed genes between NC PAC (low conc) and NC DMSO in ORG38:

```{r}
### following limma workflow
### difference between PAC and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_pacvsdmso_low)
fit <- treat(fit, lfc=log(1.2))
org38_ncPAC_low_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org38_ncPAC_low_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org38_ncPAC_low_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org38_ncPAC_low_DEgenes)))
up <- paste('• UPREGULATED:',sum(org38_ncPAC_low_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org38_ncPAC_low_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

###### high_conc
Differentially expressed genes between NC PAC (high conc) and NC DMSO in ORG38:

```{r}
### following limma workflow
### difference between PAC and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_pacvsdmso_high)
fit <- treat(fit, lfc=log(1.2))
org38_ncPAC_high_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org38_ncPAC_high_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org38_ncPAC_high_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org38_ncPAC_high_DEgenes)))
up <- paste('• UPREGULATED:',sum(org38_ncPAC_high_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org38_ncPAC_high_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

##### NC_SN38-NC_DMSO {.tabset}

###### low_conc

Differentially expressed genes between NC SN38 (low) and NC DMSO in ORG38:

```{r}
### following limma workflow
### difference between SN38 and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_sn38vsdmso_low)
fit <- treat(fit, lfc=log(1.2))
org38_ncSN38_low_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org38_ncSN38_low_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org38_ncSN38_low_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org38_ncSN38_low_DEgenes)))
up <- paste('• UPREGULATED:',sum(org38_ncSN38_low_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org38_ncSN38_low_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

###### high_conc

Differentially expressed genes between NC SN38 (high) and NC DMSO in ORG38:

```{r}
### following limma workflow
### difference between SN38 and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_sn38vsdmso_high)
fit <- treat(fit, lfc=log(1.2))
org38_ncSN38_high_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org38_ncSN38_high_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org38_ncSN38_high_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org38_ncSN38_high_DEgenes)))
up <- paste('• UPREGULATED:',sum(org38_ncSN38_high_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org38_ncSN38_high_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

##### PAC_PAC - PAC_DMSO {.tabset}

###### low_conc

Differentially expressed genes between PAC_PAC (low) and PAC_DMSO in ORG38:

```{r}
### following limma workflow
### difference between PAC and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, PAC_pacvsdmso_low)
fit <- treat(fit, lfc=log(1.2))
org38_paPAC_low_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org38_paPAC_low_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org38_paPAC_low_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org38_paPAC_low_DEgenes)))
up <- paste('• UPREGULATED:',sum(org38_paPAC_low_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org38_paPAC_low_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

###### high_conc

Differentially expressed genes between PAC_PAC (high) and PAC_DMSO in ORG38:

```{r}
### following limma workflow
### difference between PAC and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, PAC_pacvsdmso_high)
fit <- treat(fit, lfc=log(1.2))
org38_paPAC_high_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org38_paPAC_high_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org38_paPAC_high_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org38_paPAC_high_DEgenes)))
up <- paste('• UPREGULATED:',sum(org38_paPAC_high_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org38_paPAC_high_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

##### SN38_SN38 - SN38_DMSO {.tabset}

###### low_conc

Differentially expressed genes between SN38_SN38 (low) and SN38_DMSO in ORG38:

```{r}
### following limma workflow
### difference between SN38 and DMSO
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, SN38_sn38vsdmso_low)
fit <- treat(fit, lfc=log(1.2))
org38_snSN38_low_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org38_snSN38_low_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org38_snSN38_low_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org38_snSN38_low_DEgenes)))
up <- paste('• UPREGULATED:',sum(org38_snSN38_low_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org38_snSN38_low_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

###### high_conc

Differentially expressed genes between SN38_SN38 and SN38_DMSO in ORG38:

```{r}
### following limma workflow
### difference between SN38 and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, SN38_sn38vsdmso_high)
fit <- treat(fit, lfc=log(1.2))
org38_snSN38_high_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org38_snSN38_high_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org38_snSN38_high_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org38_snSN38_high_DEgenes)))
up <- paste('• UPREGULATED:',sum(org38_snSN38_high_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org38_snSN38_high_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

### ORG49 

#### Design matrix and contrasts

Factors to include in design matrix: 

• SPLINTR/treatment group (splintrA/ splintrB)

• Treatment combination

• Concentration (High/Low for each treatment combination)

Design matrix:


```{r}
### selecting metadata and countdata relevant for ORG 49
### subset metadata to include only ORG49 cells 
columns <- metadata_filtered[metadata_filtered$Organoid=='Org49' & metadata_filtered$Treatment_1!="STAUR",]
sampleinfo <- columns
samples <- c(rownames(sampleinfo))
counts_matrix <- mac_filtered[,samples]
order <- colnames(counts_matrix)
sampleinfo <- sampleinfo[colnames(counts_matrix),]

### factors to include in deisgn matrix: SPLINTR group and treatment
treatment <- as.factor(sampleinfo$Treatment_Combination)
splintr <- as.factor(sampleinfo$SPLINTR)

### set up design matrix
### without intercept (means model)
design <- model.matrix(~0+treatment+splintr)
rownames(design) <- rownames(sampleinfo)
create_dt(design)
org49_design <- design

NC_pacvsdmso_low <- makeContrasts(treatmentNC_PACL_0.0022- treatmentNC_DMSO_0.2,levels=colnames(design))
NC_pacvsdmso_high <- makeContrasts(treatmentNC_PACL_0.0499- treatmentNC_DMSO_0.2,levels=colnames(design))

NC_sn38vsdmso_low <- makeContrasts(treatmentNC_SN38_0.0166-treatmentNC_DMSO_0.2,levels=colnames(design))
NC_sn38vsdmso_high <- makeContrasts(treatmentNC_SN38_0.0333-treatmentNC_DMSO_0.2,levels=colnames(design))

PAC_pacvsdmso_low <- makeContrasts(treatmentPAC_PACL_0.0022-treatmentPAC_DMSO_0.2,levels=colnames(design))
PAC_pacvsdmso_high <- makeContrasts(treatmentPAC_PACL_0.0499-treatmentPAC_DMSO_0.2,levels=colnames(design))

SN38_sn38vsdmso_low <- makeContrasts(((treatmentSN38_SN38_0.0166+treatmentSN38_high_SN38_0.0166)/2)-treatmentSN38_DMSO_0.2,levels=colnames(design))
SN38_sn38vsdmso_high <- makeContrasts(((treatmentSN38_SN38_0.0333+treatmentSN38_high_SN38_0.0333)/2)-treatmentSN38_DMSO_0.2,levels=colnames(design))

```

#### Differential Expression {.tabset}

Voom plot: 

```{r, fig.height=5, fig.width=5}
### Normalising data
counts<-counts_matrix@assays$RNA$counts
counts <- DGEList(counts = counts) %>%
  calcNormFactors(method = 'TMMwsp')
v <- voom(counts, design, plot = T)
org49_voom_mat <- v
```

RLE of all ORG49 wells before and after normalisation:

```{r, fig.width=16}
par(mfrow=c(1,2))
plotRLE(counts$counts, main='RLE of all wells - raw')
plotRLE(v$E, main='RLE of all wells - normalised')
```

##### NC_PAC-NC_DMSO {.tabset}

###### low_conc

Differentially expressed genes between NC PAC (low) and NC DMSO in ORG49:

```{r}
### following limma workflow
### difference between PAC and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_pacvsdmso_low)
fit <- treat(fit, lfc=log(1.2))
org49_ncPAC_low_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org49_ncPA_low_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org49_ncPA_low_DEgenes)))
up <- paste('• UPREGULATED:',sum(org49_ncPA_low_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org49_ncPA_low_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

###### high_conc
Differentially expressed genes between NC PAC (high) and NC DMSO in ORG49:

```{r}
### following limma workflow
### difference between PAC and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_pacvsdmso_high)
fit <- treat(fit, lfc=log(1.2))
org49_ncPAC_high_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org49_ncPAC_high_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org49_ncPAC_high_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org49_ncPAC_high_DEgenes)))
up <- paste('• UPREGULATED:',sum(org49_ncPAC_high_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org49_ncPAC_high_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

##### NC_SN38-NC_DMSO {.tabset}

###### low_conc

Differentially expressed genes between NC SN38 (low) and NC DMSO in ORG49:

```{r}
### following limma workflow
### difference between SN38 and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_sn38vsdmso_low)
fit <- treat(fit, lfc=log(1.2))
org49_ncSN38_low_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org49_ncSN38_low_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org49_ncSN38_low_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org49_ncSN38_low_DEgenes)))
up <- paste('• UPREGULATED:',sum(org49_ncSN38_low_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org49_ncSN38_low_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

###### high_conc

Differentially expressed genes between NC SN38 (high) and NC DMSO in ORG49:

```{r}
### following limma workflow
### difference between SN38 and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_sn38vsdmso_high)
fit <- treat(fit, lfc=log(1.2))
org49_ncSN38_high_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org49_ncSN38_high_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org49_ncSN38_high_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org49_ncSN38_high_DEgenes)))
up <- paste('• UPREGULATED:',sum(org49_ncSN38_high_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org49_ncSN38_high_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```


##### PAC_PAC - PAC_DMSO {.tabset}

###### low_conc

Differentially expressed genes between PAC_PAC (low) and PAC_DMSO in ORG49:

```{r}
### following limma workflow
### difference between PAC and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, PAC_pacvsdmso_low)
fit <- treat(fit, lfc=log(1.2))
all_genes <- topTreat(fit, number=Inf)
org49_paPAC_low_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org49_paPAC_low_DEgenes)))
up <- paste('• UPREGULATED:',sum(org49_paPAC_low_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org49_paPAC_low_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

###### high_conc

Differentially expressed genes between PAC_PAC (high) and PAC_DMSO in ORG49:

```{r}
### following limma workflow
### difference between PAC and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, PAC_pacvsdmso_high)
fit <- treat(fit, lfc=log(1.2))
all_genes <- topTreat(fit, number=Inf)
org49_paPAC_high_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org49_paPAC_high_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org49_paPAC_high_DEgenes)))
up <- paste('• UPREGULATED:',sum(org49_paPAC_high_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org49_paPAC_high_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

##### SN38_SN38-SN38_DMSO {.tabset}

###### low_conc

Differentially expressed genes between SN38_SN38 (low) and SN38_DMSO in ORG49:

```{r}
### following limma workflow
### difference between SN38 and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, SN38_sn38vsdmso_low)
fit <- treat(fit, lfc=log(1.2))
org49_snSN38_low_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org49_snSN38_low_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org49_ncSN38_low_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org49_snSN38_low_DEgenes)))
up <- paste('• UPREGULATED:',sum(org49_snSN38_low_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org49_snSN38_low_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

###### high_conc

Differentially expressed genes between NC SN38 (high) and NC DMSO in ORG49:

```{r}
### following limma workflow
### difference between SN38 and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, SN38_sn38vsdmso_high)
fit <- treat(fit, lfc=log(1.2))
org49_snSN38_high_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org49_snSN38_high_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org49_snSN38_high_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org49_snSN38_high_DEgenes)))
up <- paste('• UPREGULATED:',sum(org49_snSN38_high_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org49_snSN38_high_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

### ORG66 

#### Design matrix and contrasts

Factors to include in design matrix: 

• SPLINTR/treatment group (splintrA/ splintrB)

• Treatment combination

• Concentration (High/Low for each treatment combination)

Design matrix:


```{r}
### selecting metadata and countdata relevant for ORG 66
### subset metadata to include only ORG66 cells 
columns <- metadata_filtered[metadata_filtered$Organoid=='Org66' & metadata_filtered$Treatment_1!="STAUR",]
sampleinfo <- columns
samples <- c(rownames(sampleinfo))
counts_matrix <- mac_filtered[,samples]
order <- colnames(counts_matrix)
sampleinfo <- sampleinfo[colnames(counts_matrix),]

### factors to include in deisgn matrix: SPLINTR group and treatment
treatment <- as.factor(sampleinfo$Treatment_Combination)
splintr <- as.factor(sampleinfo$SPLINTR)

### set up design matrix
### without intercept (means model)
design <- model.matrix(~0+treatment+splintr)
rownames(design) <- rownames(sampleinfo)
create_dt(design)
org49_design <- design

NC_pacvsdmso_low <- makeContrasts(treatmentNC_PACL_1.5968- treatmentNC_DMSO_0.2,levels=colnames(design))
NC_pacvsdmso_high <- makeContrasts(treatmentNC_PACL_3.1936- treatmentNC_DMSO_0.2,levels=colnames(design))

NC_sn38vsdmso_low <- makeContrasts(treatmentNC_SN38_0.02-treatmentNC_DMSO_0.2,levels=colnames(design))
NC_sn38vsdmso_high <- makeContrasts(treatmentNC_SN38_0.0366-treatmentNC_DMSO_0.2,levels=colnames(design))

PAC_pacvsdmso_low <- makeContrasts(treatmentPAC_PACL_1.5968-treatmentPAC_DMSO_0.2,levels=colnames(design))
PAC_pacvsdmso_high <- makeContrasts(treatmentPAC_PACL_3.1936-treatmentPAC_DMSO_0.2,levels=colnames(design))

SN38_sn38vsdmso_low <- makeContrasts(treatmentSN38_SN38_0.02-treatmentSN38_DMSO_0.2,levels=colnames(design))
SN38_sn38vsdmso_high <-makeContrasts(treatmentSN38_SN38_0.0366-treatmentSN38_DMSO_0.2,levels=colnames(design))

```

#### Differential Expression {.tabset}

Voom plot: 

```{r, fig.height=5, fig.width=5}
### Normalising data
counts<-counts_matrix@assays$RNA$counts
counts <- DGEList(counts = counts) %>%
  calcNormFactors(method = 'TMMwsp')
v <- voom(counts, design, plot = T)
org66_voom_mat <- v
```

RLE of all ORG66 wells before and after normalisation:

```{r, fig.width=16}
par(mfrow=c(1,2))
plotRLE(counts$counts, main='RLE of all wells - raw')
plotRLE(v$E, main='RLE of all wells - normalised')
```

##### NC_PAC-NC_DMSO {.tabset}

###### low_conc

Differentially expressed genes between NC PAC (low) and NC DMSO in ORG66:

```{r}
### following limma workflow
### difference between PAC and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_pacvsdmso_low)
fit <- treat(fit, lfc=log(1.2))
org66_ncPAC_low_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org66_ncPA_low_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org66_ncPA_low_DEgenes)))
up <- paste('• UPREGULATED:',sum(org66_ncPA_low_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org66_ncPA_low_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

###### high_conc
Differentially expressed genes between NC PAC (high) and NC DMSO in ORG66:

```{r}
### following limma workflow
### difference between PAC and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_pacvsdmso_high)
fit <- treat(fit, lfc=log(1.2))
org66_ncPAC_high_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org66_ncPAC_high_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org66_ncPAC_high_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org66_ncPAC_high_DEgenes)))
up <- paste('• UPREGULATED:',sum(org66_ncPAC_high_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org66_ncPAC_high_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

##### NC_SN38-NC_DMSO {.tabset}

###### low_conc

Differentially expressed genes between NC SN38 (low) and NC DMSO in ORG66:

```{r}
### following limma workflow
### difference between SN38 and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_sn38vsdmso_low)
fit <- treat(fit, lfc=log(1.2))
org66_ncSN38_low_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org66_ncSN38_low_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org66_ncSN38_low_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org66_ncSN38_low_DEgenes)))
up <- paste('• UPREGULATED:',sum(org66_ncSN38_low_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org66_ncSN38_low_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

###### high_conc

Differentially expressed genes between NC SN38 (high) and NC DMSO in ORG66:

```{r}
### following limma workflow
### difference between SN38 and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, NC_sn38vsdmso_high)
fit <- treat(fit, lfc=log(1.2))
org66_ncSN38_high_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org66_ncSN38_high_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org66_ncSN38_high_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org66_ncSN38_high_DEgenes)))
up <- paste('• UPREGULATED:',sum(org66_ncSN38_high_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org66_ncSN38_high_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```


##### PAC_PAC - PAC_DMSO {.tabset}

###### low_conc

Differentially expressed genes between PAC_PAC (low) and PAC_DMSO in ORG66:

```{r}
### following limma workflow
### difference between PAC and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, PAC_pacvsdmso_low)
fit <- treat(fit, lfc=log(1.2))
all_genes <- topTreat(fit, number=Inf)
org66_paPAC_low_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org66_paPAC_low_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org66_paPAC_low_DEgenes)))
up <- paste('• UPREGULATED:',sum(org66_paPAC_low_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org66_paPAC_low_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

###### high_conc

Differentially expressed genes between PAC_PAC (high) and PAC_DMSO in ORG66:

```{r}
### following limma workflow
### difference between PAC and DMSO across concentrations
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, PAC_pacvsdmso_high)
fit <- treat(fit, lfc=log(1.2))
all_genes <- topTreat(fit, number=Inf)
org66_paPAC_high_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

#create_dt_genes(org66_paPAC_high_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org66_paPAC_high_DEgenes)))
up <- paste('• UPREGULATED:',sum(org66_paPAC_high_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org66_paPAC_high_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

##### SN38_SN38 - SN38_DMSO {.tabset}

###### low_conc

Differentially expressed genes between SN38_SN38 (low) and SN38_DMSO in ORG66:

```{r}
### following limma workflow
### difference between SN38 and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, SN38_sn38vsdmso_low)
fit <- treat(fit, lfc=log(1.2))
org66_snSN38_low_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org66_snSN38_low_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org66_snSN38_low_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org66_snSN38_low_DEgenes)))
up <- paste('• UPREGULATED:',sum(org66_snSN38_low_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org66_snSN38_low_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

###### high_conc

Differentially expressed genes between NC SN38 (high) and NC DMSO in ORG66:

```{r}
### following limma workflow
### difference between SN38 and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, SN38_sn38vsdmso_high)
fit <- treat(fit, lfc=log(1.2))
org66_snSN38_high_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org66_snSN38_high_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')

create_dt_genes(org66_snSN38_low_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org66_snSN38_high_DEgenes)))
up <- paste('• UPREGULATED:',sum(org66_snSN38_high_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org66_snSN38_high_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```


