---
title: "Geneset Enrichment Analysis using MSigDB"
output: html_document
date: "2026-03-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.height=4, fig.width = 6)
```

```{r}
library(tidyverse)
library(RColorBrewer)
library(fgsea)
library(msigdbr)
library(DT)
library(ggpubr)
```

```{r}
create_dt <-  function(x){
  datatable(x,
            extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
}
```

# Genesets

```{r}
#gene_sets_df <- msigdbr(species = 'Homo sapiens', category = 'C2')
gene_sets_df <- msigdbr(species = 'Homo sapiens', category = 'C2', subcollection = 'CP:REACTOME')
head(gene_sets_df)

# Convert to named list format required by fgsea
gene_sets <- gene_sets_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(gene_sets), "gene sets\n"))
```

# FGSEA Analysis

## Plate 2 {.tabset}

### ORG38 {.tabset}

#### PAC_PAC - PAC_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P2/ORG38_PACPAC.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org38_p2_pacpac_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org38_p2_pacpac_rankings) <- df$gene_symbol 

### plot rankings
org38_p2_pacpac_rankings <- sort(org38_p2_pacpac_rankings, decreasing = TRUE) 
plot(org38_p2_pacpac_rankings)

ggplot(data.frame(gene_symbol = names(org38_p2_pacpac_rankings[1:50]), ranks = org38_p2_pacpac_rankings[1:50]), aes(gene_symbol, ranks)) +
	geom_point() +
	theme_classic() +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------

org38_p2_pacpac_GSEAres <- fgsea(pathways = gene_sets, 
                 stats = org38_p2_pacpac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
```

GSEA Results:

```{r}
create_dt(org38_p2_pacpac_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org38_p2_pacpac_GSEAres[order(org38_p2_pacpac_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org38_p2_pacpac_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### SN38_SN38 - SN38_DMSO 

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P2/ORG38_SNSN38.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org38_p2_snsn38_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org38_p2_snsn38_rankings) <- df$gene_symbol 

### plot rankings
org38_p2_snsn38_rankings <- sort(org38_p2_snsn38_rankings, decreasing = TRUE) 
plot(org38_p2_snsn38_rankings)

ggplot(data.frame(gene_symbol = names(org38_p2_snsn38_rankings[1:50]), ranks = org38_p2_snsn38_rankings[1:50]), aes(gene_symbol, ranks)) +
	geom_point() +
	theme_classic() +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------

org38_p2_snsn38_GSEAres <- fgsea(pathways = gene_sets, 
                 stats = org38_p2_snsn38_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
```

GSEA Results:

```{r}
create_dt(org38_p2_snsn38_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org38_p2_snsn38_GSEAres[order(org38_p2_snsn38_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org38_p2_snsn38_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### PAC_DMSO - NC_DMSO 

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P2/ORG38_PDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org38_p2_pdmso_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org38_p2_pdmso_rankings) <- df$gene_symbol 

### plot rankings
org38_p2_pdmso_rankings <- sort(org38_p2_pdmso_rankings, decreasing = TRUE) 
plot(org38_p2_pdmso_rankings)

ggplot(data.frame(gene_symbol = names(org38_p2_pdmso_rankings[1:50]), ranks = org38_p2_pdmso_rankings[1:50]), aes(gene_symbol, ranks)) +
	geom_point() +
	theme_classic() +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------

org38_p2_pdmso_GSEAres <- fgsea(pathways = gene_sets, 
                 stats = org38_p2_pdmso_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
```

GSEA Results:

```{r}
create_dt(org38_p2_pdmso_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org38_p2_pdmso_GSEAres[order(org38_p2_pdmso_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org38_p2_pdmso_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### SN38_DMSO - NC_DMSO 

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P2/ORG38_SNDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org38_p2_sndmso_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org38_p2_sndmso_rankings) <- df$gene_symbol 

### plot rankings
org38_p2_sndmso_rankings <- sort(org38_p2_sndmso_rankings, decreasing = TRUE) 
plot(org38_p2_sndmso_rankings)

ggplot(data.frame(gene_symbol = names(org38_p2_sndmso_rankings[1:50]), ranks = org38_p2_sndmso_rankings[1:50]), aes(gene_symbol, ranks)) +
	geom_point() +
	theme_classic() +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------

org38_p2_sndmso_GSEAres <- fgsea(pathways = gene_sets, 
                 stats = org38_p2_sndmso_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
```

GSEA Results:

```{r}
create_dt(org38_p2_sndmso_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org38_p2_sndmso_GSEAres[order(org38_p2_sndmso_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org38_p2_sndmso_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

### ORG49 {.tabset}

#### PAC_PAC - PAC_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P2/ORG49_PACPAC.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org49_p2_pacpac_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org49_p2_pacpac_rankings) <- df$gene_symbol 

### plot rankings
org49_p2_pacpac_rankings <- sort(org49_p2_pacpac_rankings, decreasing = TRUE) 
plot(org49_p2_pacpac_rankings)

ggplot(data.frame(gene_symbol = names(org49_p2_pacpac_rankings[1:50]), ranks = org49_p2_pacpac_rankings[1:50]), aes(gene_symbol, ranks)) +
	geom_point() +
	theme_classic() +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------

org49_p2_pacpac_GSEAres <- fgsea(pathways = gene_sets, 
                 stats = rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
```

GSEA Results:

```{r}
create_dt(org49_p2_pacpac_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org49_p2_pacpac_GSEAres[order(org49_p2_pacpac_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org49_p2_pacpac_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### SN38_SN38 - SN38_DMSO 

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P2/ORG49_SNSN38.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org49_p2_snsn38_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org49_p2_snsn38_rankings) <- df$gene_symbol 

### plot rankings
org49_p2_snsn38_rankings <- sort(org49_p2_snsn38_rankings, decreasing = TRUE) 
plot(org49_p2_snsn38_rankings)

ggplot(data.frame(gene_symbol = names(org49_p2_snsn38_rankings[1:50]), ranks = org49_p2_snsn38_rankings[1:50]), aes(gene_symbol, ranks)) +
	geom_point() +
	theme_classic() +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------

org49_p2_snsn38_GSEAres <- fgsea(pathways = gene_sets, 
                 stats = org49_p2_snsn38_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
```

GSEA Results:

```{r}
create_dt(org49_p2_snsn38_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org49_p2_snsn38_GSEAres[order(org49_p2_snsn38_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org49_p2_snsn38_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### PAC_DMSO - NC_DMSO 

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P2/ORG49_PDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org49_p2_pdmso_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org49_p2_pdmso_rankings) <- df$gene_symbol 

### plot rankings
org49_p2_pdmso_rankings <- sort(org49_p2_pdmso_rankings, decreasing = TRUE) 
plot(org49_p2_pdmso_rankings)

ggplot(data.frame(gene_symbol = names(org49_p2_pdmso_rankings[1:50]), ranks = org49_p2_pdmso_rankings[1:50]), aes(gene_symbol, ranks)) +
	geom_point() +
	theme_classic() +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------

org49_p2_pdmso_GSEAres <- fgsea(pathways = gene_sets, 
                 stats = org49_p2_pdmso_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
```

GSEA Results:

```{r}
create_dt(org49_p2_pdmso_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org49_p2_pdmso_GSEAres[order(org49_p2_pdmso_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org49_p2_pdmso_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### SN38_DMSO - NC_DMSO 

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P2/ORG49_SNDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org49_p2_sndmso_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org49_p2_sndmso_rankings) <- df$gene_symbol 

### plot rankings
org49_p2_sndmso_rankings <- sort(org49_p2_sndmso_rankings, decreasing = TRUE) 
plot(org49_p2_sndmso_rankings)

ggplot(data.frame(gene_symbol = names(org49_p2_sndmso_rankings[1:50]), ranks = org49_p2_sndmso_rankings[1:50]), aes(gene_symbol, ranks)) +
	geom_point() +
	theme_classic() +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------

org49_p2_sndmso_GSEAres <- fgsea(pathways = gene_sets, 
                 stats = org49_p2_sndmso_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
```

GSEA Results:

```{r}
create_dt(org49_p2_sndmso_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org49_p2_sndmso_GSEAres[order(org49_p2_sndmso_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org49_p2_sndmso_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

### ORG66 {.tabset}

#### PAC_PAC - PAC_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P2/ORG66_PACPAC.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org66_p2_pacpac_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org66_p2_pacpac_rankings) <- df$gene_symbol 

### plot rankings
org66_p2_pacpac_rankings <- sort(org66_p2_pacpac_rankings, decreasing = TRUE) 
plot(org66_p2_pacpac_rankings)

ggplot(data.frame(gene_symbol = names(org66_p2_pacpac_rankings[1:50]), ranks = org66_p2_pacpac_rankings[1:50]), aes(gene_symbol, ranks)) +
	geom_point() +
	theme_classic() +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------

org66_p2_pacpac_GSEAres <- fgsea(pathways = gene_sets, 
                 stats = org66_p2_pacpac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
```

GSEA Results:

```{r}
create_dt(org66_p2_pacpac_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org66_p2_pacpac_GSEAres[order(org66_p2_pacpac_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org66_p2_pacpac_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### SN38_SN38 - SN38_DMSO 

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P2/ORG66_SNSN38.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org66_p2_snsn38_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org66_p2_snsn38_rankings) <- df$gene_symbol 

### plot rankings
org66_p2_snsn38_rankings <- sort(org66_p2_snsn38_rankings, decreasing = TRUE) 
plot(org66_p2_snsn38_rankings)

ggplot(data.frame(gene_symbol = names(org66_p2_snsn38_rankings[1:50]), ranks = org66_p2_snsn38_rankings[1:50]), aes(gene_symbol, ranks)) +
	geom_point() +
	theme_classic() +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------

org66_p2_snsn38_GSEAres <- fgsea(pathways = gene_sets, 
                 stats = org66_p2_snsn38_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
```

GSEA Results:

```{r}
create_dt(org66_p2_snsn38_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org66_p2_snsn38_GSEAres[order(org66_p2_snsn38_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org66_p2_snsn38_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### PAC_DMSO - NC_DMSO 

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P2/ORG66_PDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org66_p2_pdmso_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org66_p2_pdmso_rankings) <- df$gene_symbol 

### plot rankings
org66_p2_pdmso_rankings <- sort(org66_p2_pdmso_rankings, decreasing = TRUE) 
plot(org66_p2_pdmso_rankings)

ggplot(data.frame(gene_symbol = names(org66_p2_pdmso_rankings[1:50]), ranks = org66_p2_pdmso_rankings[1:50]), aes(gene_symbol, ranks)) +
	geom_point() +
	theme_classic() +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------

org66_p2_pdmso_GSEAres <- fgsea(pathways = gene_sets, 
                 stats = org66_p2_pdmso_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
```

GSEA Results:

```{r}
create_dt(org66_p2_pdmso_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org66_p2_pdmso_GSEAres[order(org66_p2_pdmso_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org66_p2_pdmso_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### SN38_DMSO - NC_DMSO 

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P2/ORG66_SNDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org66_p2_sndmso_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org66_p2_sndmso_rankings) <- df$gene_symbol 

### plot rankings
org66_p2_sndmso_rankings <- sort(org66_p2_sndmso_rankings, decreasing = TRUE) 
plot(org66_p2_sndmso_rankings)

ggplot(data.frame(gene_symbol = names(org66_p2_sndmso_rankings[1:50]), ranks = org66_p2_sndmso_rankings[1:50]), aes(gene_symbol, ranks)) +
	geom_point() +
	theme_classic() +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------

org66_p2_sndmso_GSEAres <- fgsea(pathways = gene_sets, 
                 stats = org66_p2_sndmso_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)
```

GSEA Results:

```{r}
create_dt(org66_p2_sndmso_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org66_p2_sndmso_GSEAres[order(org66_p2_sndmso_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org66_p2_sndmso_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

### All organoids {.tabset}

#### PAC_PAC - PAC_DMSO 

*Common pathways between ORG38 & ORG49:*

```{r}
list1 <- org38_p2_pacpac_GSEAres[org38_p2_pacpac_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_p2_pacpac_GSEAres[org49_p2_pacpac_GSEAres$padj<0.05,c('pathway')]
list3 <- org66_p2_pacpac_GSEAres[org66_p2_pacpac_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway, list2$pathway)
common_strings
```

#### SN38_SN38 - SN38_DMSO 

*Common pathways between ORG38 & ORG49:*

```{r}
list1 <- org38_p2_snsn38_GSEAres[org38_p2_snsn38_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_p2_snsn38_GSEAres[org49_p2_snsn38_GSEAres$padj<0.05,c('pathway')]
list3 <- org66_p2_snsn38_GSEAres[org66_p2_snsn38_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list2$pathway, list1$pathway)
common_strings
```

#### PAC_DMSO - NC_DMSO 

*Common pathways between ORG38 & ORG49:*

```{r}
list1 <- org38_p2_pdmso_GSEAres[org38_p2_pdmso_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_p2_pdmso_GSEAres[org49_p2_pdmso_GSEAres$padj<0.05,c('pathway')]
list3 <- org66_p2_pdmso_GSEAres[org66_p2_pdmso_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway, list2$pathway)
common_strings
```

#### SN38_DMSO - NC_DMSO 

*Common pathways between ORG38 & ORG66:*

```{r}
list1 <- org38_p2_sndmso_GSEAres[org38_p2_sndmso_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_p2_sndmso_GSEAres[org49_p2_sndmso_GSEAres$padj<0.05,c('pathway')]
list3 <- org66_p2_sndmso_GSEAres[org66_p2_sndmso_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway, list3$pathway)
common_strings
```

### Between Treatments {.tabset}

#### ORG38 Short term response 

```{r}
list1 <- org38_p2_pacpac_GSEAres[org38_p2_pacpac_GSEAres$padj<0.05,c('pathway')]
list2 <- org38_p2_snsn38_GSEAres[org38_p2_snsn38_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```

#### ORG38 Long term changes 

```{r}
list1 <- org38_p2_pdmso_GSEAres[org38_p2_pdmso_GSEAres$padj<0.05,c('pathway')]
list2 <- org38_p2_sndmso_GSEAres[org38_p2_sndmso_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```

#### ORG49 Short term response 

```{r}
list1 <- org49_p2_pacpac_GSEAres[org49_p2_pacpac_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_p2_snsn38_GSEAres[org49_p2_snsn38_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```


#### ORG49 Long term changes 

```{r}
list1 <- org49_p2_pdmso_GSEAres[org49_p2_pdmso_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_p2_sndmso_GSEAres[org49_p2_sndmso_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```

#### ORG66 Short term response 

```{r}
list1 <- org66_p2_pacpac_GSEAres[org66_p2_pacpac_GSEAres$padj<0.05,c('pathway')]
list2 <- org66_p2_snsn38_GSEAres[org66_p2_snsn38_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```

#### ORG66 Long term changes 

```{r}
list1 <- org66_p2_pdmso_GSEAres[org66_p2_pdmso_GSEAres$padj<0.05,c('pathway')]
list2 <- org66_p2_sndmso_GSEAres[org66_p2_sndmso_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```


## Plate 3 {.tabset}

### ORG38 {.tabset}

#### PAC_PAC - PAC_DMSO 

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P3/ORG38_paPAC.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
### FGSEA needs a named vector with the rankings and the genes as names
### We will use  signed p value as ranking: sign(df$log2fc)*(-log10(df$pval)))
# Note that you can also just use the log2FC only - but it will not take into account 
# genes with a large FC but non-significant

org38_p3_paPAC_rankings <- sign(df$logFC)*(-log10(df$P.Value)) # we will use the signed p values from spatial DGE as ranking
names(org38_p3_paPAC_rankings) <- df$gene_symbol # genes as names#

### plot rankings
org38_p3_paPAC_rankings <- sort(org38_p3_paPAC_rankings, decreasing = TRUE) # sort genes by ranking
plot(org38_p3_paPAC_rankings)

# # Some genes have such low p values that the signed pval is +- inf, we need to change it to the maximum * constant to avoid problems with fgsea
# max_ranking <- max(rankings[is.finite(rankings)])
# min_ranking <- min(rankings[is.finite(rankings)])
# rankings <- replace(rankings, rankings > max_ranking, max_ranking * 10)
# rankings <- replace(rankings, rankings < min_ranking, min_ranking * 10)
# rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking

ggplot(data.frame(gene_symbol = names(org38_p3_paPAC_rankings[1:50]), ranks = org38_p3_paPAC_rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
# Easy peasy! Run fgsea with the pathways 
org38_paPAC_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets, # List of gene sets to check
                 stats = org38_p3_paPAC_rankings,
                 scoreType = 'std', # in this case we have both pos and neg rankings. if only pos or neg, set to 'pos', 'neg'
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) # for parallelisation
)
```

GSEA Results:

```{r}
create_dt(org38_paPAC_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org38_paPAC_GSEAres[order(org38_paPAC_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org38_p3_paPAC_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### SN38_SN38 - SN38_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P3/org38_snSN38.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
### FGSEA needs a named vector with the rankings and the genes as names
### We will use  signed p value as ranking: sign(df$log2fc)*(-log10(df$pval)))
# Note that you can also just use the log2FC only - but it will not take into account 
# genes with a large FC but non-significant

org38_p3_snSN38_rankings <- sign(df$logFC)*(-log10(df$P.Value)) # we will use the signed p values from spatial DGE as ranking
names(org38_p3_snSN38_rankings) <- df$gene_symbol # genes as names#

### plot rankings
org38_p3_snSN38_rankings <- sort(org38_p3_snSN38_rankings, decreasing = TRUE) # sort genes by ranking
plot(org38_p3_snSN38_rankings)

# # Some genes have such low p values that the signed pval is +- inf, we need to change it to the maximum * constant to avoid problems with fgsea
# max_ranking <- max(rankings[is.finite(rankings)])
# min_ranking <- min(rankings[is.finite(rankings)])
# rankings <- replace(rankings, rankings > max_ranking, max_ranking * 10)
# rankings <- replace(rankings, rankings < min_ranking, min_ranking * 10)
# rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking

ggplot(data.frame(gene_symbol = names(org38_p3_snSN38_rankings[1:50]), ranks = org38_p3_snSN38_rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
# Easy peasy! Run fgsea with the pathways 
org38_snSN38_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = org38_p3_snSN38_rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org38_snSN38_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org38_snSN38_GSEAres[order(org38_snSN38_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org38_p3_snSN38_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### PAC_DMSO - NC_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P3/ORG38_PDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
### FGSEA needs a named vector with the rankings and the genes as names
### We will use  signed p value as ranking: sign(df$log2fc)*(-log10(df$pval)))
# Note that you can also just use the log2FC only - but it will not take into account 
# genes with a large FC but non-significant

org38_p3_pdmso_rankings <- sign(df$logFC)*(-log10(df$P.Value)) # we will use the signed p values from spatial DGE as ranking
names(org38_p3_pdmso_rankings) <- df$gene_symbol # genes as names#

### plot rankings
org38_p3_pdmso_rankings <- sort(org38_p3_pdmso_rankings, decreasing = TRUE) # sort genes by ranking
plot(org38_p3_pdmso_rankings)

# # Some genes have such low p values that the signed pval is +- inf, we need to change it to the maximum * constant to avoid problems with fgsea
# max_ranking <- max(rankings[is.finite(rankings)])
# min_ranking <- min(rankings[is.finite(rankings)])
# rankings <- replace(rankings, rankings > max_ranking, max_ranking * 10)
# rankings <- replace(rankings, rankings < min_ranking, min_ranking * 10)
# rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking

ggplot(data.frame(gene_symbol = names(org38_p3_pdmso_rankings[1:50]), ranks = org38_p3_pdmso_rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
# Easy peasy! Run fgsea with the pathways 
org38_PDMSO_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = org38_p3_pdmso_rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org38_PDMSO_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org38_PDMSO_GSEAres[order(org38_PDMSO_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org38_p3_pdmso_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### SN38_DMSO - NC_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P3/ORG38_SNDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org38_p3_sndmso_rankings <- sign(df$logFC)*(-log10(df$P.Value)) # we will use the signed p values from spatial DGE as ranking
names(org38_p3_sndmso_rankings) <- df$gene_symbol # genes as names#

### plot rankings
org38_p3_sndmso_rankings <- sort(org38_p3_sndmso_rankings, decreasing = TRUE) # sort genes by ranking
plot(org38_p3_sndmso_rankings)

ggplot(data.frame(gene_symbol = names(org38_p3_sndmso_rankings[1:50]), ranks = org38_p3_sndmso_rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org38_SNDMSO_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = org38_p3_sndmso_rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org38_SNDMSO_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org38_SNDMSO_GSEAres[order(org38_SNDMSO_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org38_p3_sndmso_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

### ORG49 {.tabset}

#### PAC_PAC - PAC_DMSO 

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P3/ORG49_paPAC.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org49_p3_paPAC_rankings <- sign(df$logFC)*(-log10(df$P.Value))
names(org49_p3_paPAC_rankings) <- df$gene_symbol

### plot rankings
org49_p3_paPAC_rankings <- sort(org49_p3_paPAC_rankings, decreasing = TRUE) # sort genes by ranking
plot(org49_p3_paPAC_rankings)

ggplot(data.frame(gene_symbol = names(org49_p3_paPAC_rankings[1:50]), ranks = org49_p3_paPAC_rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org49_paPAC_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = org49_p3_paPAC_rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org49_paPAC_GSEAres)
```

#### SN38_SN38 - SN38_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P3/org49_snSN38.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org49_p3_snSN38_rankings <- sign(df$logFC)*(-log10(df$P.Value))
names(org49_p3_snSN38_rankings) <- df$gene_symbol

### plot rankings
org49_p3_snSN38_rankings <- sort(org49_p3_snSN38_rankings, decreasing = TRUE) # sort genes by ranking
plot(org49_p3_snSN38_rankings)

ggplot(data.frame(gene_symbol = names(org49_p3_snSN38_rankings[1:50]), ranks = org49_p3_snSN38_rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org49_snSN38_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = org49_p3_snSN38_rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org49_snSN38_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org49_snSN38_GSEAres[order(org49_snSN38_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org49_p3_snSN38_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### PAC_DMSO - NC_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P3/ORG49_PDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)


### 2. Prepare ranked list of genes -----------------------------------------------
org49_p3_pdmso_rankings <- sign(df$logFC)*(-log10(df$P.Value))
names(org49_p3_pdmso_rankings) <- df$gene_symbol

### plot rankings
org49_p3_pdmso_rankings <- sort(org49_p3_pdmso_rankings, decreasing = TRUE) # sort genes by ranking
plot(org49_p3_pdmso_rankings)

ggplot(data.frame(gene_symbol = names(ranorg49_p3_pdmso_rankingskings[1:50]), ranks = org49_p3_pdmso_rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org49_PDMSO_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = org49_p3_pdmso_rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org49_PDMSO_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org49_PDMSO_GSEAres[order(org49_PDMSO_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org49_p3_pdmso_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)

```

#### SN38_DMSO - NC_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P3/ORG49_SNDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org49_p3_sndmso_rankings <- sign(df$logFC)*(-log10(df$P.Value))
names(org49_p3_sndmso_rankings) <- df$gene_symbol

### plot rankings
org49_p3_sndmso_rankings <- sort(org49_p3_sndmso_rankings, decreasing = TRUE) # sort genes by ranking
plot(org49_p3_sndmso_rankings)

ggplot(data.frame(gene_symbol = names(org49_p3_sndmso_rankings[1:50]), ranks = org49_p3_sndmso_rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org49_SNDMSO_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = org49_p3_sndmso_rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org49_SNDMSO_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org49_SNDMSO_GSEAres[order(org49_SNDMSO_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org49_p3_sndmso_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

### ORG66 {.tabset}

#### PAC_PAC - PAC_DMSO 

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P3/ORG66_paPAC.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org66_p3_paPAC_rankings <- sign(df$logFC)*(-log10(df$P.Value))
names(org66_p3_paPAC_rankings) <- df$gene_symbol

### plot rankings
org66_p3_paPAC_rankings <- sort(org66_p3_paPAC_rankings, decreasing = TRUE) # sort genes by ranking
plot(org66_p3_paPAC_rankings)

ggplot(data.frame(gene_symbol = names(org66_p3_paPAC_rankings[1:50]), ranks = org66_p3_paPAC_rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org66_paPAC_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = org66_p3_paPAC_rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org66_paPAC_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org66_paPAC_GSEAres[order(org66_paPAC_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org66_p3_paPAC_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### SN38_SN38 - SN38_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P3/org66_snSN38.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org66_p3_snSN38_rankings <- sign(df$logFC)*(-log10(df$P.Value))
names(org66_p3_snSN38_rankings) <- df$gene_symbol

### plot rankings
org66_p3_snSN38_rankings <- sort(org66_p3_snSN38_rankings, decreasing = TRUE) # sort genes by ranking
plot(org66_p3_snSN38_rankings)

ggplot(data.frame(gene_symbol = names(org66_p3_snSN38_rankings[1:50]), ranks = org66_p3_snSN38_rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org66_snSN38_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = org66_p3_snSN38_rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org66_snSN38_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org66_snSN38_GSEAres[order(org66_snSN38_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org66_p3_snSN38_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]],nrow=3, ncol=3)
```

#### PAC_DMSO - NC_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P3/ORG66_PDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)


### 2. Prepare ranked list of genes -----------------------------------------------
org66_p3_pdmso_rankings <- sign(df$logFC)*(-log10(df$P.Value))
names(org66_p3_pdmso_rankings) <- df$gene_symbol

### plot rankings
org66_p3_pdmso_rankings <- sort(org66_p3_pdmso_rankings, decreasing = TRUE) # sort genes by ranking
plot(org66_p3_pdmso_rankings)

ggplot(data.frame(gene_symbol = names(org66_p3_pdmso_rankings[1:50]), ranks = org66_p3_pdmso_rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org66_PDMSO_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = org66_p3_pdmso_rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org66_PDMSO_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org66_PDMSO_GSEAres[order(org66_PDMSO_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org66_p3_pdmso_rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### SN38_DMSO - NC_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P3/ORG66_SNDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org66_p3_sndmso_rankings <- sign(df$logFC)*(-log10(df$P.Value))
names(org66_p3_sndmso_rankings) <- df$gene_symbol

### plot rankings
org66_p3_sndmso_rankings <- sort(org66_p3_sndmso_rankings, decreasing = TRUE) # sort genes by ranking
plot(org66_p3_sndmso_rankings)

ggplot(data.frame(gene_symbol = names(org66_p3_sndmso_rankings[1:50]), ranks = org66_p3_sndmso_rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org66_SNDMSO_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = org66_p3_sndmso_rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org66_SNDMSO_GSEAres)
```

```{r, fig.height = 9, fig.width=20}
top10 <- org66_SNDMSO_GSEAres[order(org66_SNDMSO_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], org66_p3_sndmso_rankings) +
    labs(title = substr(pw, 10,10000000))
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

### All organoids {.tabset}

#### PAC_PAC - PAC_DMSO 

*Common pathways between ORG38 & ORG66:*

```{r}
list1 <- org38_paPAC_GSEAres[org38_paPAC_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_paPAC_GSEAres[org49_paPAC_GSEAres$padj<0.05,c('pathway')]
list3 <- org66_paPAC_GSEAres[org66_paPAC_GSEAres$padj<0.05,c('pathway')]

common_strings <- Reduce(intersect, list(list1$pathway, list3$pathway))
common_strings
```

#### SN38_SN38 - SN38_DMSO 

```{r}
list1 <- org38_snSN38_GSEAres[org38_snSN38_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_snSN38_GSEAres[org49_snSN38_GSEAres$padj<0.05,c('pathway')]
list3 <- org66_snSN38_GSEAres[org66_snSN38_GSEAres$padj<0.05,c('pathway')]

common_strings <- Reduce(intersect, list(list1$pathway, list2$pathway, list3$pathway))
common_strings
```

#### PAC_DMSO - NC_DMSO 

```{r}
list1 <- org38_PDMSO_GSEAres[org38_PDMSO_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_PDMSO_GSEAres[org49_PDMSO_GSEAres$padj<0.05,c('pathway')]
list3 <- org66_PDMSO_GSEAres[org66_PDMSO_GSEAres$padj<0.05,c('pathway')]
```

*Common pathways between ORG38 & ORG66:*

```{r}
common_strings <- Reduce(intersect, list(list1$pathway, list3$pathway))
common_strings
```

*Common pathways between ORG38 & ORG49:*

```{r}
common_strings <- Reduce(intersect, list(list1$pathway, list2$pathway))
common_strings
```

#### SN38_DMSO - NC_DMSO 

```{r}
list1 <- org38_SNDMSO_GSEAres[org38_SNDMSO_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_SNDMSO_GSEAres[org49_SNDMSO_GSEAres$padj<0.05,c('pathway')]
list3 <- org66_SNDMSO_GSEAres[org66_SNDMSO_GSEAres$padj<0.05,c('pathway')]

common_strings <- Reduce(intersect, list(list1$pathway,list2$pathway, list3$pathway))
common_strings
```

### Between Treatments {.tabset}

#### ORG38 Short term response 

```{r}
list1 <- org38_paPAC_GSEAres[org38_paPAC_GSEAres$padj<0.05,c('pathway')]
list2 <- org38_snSN38_GSEAres[org38_snSN38_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```

#### ORG38 Long term changes 

```{r}
list1 <- org38_PDMSO_GSEAres[org38_PDMSO_GSEAres$padj<0.05,c('pathway')]
list2 <- org38_SNDMSO_GSEAres[org38_SNDMSO_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```

#### ORG49 Short term response 

```{r}
list1 <- org49_paPAC_GSEAres[org49_paPAC_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_snSN38_GSEAres[org49_snSN38_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```


#### ORG49 Long term changes 

```{r}
list1 <- org49_PDMSO_GSEAres[org49_PDMSO_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_SNDMSO_GSEAres[org49_SNDMSO_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```

#### ORG66 Short term response 

```{r}
list1 <- org66_paPAC_GSEAres[org66_paPAC_GSEAres$padj<0.05,c('pathway')]
list2 <- org66_snSN38_GSEAres[org66_snSN38_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```

#### ORG66 Long term changes 

```{r}
list1 <- org66_PDMSO_GSEAres[org66_PDMSO_GSEAres$padj<0.05,c('pathway')]
list2 <- org66_SNDMSO_GSEAres[org66_SNDMSO_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```


## Comparing between T2 and T3 {.tabset}

### ORG38 {.tabset}

#### short-term PAC

```{r}
list1 <- org38_p2_pacpac_GSEAres[org38_p2_pacpac_GSEAres$padj<0.05,c('pathway')]
list2 <- org38_paPAC_GSEAres[org38_paPAC_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```


#### short-term SN38

```{r}
list1 <- org38_p2_snsn38_GSEAres[org38_p2_snsn38_GSEAres$padj<0.05,c('pathway')]
list2 <- org38_snSN38_GSEAres[org38_snSN38_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```

#### long-term PAC

```{r}
list1 <- org38_p2_pdmso_GSEAres[org38_p2_pdmso_GSEAres$padj<0.05,c('pathway')]
list2 <- org38_PDMSO_GSEAres[org38_PDMSO_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```

#### long-term SN38

```{r}
list1 <- org38_p2_sndmso_GSEAres[org38_p2_sndmso_GSEAres$padj<0.05,c('pathway')]
list2 <- org38_SNDMSO_GSEAres[org38_SNDMSO_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```

### ORG49 {.tabset}

#### short-term PAC

```{r}
list1 <- org49_p2_pacpac_GSEAres[org49_p2_pacpac_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_paPAC_GSEAres[org49_paPAC_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```


#### short-term SN38

```{r}
list1 <- org49_p2_snsn38_GSEAres[org49_p2_snsn38_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_snSN38_GSEAres[org49_snSN38_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```

#### long-term PAC

```{r}
list1 <- org49_p2_pdmso_GSEAres[org49_p2_pdmso_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_PDMSO_GSEAres[org49_PDMSO_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```

#### long-term SN38

```{r}
list1 <- org49_p2_sndmso_GSEAres[org49_p2_sndmso_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_SNDMSO_GSEAres[org49_SNDMSO_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```

### ORG66 {.tabset}

#### short-term PAC

```{r}
list1 <- org66_p2_pacpac_GSEAres[org66_p2_pacpac_GSEAres$padj<0.05,c('pathway')]
list2 <- org66_paPAC_GSEAres[org66_paPAC_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```


#### short-term SN38

```{r}
list1 <- org66_p2_snsn38_GSEAres[org66_p2_snsn38_GSEAres$padj<0.05,c('pathway')]
list2 <- org66_snSN38_GSEAres[org66_snSN38_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```

#### long-term PAC

```{r}
list1 <- org66_p2_pdmso_GSEAres[org66_p2_pdmso_GSEAres$padj<0.05,c('pathway')]
list2 <- org66_PDMSO_GSEAres[org66_PDMSO_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```

#### long-term SN38

```{r}
list1 <- org66_p2_sndmso_GSEAres[org66_p2_sndmso_GSEAres$padj<0.05,c('pathway')]
list2 <- org66_SNDMSO_GSEAres[org66_SNDMSO_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1$pathway,list2$pathway)
common_strings
```

### All organoids {.tabset}

#### short-term pac

```{r}
list1 <- org38_p2_pacpac_GSEAres[org38_p2_pacpac_GSEAres$padj<0.05,c('pathway')]
list2 <- org38_paPAC_GSEAres[org38_paPAC_GSEAres$padj<0.05,c('pathway')]
list3 <- org49_p2_pacpac_GSEAres[org49_p2_pacpac_GSEAres$padj<0.05,c('pathway')]
list4 <- org49_paPAC_GSEAres[org49_paPAC_GSEAres$padj<0.05,c('pathway')]
list5 <- org66_p2_pacpac_GSEAres[org66_p2_pacpac_GSEAres$padj<0.05,c('pathway')]
list6 <- org66_paPAC_GSEAres[org66_paPAC_GSEAres$padj<0.05,c('pathway')]

common_strings <- Reduce(intersect, list(list1$pathway,list2$pathway, list3$pathway, list4$pathway, list5$pathway, list6$pathway))
common_strings
```

#### short-term SN38

```{r}
list1 <- org38_p2_snsn38_GSEAres[org38_p2_snsn38_GSEAres$padj<0.05,c('pathway')]
list2 <- org38_snSN38_GSEAres[org38_snSN38_GSEAres$padj<0.05,c('pathway')]
list3 <- org49_p2_snsn38_GSEAres[org49_p2_snsn38_GSEAres$padj<0.05,c('pathway')]
list4 <- org49_snSN38_GSEAres[org49_snSN38_GSEAres$padj<0.05,c('pathway')]
list5 <- org66_p2_snsn38_GSEAres[org66_p2_snsn38_GSEAres$padj<0.05,c('pathway')]
list6 <- org66_snSN38_GSEAres[org66_snSN38_GSEAres$padj<0.05,c('pathway')]

common_strings <- Reduce(intersect, list(list1$pathway,list2$pathway, list3$pathway, list4$pathway, list5$pathway, list6$pathway))
common_strings
```

#### long-term pac

```{r}
list1 <- org38_p2_pdmso_GSEAres[org38_p2_pdmso_GSEAres$padj<0.05,c('pathway')]
list2 <- org38_PDMSO_GSEAres[org38_PDMSO_GSEAres$padj<0.05,c('pathway')]
list3 <- org49_p2_pdmso_GSEAres[org49_p2_pdmso_GSEAres$padj<0.05,c('pathway')]
list4 <- org49_PDMSO_GSEAres[org49_PDMSO_GSEAres$padj<0.05,c('pathway')]
list5 <- org66_p2_pdmso_GSEAres[org66_p2_pdmso_GSEAres$padj<0.05,c('pathway')]
list6 <- org66_PDMSO_GSEAres[org66_PDMSO_GSEAres$padj<0.05,c('pathway')]

common_strings <- Reduce(intersect, list(list1$pathway,list2$pathway, list3$pathway, list4$pathway, list5$pathway, list6$pathway))
common_strings
```

#### long-term SN38

```{r}
list1 <- org38_p2_sndmso_GSEAres[org38_p2_sndmso_GSEAres$padj<0.05,c('pathway')]
list2 <- org38_SNDMSO_GSEAres[org38_SNDMSO_GSEAres$padj<0.05,c('pathway')]
list3 <- org49_p2_sndmso_GSEAres[org49_p2_sndmso_GSEAres$padj<0.05,c('pathway')]
list4 <- org49_SNDMSO_GSEAres[org49_SNDMSO_GSEAres$padj<0.05,c('pathway')]
list5 <- org66_p2_sndmso_GSEAres[org66_p2_sndmso_GSEAres$padj<0.05,c('pathway')]
list6 <- org66_SNDMSO_GSEAres[org66_SNDMSO_GSEAres$padj<0.05,c('pathway')]

common_strings <- Reduce(intersect, list(list1$pathway,list2$pathway, list3$pathway, list4$pathway, list5$pathway, list6$pathway))
common_strings
```

## Plotting

### SN38_DMSO - NC_DMSO (T2 vs T3) {.tabset}

For pathways associated with long-term SN38 treatement (SN38_DMSO - NC_DMSO) that are enriched in all organoids across both T2 and T3: 

#### ORG38

Activation of gene expression by SREBF/SREBP:

```{r}
p1 <- plotEnrichment(gene_sets[['REACTOME_ACTIVATION_OF_GENE_EXPRESSION_BY_SREBF_SREBP']],
               org38_p2_snsn38_rankings)

p2 <- plotEnrichment(gene_sets[['REACTOME_ACTIVATION_OF_GENE_EXPRESSION_BY_SREBF_SREBP']],
               org38_p3_snSN38_rankings) 

p3 <- ggarrange(p1, p2, nrow=1, ncol=2)

annotate_figure(p3, top = text_grob("ACTIVATION OF GENE EXPRESSION \nBY SREBF/SREBP (T2 VS T3)", 
               color = "black", face = "bold", size = 16))
```

Metabolism of steroids:

```{r}
p1 <- plotEnrichment(gene_sets[['REACTOME_METABOLISM_OF_STEROIDS']],
               org38_p2_snsn38_rankings) 

p2 <- plotEnrichment(gene_sets[['REACTOME_METABOLISM_OF_STEROIDS']],
               org38_p3_snSN38_rankings) 

p3 <- ggarrange(p1, p2, nrow=1, ncol=2)

annotate_figure(p3, top = text_grob("METABOLISM OF STEROIDS (T2 VS T3)", 
               color = "black", face = "bold", size = 16))
```

Regulation of cholesterol biosynthesis by SREBP/SREBF:

```{r}
p1 <- plotEnrichment(gene_sets[['REACTOME_REGULATION_OF_CHOLESTEROL_BIOSYNTHESIS_BY_SREBP_SREBF']],
               org38_p2_snsn38_rankings) 

p2 <- plotEnrichment(gene_sets[['REACTOME_REGULATION_OF_CHOLESTEROL_BIOSYNTHESIS_BY_SREBP_SREBF']],
               org38_p3_snSN38_rankings) 

p3 <- ggarrange(p1, p2, nrow=1, ncol=2)

annotate_figure(p3, top = text_grob("REGULATION OF CHOLESTEROL BIOSYNTHESIS \nBY SREBP/SREBF (T2 VS T3)", 
               color = "black", face = "bold", size = 16))
```

#### ORG49

Activation of gene expression by SREBF/SREBP:

```{r}
p1 <- plotEnrichment(gene_sets[['REACTOME_ACTIVATION_OF_GENE_EXPRESSION_BY_SREBF_SREBP']],
               org49_p2_snsn38_rankings)

p2 <- plotEnrichment(gene_sets[['REACTOME_ACTIVATION_OF_GENE_EXPRESSION_BY_SREBF_SREBP']],
               org49_p3_snSN38_rankings)

p3 <- ggarrange(p1, p2, nrow=1, ncol=2)

annotate_figure(p3, top = text_grob("ACTIVATION OF GENE EXPRESSION \nBY SREBF/SREBP (T2 VS T3)", 
               color = "black", face = "bold", size = 16))
```

Metabolism of steroids:

```{r}
p1 <- plotEnrichment(gene_sets[['REACTOME_METABOLISM_OF_STEROIDS']],
               org49_p2_snsn38_rankings) 

p2 <- plotEnrichment(gene_sets[['REACTOME_METABOLISM_OF_STEROIDS']],
               org49_p3_snSN38_rankings)

p3 <- ggarrange(p1, p2, nrow=1, ncol=2)

annotate_figure(p3, top = text_grob("METABOLISM OF STEROIDS (T2 VS T3)", 
               color = "black", face = "bold", size = 16))
```

Regulation of cholesterol biosynthesis by SREBP/SREBF:

```{r}
p1 <- plotEnrichment(gene_sets[['REACTOME_REGULATION_OF_CHOLESTEROL_BIOSYNTHESIS_BY_SREBP_SREBF']],
               org49_p2_snsn38_rankings) 

p2 <- plotEnrichment(gene_sets[['REACTOME_REGULATION_OF_CHOLESTEROL_BIOSYNTHESIS_BY_SREBP_SREBF']],
               org49_p3_snSN38_rankings) 

p3 <- ggarrange(p1, p2, nrow=1, ncol=2)

annotate_figure(p3, top = text_grob("REGULATION OF CHOLESTEROL BIOSYNTHESIS \nBY SREBP/SREBF (T2 VS T3)", 
               color = "black", face = "bold", size = 16))
```


#### ORG66

Activation of gene expression by SREBF/SREBP:

```{r}
p1 <- plotEnrichment(gene_sets[['REACTOME_ACTIVATION_OF_GENE_EXPRESSION_BY_SREBF_SREBP']],
               org66_p2_snsn38_rankings) 

p2 <- plotEnrichment(gene_sets[['REACTOME_ACTIVATION_OF_GENE_EXPRESSION_BY_SREBF_SREBP']],
               org66_p3_snSN38_rankings) 

p3 <- ggarrange(p1, p2, nrow=1, ncol=2)

annotate_figure(p3, top = text_grob("ACTIVATION OF GENE EXPRESSION \nBY SREBF/SREBP (T2 VS T3)", 
               color = "black", face = "bold", size = 16))
```

Metabolism of steroids:

```{r}
p1 <- plotEnrichment(gene_sets[['REACTOME_METABOLISM_OF_STEROIDS']],
               org66_p2_snsn38_rankings) 

p2 <- plotEnrichment(gene_sets[['REACTOME_METABOLISM_OF_STEROIDS']],
               org66_p3_snSN38_rankings) 

p3 <- ggarrange(p1, p2, nrow=1, ncol=2)

annotate_figure(p3, top = text_grob("METABOLISM OF STEROIDS (T2 VS T3)", 
               color = "black", face = "bold", size = 16))
```

Regulation of cholesterol biosynthesis by SREBP/SREBF:

```{r}
p1 <- plotEnrichment(gene_sets[['REACTOME_REGULATION_OF_CHOLESTEROL_BIOSYNTHESIS_BY_SREBP_SREBF']],
               org66_p2_snsn38_rankings) 

p2 <- plotEnrichment(gene_sets[['REACTOME_REGULATION_OF_CHOLESTEROL_BIOSYNTHESIS_BY_SREBP_SREBF']],
               org66_p3_snSN38_rankings)

p3 <- ggarrange(p1, p2, nrow=1, ncol=2)

annotate_figure(p3, top = text_grob("REGULATION OF CHOLESTEROL BIOSYNTHESIS \nBY SREBP/SREBF (T2 VS T3)", 
               color = "black", face = "bold", size = 16))
```

```{r}
### Code for counting elading edge genes
library(purrr)


leading_edges <- lapply(org38_SNDMSO_GSEAres[org38_SNDMSO_GSEAres$padj< 0.01, ]$leadingEdge, unlist)

# count how often genes appear
gene_counts <- table(unlist(leading_edges))

# top recurring genes
head(sort(gene_counts, decreasing = TRUE), 20)

```

# Filtering apthways 

## P2

### ORG38 {.tabset}

#### PAC_PAC 

```{r}
collapsed <- fgsea::collapsePathways(org38_p2_pacpac_GSEAres[order(pval)][padj < 0.1], gene_sets, org38_p2_pacpac_rankings)
org38_p2_pacpac_GSEAres_main <- org38_p2_pacpac_GSEAres[org38_p2_pacpac_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org38_p2_pacpac_GSEAres_main)
```

#### SN38_SN38

```{r}
collapsed <- fgsea::collapsePathways(org38_p2_snsn38_GSEAres[order(pval)][padj < 0.1], gene_sets, org38_p2_snsn38_rankings)
org38_p2_snsn38_GSEAres_main <- org38_p2_snsn38_GSEAres[org38_p2_snsn38_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org38_p2_snsn38_GSEAres_main)
```

#### PAC_DMSO - NC_DMSO

```{r}
collapsed <- fgsea::collapsePathways(org38_p2_pdmso_GSEAres[order(pval)][padj < 0.1], gene_sets, org38_p2_pdmso_rankings)
org38_p2_pdmso_GSEAres_main <- org38_p2_pdmso_GSEAres[org38_p2_pdmso_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org38_p2_pdmso_GSEAres_main)
```

#### SN38_DMSO - NC_DMSO

```{r}
collapsed <- fgsea::collapsePathways(org38_p2_sndmso_GSEAres[order(pval)][padj < 0.1], gene_sets, org38_p2_sndmso_rankings)
org38_p2_sndmso_GSEAres_main <- org38_p2_sndmso_GSEAres[org38_p2_sndmso_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org38_p2_sndmso_GSEAres)
```

### ORG49 {.tabset}

#### PAC_PAC 

```{r}
collapsed <- fgsea::collapsePathways(org49_p2_pacpac_GSEAres[order(pval)][padj < 0.1], gene_sets, org49_p2_pacpac_rankings)
org49_p2_pacpac_GSEAres_main <- org49_p2_pacpac_GSEAres[org49_p2_pacpac_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org49_p2_pacpac_GSEAres_main)
```

#### SN38_SN38

```{r}
collapsed <- fgsea::collapsePathways(org49_p2_snsn38_GSEAres[order(pval)][padj < 0.1], gene_sets, org49_p2_snsn38_rankings)
org49_p2_snsn38_GSEAres_main <- org49_p2_snsn38_GSEAres[org49_p2_snsn38_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org49_p2_snsn38_GSEAres_main)
```

#### PAC_DMSO - NC_DMSO

```{r}
collapsed <- fgsea::collapsePathways(org49_p2_pdmso_GSEAres[order(pval)][padj < 0.1], gene_sets, org49_p2_pdmso_rankings)
org49_p2_pdmso_GSEAres_main <- org49_p2_pdmso_GSEAres[org49_p2_pdmso_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org49_p2_pdmso_GSEAres_main)
```

#### SN38_DMSO - NC_DMSO

```{r}
collapsed <- fgsea::collapsePathways(org49_p2_sndmso_GSEAres[order(pval)][padj < 0.1], gene_sets, org49_p2_sndmso_rankings)
org49_p2_sndmso_GSEAres_main <- org49_p2_sndmso_GSEAres[org49_p2_sndmso_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org49_p2_sndmso_GSEAres)
```


### ORG66 {.tabset}

#### PAC_PAC 

```{r}
collapsed <- fgsea::collapsePathways(org66_p2_pacpac_GSEAres[order(pval)][padj < 0.1], gene_sets, org66_p2_pacpac_rankings)
org66_p2_pacpac_GSEAres_main <- org66_p2_pacpac_GSEAres[org66_p2_pacpac_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org66_p2_pacpac_GSEAres_main)
```

#### SN38_SN38

```{r}
collapsed <- fgsea::collapsePathways(org66_p2_snsn38_GSEAres[order(pval)][padj < 0.1], gene_sets, org66_p2_snsn38_rankings)
org66_p2_snsn38_GSEAres_main <- org66_p2_snsn38_GSEAres[org66_p2_snsn38_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org66_p2_snsn38_GSEAres_main)
```

#### PAC_DMSO - NC_DMSO

```{r}
collapsed <- fgsea::collapsePathways(org66_p2_pdmso_GSEAres[order(pval)][padj < 0.1], gene_sets, org66_p2_pdmso_rankings)
org66_p2_pdmso_GSEAres_main <- org66_p2_pdmso_GSEAres[org66_p2_pdmso_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org66_p2_pdmso_GSEAres_main)
```

#### SN38_DMSO - NC_DMSO

```{r}
collapsed <- fgsea::collapsePathways(org66_p2_sndmso_GSEAres[order(pval)][padj < 0.1], gene_sets, org66_p2_sndmso_rankings)
org66_p2_sndmso_GSEAres_main <- org66_p2_sndmso_GSEAres[org66_p2_sndmso_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org66_p2_sndmso_GSEAres)
```


## P3

### ORG38 {.tabset}

#### PAC_PAC 

```{r}
collapsed <- fgsea::collapsePathways(org38_paPAC_GSEAres[order(pval)][padj < 0.1], gene_sets, org38_p3_paPAC_rankings)
org38_paPAC_GSEAres_main <- org38_paPAC_GSEAres[org38_paPAC_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org38_paPAC_GSEAres_main)
```

#### SN38_SN38

```{r}
collapsed <- fgsea::collapsePathways(org38_snSN38_GSEAres[order(pval)][padj < 0.1], gene_sets, org38_p3_snSN38_rankings)
org38_snSN38_GSEAres_main <- org38_snSN38_GSEAres[org38_snSN38_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org38_snSN38_GSEAres_main)
```

#### PAC_DMSO - NC_DMSO

```{r}
collapsed <- fgsea::collapsePathways(org38_PDMSO_GSEAres[order(pval)][padj < 0.1], gene_sets, org38_p3_pdmso_rankings)
org38_PDMSO_GSEAres_main <- org38_PDMSO_GSEAres[org38_PDMSO_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org38_PDMSO_GSEAres_main)
```

#### SN38_DMSO - NC_DMSO

```{r}
collapsed <- fgsea::collapsePathways(org38_SNDMSO_GSEAres[order(pval)][padj < 0.1], gene_sets, org38_p3_sndmso_rankings)
org38_SNDMSO_GSEAres_main <- org38_SNDMSO_GSEAres[org38_SNDMSO_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org38_SNDMSO_GSEAres_main)
```

### ORG49 {.tabset}

#### PAC_PAC 

```{r}
collapsed <- fgsea::collapsePathways(org49_paPAC_GSEAres[order(pval)][padj < 0.1], gene_sets, org49_p3_paPAC_rankings)
org49_paPAC_GSEAres_main <- org49_paPAC_GSEAres[org49_paPAC_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org49_paPAC_GSEAres_main)
```

#### SN38_SN38

```{r}
collapsed <- fgsea::collapsePathways(org49_snSN38_GSEAres[order(pval)][padj < 0.1], gene_sets, org49_p3_snSN38_rankings)
org49_snSN38_GSEAres_main <- org49_snSN38_GSEAres[org49_snSN38_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org49_snSN38_GSEAres_main)
```

#### PAC_DMSO - NC_DMSO

```{r}
collapsed <- fgsea::collapsePathways(org49_PDMSO_GSEAres[order(pval)][padj < 0.1], gene_sets, org49_p3_pdmso_rankings)
org49_PDMSO_GSEAres_main <- org49_PDMSO_GSEAres[org49_PDMSO_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org49_PDMSO_GSEAres_main)
```

#### SN38_DMSO - NC_DMSO

```{r}
collapsed <- fgsea::collapsePathways(org49_SNDMSO_GSEAres[order(pval)][padj < 0.1], gene_sets, org49_p3_sndmso_rankings)
org49_SNDMSO_GSEAres_main <- org49_SNDMSO_GSEAres[org49_SNDMSO_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org49_SNDMSO_GSEAres_main)
```

### ORG66 {.tabset}

#### PAC_PAC 

```{r}
collapsed <- fgsea::collapsePathways(org66_paPAC_GSEAres[order(pval)][padj < 0.1], gene_sets, org66_p3_paPAC_rankings)
org66_paPAC_GSEAres_main <- org66_paPAC_GSEAres[org66_paPAC_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org66_paPAC_GSEAres_main)
```

#### SN38_SN38

```{r}
collapsed <- fgsea::collapsePathways(org66_snSN38_GSEAres[order(pval)][padj < 0.1], gene_sets, org66_p3_snSN38_rankings)
org66_snSN38_GSEAres_main <- org66_snSN38_GSEAres[org66_snSN38_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org66_snSN38_GSEAres_main)
```

#### PAC_DMSO - NC_DMSO

```{r}
collapsed <- fgsea::collapsePathways(org66_PDMSO_GSEAres[order(pval)][padj < 0.1], gene_sets, org66_p3_pdmso_rankings)
org66_PDMSO_GSEAres_main <- org66_PDMSO_GSEAres[org66_PDMSO_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org66_PDMSO_GSEAres_main)
```

#### SN38_DMSO - NC_DMSO

```{r}
collapsed <- fgsea::collapsePathways(org66_SNDMSO_GSEAres[order(pval)][padj < 0.1], gene_sets, org66_p3_sndmso_rankings)
org66_SNDMSO_GSEAres_main <- org66_SNDMSO_GSEAres[org66_SNDMSO_GSEAres$pathway %in% collapsed$mainPathways, ]

create_dt(org66_SNDMSO_GSEAres_main)
```

## Building pathway/NES matrix {.tabset}

```{r}
### Extract NES - fill missing with 0
get_NES <- function(res, pathways) {
  out <- setNames(rep(0, length(pathways)), pathways)
  match_idx <- match(res$pathway, pathways)
  out[match_idx[!is.na(match_idx)]] <- res$NES[!is.na(match_idx)]
  return(out)
}

col_palette <- colorRampPalette(brewer.pal(11, "RdBu"))(100)
```

### PAC-PAC - PAC_DMSO {.tabset}

#### T2

```{r, fig.height = 10, fig.width = 10}
### select top pathways from collapsed dataframes
top38 <- org38_p2_pacpac_GSEAres_main[order(org38_p2_pacpac_GSEAres_main$padj)][1:10, ]
top49 <- org49_p2_pacpac_GSEAres_main[order(org49_p2_pacpac_GSEAres_main$padj)][1:10, ]
top66 <- org66_p2_pacpac_GSEAres_main[order(org66_p2_pacpac_GSEAres_main$padj)][1:10, ]

### get union of pathways
all_pathways <- unique(c(
  top38$pathway,
  top49$pathway,
  top66$pathway
))

NES_matrix <- cbind(
  org38 = get_NES(org38_p2_pacpac_GSEAres_main, all_pathways),
  org49 = get_NES(org49_p2_pacpac_GSEAres_main, all_pathways),
  org66 = get_NES(org66_p2_pacpac_GSEAres_main, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette)
)
```

#### T3

```{r, fig.height = 10, fig.width = 10}
### select top pathways from collapsed dataframes
top38 <- org38_paPAC_GSEAres_main[order(org38_paPAC_GSEAres_main$padj)][1:10, ]
top49 <- org49_paPAC_GSEAres_main[order(org49_paPAC_GSEAres_main$padj)][1:10, ]
top66 <- org66_paPAC_GSEAres_main[order(org66_paPAC_GSEAres_main$padj)][1:10, ]

### get union of pathways
all_pathways <- unique(c(
  top38$pathway,
  top49$pathway,
  top66$pathway
))

NES_matrix <- cbind(
  org38 = get_NES(org38_paPAC_GSEAres_main, all_pathways),
  org49 = get_NES(org49_paPAC_GSEAres_main, all_pathways),
  org66 = get_NES(org66_paPAC_GSEAres_main, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette)
)
```

### SN38_SN38 - SN38_DMSO {.tabset}

#### T2

```{r, fig.height = 10, fig.width = 15}
### select top pathways from collapsed dataframes
top38 <- org38_p2_snsn38_GSEAres_main[order(org38_p2_snsn38_GSEAres_main$padj)][1:20, ]
top49 <- org49_p2_snsn38_GSEAres_main[order(org49_p2_snsn38_GSEAres_main$padj)][1:20, ]
top66 <- org66_p2_snsn38_GSEAres_main[order(org66_p2_snsn38_GSEAres_main$padj)][1:20, ]

### get union of pathways
all_pathways <- unique(c(
  top38$pathway,
  top49$pathway,
  top66$pathway
))

NES_matrix <- cbind(
  org38 = get_NES(org38_p2_snsn38_GSEAres_main, all_pathways),
  org49 = get_NES(org49_p2_snsn38_GSEAres_main, all_pathways),
  org66 = get_NES(org66_p2_snsn38_GSEAres_main, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette)
)
```
```{r, fig.height = 10, fig.width = 15}
### select top pathways from collapsed dataframes
top38 <- org38_p2_snsn38_GSEAres[order(org38_p2_snsn38_GSEAres$padj)][1:15, ]
top49 <- org49_p2_snsn38_GSEAres[order(org49_p2_snsn38_GSEAres$padj)][1:15, ]
top66 <- org66_p2_snsn38_GSEAres[order(org66_p2_snsn38_GSEAres$padj)][1:15, ]

### get union of pathways
all_pathways <- unique(c(
  top38$pathway,
  top49$pathway,
  top66$pathway
))

NES_matrix <- cbind(
  org38 = get_NES(org38_p2_snsn38_GSEAres, all_pathways),
  org49 = get_NES(org49_p2_snsn38_GSEAres, all_pathways),
  org66 = get_NES(org66_p2_snsn38_GSEAres, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette),
  border_color = 'NA'
)
```

#### T3

```{r, fig.height = 10, fig.width = 15}
### select top pathways from collapsed dataframes
top38 <- org38_snSN38_GSEAres_main[order(org38_snSN38_GSEAres_main$padj)][1:20, ]
top49 <- org49_snSN38_GSEAres_main[order(org49_snSN38_GSEAres_main$padj)][1:20, ]
top66 <- org66_snSN38_GSEAres_main[order(org66_snSN38_GSEAres_main$padj)][1:20, ]

### get union of pathways
all_pathways <- unique(c(
  top38$pathway,
  top49$pathway,
  top66$pathway
))

NES_matrix <- cbind(
  org38 = get_NES(org38_snSN38_GSEAres_main, all_pathways),
  org49 = get_NES(org49_snSN38_GSEAres_main, all_pathways),
  org66 = get_NES(org66_snSN38_GSEAres_main, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette)
)
```

```{r, fig.height = 10, fig.width = 15}
### select top pathways from collapsed dataframes
top38 <- org38_snSN38_GSEAres[order(org38_snSN38_GSEAres$padj)][1:15, ]
top49 <- org49_snSN38_GSEAres[order(org49_snSN38_GSEAres$padj)][1:15, ]
top66 <- org66_snSN38_GSEAres[order(org66_snSN38_GSEAres$padj)][1:15, ]

### get union of pathways
all_pathways <- unique(c(
  top38$pathway,
  top49$pathway,
  top66$pathway
))

NES_matrix <- cbind(
  org38 = get_NES(org38_snSN38_GSEAres, all_pathways),
  org49 = get_NES(org49_snSN38_GSEAres, all_pathways),
  org66 = get_NES(org66_snSN38_GSEAres, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette),
  border_color = 'NA'
)
```

### PAC_DMSO - NC_DMSO {.tabset}

#### T2

```{r, fig.height = 10, fig.width = 15}
### select top pathways from collapsed dataframes
top38 <- org38_p2_pdmso_GSEAres_main[order(org38_p2_pdmso_GSEAres_main$padj)][1:10, ]
top49 <- org49_p2_pdmso_GSEAres_main[order(org49_p2_pdmso_GSEAres_main$padj)][1:10, ]
top66 <- org66_p2_pdmso_GSEAres_main[order(org66_p2_pdmso_GSEAres_main$padj)][1:10, ]

### get union of pathways
all_pathways <- unique(c(
  top38$pathway,
  top49$pathway,
  top66$pathway
))

NES_matrix <- cbind(
  org38 = get_NES(org38_p2_pdmso_GSEAres_main, all_pathways),
  org49 = get_NES(org49_p2_pdmso_GSEAres_main, all_pathways),
  org66 = get_NES(org66_p2_pdmso_GSEAres_main, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette)
)
```

#### T3

```{r, fig.height = 10, fig.width = 15}
### select top pathways from collapsed dataframes
top38 <- org38_PDMSO_GSEAres_main[order(org38_PDMSO_GSEAres_main$padj)][1:10, ]
top49 <- org49_PDMSO_GSEAres_main[order(org49_PDMSO_GSEAres_main$padj)][1:10, ]
top66 <- org66_PDMSO_GSEAres_main[order(org66_PDMSO_GSEAres_main$padj)][1:10, ]

### get union of pathways
all_pathways <- unique(c(
  top38$pathway,
  top49$pathway,
  top66$pathway
))

NES_matrix <- cbind(
  org38 = get_NES(org38_PDMSO_GSEAres_main, all_pathways),
  org49 = get_NES(org49_PDMSO_GSEAres_main, all_pathways),
  org66 = get_NES(org66_PDMSO_GSEAres_main, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette)
)
```

### SN38_DMSO - NC_DMSO {.tabset}

#### T2

```{r, fig.height = 10, fig.width = 15}
### select top pathways from collapsed dataframes
top38 <- org38_p2_sndmso_GSEAres_main[order(org38_p2_sndmso_GSEAres_main$padj)][1:20, ]
top49 <- org49_p2_sndmso_GSEAres_main[order(org49_p2_sndmso_GSEAres_main$padj)][1:20, ]
top66 <- org66_p2_sndmso_GSEAres_main[order(org66_p2_sndmso_GSEAres_main$padj)][1:20, ]

### get union of pathways
all_pathways <- unique(c(
  top38$pathway,
  top49$pathway,
  top66$pathway
))

NES_matrix <- cbind(
  org38 = get_NES(org38_p2_sndmso_GSEAres_main, all_pathways),
  org49 = get_NES(org49_p2_sndmso_GSEAres_main, all_pathways),
  org66 = get_NES(org66_p2_sndmso_GSEAres_main, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette)
)
```
```{r, fig.height = 10, fig.width = 15}
### select top pathways from collapsed dataframes
top38 <- org38_p2_sndmso_GSEAres[order(org38_p2_sndmso_GSEAres$padj)][1:10, ]
top49 <- org49_p2_sndmso_GSEAres[order(org49_p2_sndmso_GSEAres$padj)][1:10, ]
top66 <- org66_p2_sndmso_GSEAres[order(org66_p2_sndmso_GSEAres$padj)][1:10, ]

### get union of pathways
all_pathways <- unique(c(
  top38$pathway,
  top49$pathway,
  top66$pathway
))

NES_matrix <- cbind(
  org38 = get_NES(org38_p2_sndmso_GSEAres, all_pathways),
  org49 = get_NES(org49_p2_sndmso_GSEAres, all_pathways),
  org66 = get_NES(org66_p2_sndmso_GSEAres, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette),
  border_color = 'NA'
)
```

#### T3

```{r, fig.height = 10, fig.width = 15}
### select top pathways from collapsed dataframes
top38 <- org38_SNDMSO_GSEAres_main[order(org38_SNDMSO_GSEAres_main$padj)][1:10, ]
top49 <- org49_SNDMSO_GSEAres_main[order(org49_SNDMSO_GSEAres_main$padj)][1:10, ]
top66 <- org66_SNDMSO_GSEAres_main[order(org66_SNDMSO_GSEAres_main$padj)][1:10, ]

### get union of pathways
all_pathways <- unique(c(
  top38$pathway,
  top49$pathway,
  top66$pathway
))

NES_matrix <- cbind(
  org38 = get_NES(org38_SNDMSO_GSEAres_main, all_pathways),
  org49 = get_NES(org49_SNDMSO_GSEAres_main, all_pathways),
  org66 = get_NES(org66_SNDMSO_GSEAres_main, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette)
)
```
```{r, fig.height = 10, fig.width = 15}
### select top pathways from collapsed dataframes
top38 <- org38_SNDMSO_GSEAres[order(org38_SNDMSO_GSEAres$padj)][1:10, ]
top49 <- org49_SNDMSO_GSEAres[order(org49_SNDMSO_GSEAres$padj)][1:10, ]
top66 <- org66_SNDMSO_GSEAres[order(org66_SNDMSO_GSEAres$padj)][1:10, ]

### get union of pathways
all_pathways <- unique(c(
  top38$pathway,
  top49$pathway,
  top66$pathway
))

NES_matrix <- cbind(
  org38 = get_NES(org38_SNDMSO_GSEAres, all_pathways),
  org49 = get_NES(org49_SNDMSO_GSEAres, all_pathways),
  org66 = get_NES(org66_SNDMSO_GSEAres, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette),
  border_color = 'NA'
)
```

## {-}

```{r, fig.height = 10, fig.width = 15}
### select top pathways from collapsed dataframes

org38_p2_short <- org38_p2_snsn38_GSEAres_main[order(org38_p2_snsn38_GSEAres_main$padj)][1:20, ]
org38_p3_short <- org38_snSN38_GSEAres_main[order(org38_snSN38_GSEAres_main$padj)][1:20, ]
org38_p2_long <- org38_p2_sndmso_GSEAres_main[order(org38_p2_snsn38_GSEAres_main$padj)][1:20, ]
org38_p3_long <- org38_SNDMSO_GSEAres_main[order(org38_SNDMSO_GSEAres_main$padj)][1:20, ]



### get union of pathways
all_pathways <- unique(c(
  org38_p2_short$pathway,
  org38_p3_short$pathway,
  org38_p2_long$pathway, 
  org38_p3_long$pathway
))

NES_matrix <- cbind(
  `ORG38 short-term response (P2)` = get_NES(org38_p2_snsn38_GSEAres_main, all_pathways),
  `ORG38 short-term response (P3)` = get_NES(org38_snSN38_GSEAres_main, all_pathways),
  `ORG38 long-term response (P2)` = get_NES(org38_p2_sndmso_GSEAres_main, all_pathways),
  `ORG38 long-term response (P3)`= get_NES(org38_SNDMSO_GSEAres_main, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette),
  border_color = 'NA'
)
```



```{r, fig.height = 10, fig.width = 15}
### select top pathways from collapsed dataframes

org49_p2_short <- org49_p2_snsn38_GSEAres_main[order(org49_p2_snsn38_GSEAres_main$padj)][1:10, ]
org49_p3_short <- org49_snSN38_GSEAres_main[order(org49_snSN38_GSEAres_main$padj)][1:10, ]
org49_p2_long <- org49_p2_sndmso_GSEAres_main[order(org49_p2_snsn38_GSEAres_main$padj)][1:10, ]
org49_p3_long <- org49_SNDMSO_GSEAres_main[order(org49_SNDMSO_GSEAres_main$padj)][1:10, ]



### get union of pathways
all_pathways <- unique(c(
  org49_p2_short$pathway,
  org49_p3_short$pathway,
  org49_p2_long$pathway, 
  org49_p3_long$pathway
))

NES_matrix <- cbind(
  `ORG49 short-term response (P2)` = get_NES(org49_p2_snsn38_GSEAres_main, all_pathways),
  `ORG49 short-term response (P3)` = get_NES(org49_snSN38_GSEAres_main, all_pathways),
  `ORG49 long-term response (P2)` = get_NES(org49_p2_sndmso_GSEAres_main, all_pathways),
  `ORG49 long-term response (P3)`= get_NES(org49_SNDMSO_GSEAres_main, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette),
  border_color = 'NA'
)
```
```{r, fig.height = 10, fig.width = 15}
### select top pathways from collapsed dataframes

org49_p2_short <- org49_p2_snsn38_GSEAres[order(org49_p2_snsn38_GSEAres$padj)][1:10, ]
org49_p3_short <- org49_snSN38_GSEAres[order(org49_snSN38_GSEAres$padj)][1:10, ]
org49_p2_long <- org49_p2_sndmso_GSEAres[order(org49_p2_sndmso_GSEAres$padj)][1:10, ]
org49_p3_long <- org49_SNDMSO_GSEAres[order(org49_SNDMSO_GSEAres$padj)][1:10, ]



### get union of pathways
all_pathways <- unique(c(
  org49_p2_short$pathway,
  org49_p3_short$pathway,
  org49_p2_long$pathway, 
  org49_p3_long$pathway
))

NES_matrix <- cbind(
  `ORG49 short-term response (P2)` = get_NES(org49_p2_snsn38_GSEAres, all_pathways),
  `ORG49 short-term response (P3)` = get_NES(org49_snSN38_GSEAres, all_pathways),
  `ORG49 long-term response (P2)` = get_NES(org49_p2_sndmso_GSEAres, all_pathways),
  `ORG49 long-term response (P3)`= get_NES(org49_SNDMSO_GSEAres, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette),
  border_color = 'NA'
)
```

```{r, fig.height = 10, fig.width = 10}
### select top pathways from collapsed dataframes

org66_p2_short <- org66_p2_snsn38_GSEAres_main[order(org66_p2_snsn38_GSEAres_main$padj)][1:5, ]
org66_p3_short <- org66_snSN38_GSEAres_main[order(org66_snSN38_GSEAres_main$padj)][1:5, ]
org66_p2_long <- org66_p2_sndmso_GSEAres_main[order(org66_p2_snsn38_GSEAres_main$padj)][1:5, ]
org66_p3_long <- org66_SNDMSO_GSEAres_main[order(org66_SNDMSO_GSEAres_main$padj)][1:5, ]



### get union of pathways
all_pathways <- unique(c(
  org66_p2_short$pathway,
  org66_p3_short$pathway,
  org66_p2_long$pathway, 
  org66_p3_long$pathway
))

NES_matrix <- cbind(
  `ORG66 short-term response (P2)` = get_NES(org66_p2_snsn38_GSEAres_main, all_pathways),
  `ORG66 short-term response (P3)` = get_NES(org66_snSN38_GSEAres_main, all_pathways),
  `ORG66 long-term response (P2)` = get_NES(org66_p2_sndmso_GSEAres_main, all_pathways),
  `ORG66 long-term response (P3)`= get_NES(org66_SNDMSO_GSEAres_main, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette),
  border_color = 'NA'
)
```

```{r, fig.height = 6, fig.width = 10}
### select top pathways from collapsed dataframes

org66_p2_short <- org66_p2_snsn38_GSEAres[order(org66_p2_snsn38_GSEAres$padj)][1:5, ]
org66_p3_short <- org66_snSN38_GSEAres[order(org66_snSN38_GSEAres$padj)][1:5, ]
org66_p2_long <- org66_p2_sndmso_GSEAres[order(org66_p2_sndmso_GSEAres$padj)][1:5, ]
org66_p3_long <- org66_SNDMSO_GSEAres[order(org66_SNDMSO_GSEAres$padj)][1:5, ]



### get union of pathways
all_pathways <- unique(c(
  org66_p2_short$pathway,
  org66_p3_short$pathway,
  org66_p2_long$pathway, 
  org66_p3_long$pathway
))

NES_matrix <- cbind(
  `ORG66 short-term response (P2)` = get_NES(org66_p2_snsn38_GSEAres, all_pathways),
  `ORG66 short-term response (P3)` = get_NES(org66_snSN38_GSEAres, all_pathways),
  `ORG66 long-term response (P2)` = get_NES(org66_p2_sndmso_GSEAres, all_pathways),
  `ORG66 long-term response (P3)`= get_NES(org66_SNDMSO_GSEAres, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette),
  border_color = 'NA'
)
```
# DSigDB

```{r}
dsigdb_sets <- gmtPathways('../data/DSigDB_All.gmt')
```

#### SN38_SN38 - SN38_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P3/org38_snSN38.csv')
df <- df %>%
  rename(gene_symbol = X)
df <- df[df$logFC < 0, ]
head(df)


org38_p3_snSN38_rankings <- sign(df$logFC)*(-log10(df$P.Value)) # we will use the signed p values from spatial DGE as ranking
names(org38_p3_snSN38_rankings) <- df$gene_symbol # genes as names#

### plot rankings
org38_p3_snSN38_rankings <- sort(org38_p3_snSN38_rankings, decreasing = TRUE) # sort genes by ranking
plot(org38_p3_snSN38_rankings)

ggplot(data.frame(gene_symbol = names(org38_p3_snSN38_rankings[1:50]), ranks = org38_p3_snSN38_rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
# Easy peasy! Run fgsea with the pathways 
test_res <- suppressMessages(
  fgsea(pathways = dsigdb_sets,
                 stats = org38_p3_snSN38_rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(test_res)
```



#### SN38_DMSO - NC_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P3/ORG38_SNDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
df <- df[df$logFC > 0, ]
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org38_p3_sndmso_rankings <- sign(df$logFC)*(-log10(df$P.Value)) # we will use the signed p values from spatial DGE as ranking
names(org38_p3_sndmso_rankings) <- df$gene_symbol # genes as names#

### plot rankings
org38_p3_sndmso_rankings <- sort(org38_p3_sndmso_rankings, decreasing = TRUE) # sort genes by ranking
plot(org38_p3_sndmso_rankings)

ggplot(data.frame(gene_symbol = names(org38_p3_sndmso_rankings[1:50]), ranks = org38_p3_sndmso_rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
test_res_2 <- suppressMessages(
  fgsea(pathways = dsigdb_sets,
                 stats = org38_p3_sndmso_rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(test_res_2)
```

## Poster plots {.tabset}

### NES Heatmaps

```{r, fig.height = 10, fig.width = 15}
### select top pathways from collapsed dataframes

org38_p2_short <- org38_p2_snsn38_GSEAres[order(org38_p2_snsn38_GSEAres$padj)][padj<0.05, ]
org38_p3_short <- org38_snSN38_GSEAres[order(org38_snSN38_GSEAres$padj)][padj<0.05, ]
org38_p2_long <- org38_p2_sndmso_GSEAres[order(org38_p2_sndmso_GSEAres$padj)][padj<0.05, ]
org38_p3_long <- org38_SNDMSO_GSEAres[order(org38_SNDMSO_GSEAres$padj)][padj<0.05, ]



### get union of pathways
all_pathways <- unique(c(
  org38_p2_short$pathway,
  org38_p3_short$pathway,
  org38_p2_long$pathway, 
  org38_p3_long$pathway
))


main_pathways <- all_pathways[1:25]


NES_matrix <- cbind(
  `ORG38 short-term response (P2)` = get_NES(org38_p2_snsn38_GSEAres, main_pathways),
  `ORG38 short-term response (P3)` = get_NES(org38_snSN38_GSEAres, main_pathways),
  `ORG38 long-term response (P2)` = get_NES(org38_p2_sndmso_GSEAres, main_pathways),
  `ORG38 long-term response (P3)`= get_NES(org38_SNDMSO_GSEAres, main_pathways))

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

rownames(NES_matrix)[20] <- 'SIGNALING BY PIEZO1 AND INTEGRINS IN ENDOTHELIAL CELLS'
rownames(NES_matrix)[23] <- 'DISEASES OF SIGNAL TRANSDUCTION BY GROWTH FACTOR RECEPTORS'
### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette),
  border_color = 'NA', 
  cellwidth = 100,
  cellheight = 20
)
```

```{r, fig.height = 10, fig.width = 15}
### select top pathways from collapsed dataframes

org49_p2_short <- org49_p2_snsn38_GSEAres[order(org49_p2_snsn38_GSEAres$padj)][padj<0.05, ]
org49_p3_short <- org49_snSN38_GSEAres[order(org49_snSN38_GSEAres$padj)][padj<0.05, ]
org49_p2_long <- org49_p2_sndmso_GSEAres[order(org49_p2_sndmso_GSEAres$padj)][padj<0.05, ]
org49_p3_long <- org49_SNDMSO_GSEAres[order(org49_SNDMSO_GSEAres$padj)][padj<0.05, ]



### get union of pathways
all_pathways <- unique(c(
  org49_p2_short$pathway,
  org49_p3_short$pathway,
  org49_p2_long$pathway, 
  org49_p3_long$pathway
))

all_pathways <- all_pathways[1:25]

NES_matrix <- cbind(
  `ORG49 short-term response (P2)` = get_NES(org49_p2_snsn38_GSEAres, all_pathways),
  `ORG49 short-term response (P3)` = get_NES(org49_snSN38_GSEAres, all_pathways),
  `ORG49 long-term response (P2)` = get_NES(org49_p2_sndmso_GSEAres, all_pathways),
  `ORG49 long-term response (P3)`= get_NES(org49_SNDMSO_GSEAres, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette),
  border_color = 'NA', 
  cellwidth = 100,
  cellheight = 20
)
```

```{r, fig.height = 10, fig.width = 15}
### select top pathways from collapsed dataframes

org66_p2_short <- org66_p2_snsn38_GSEAres[order(org66_p2_snsn38_GSEAres$padj)][padj<0.05,]
org66_p3_short <- org66_snSN38_GSEAres[order(org66_snSN38_GSEAres$padj)][padj<0.05,]
org66_p2_long <- org66_p2_sndmso_GSEAres[order(org66_p2_sndmso_GSEAres$padj)][padj<0.05,]
org66_p3_long <- org66_SNDMSO_GSEAres[order(org66_SNDMSO_GSEAres$padj)][padj<0.05,]



### get union of pathways
all_pathways <- unique(c(
  org66_p2_short$pathway,
  org66_p3_short$pathway,
  org66_p2_long$pathway, 
  org66_p3_long$pathway
))

all_pathways <- all_pathways[1:25]

NES_matrix <- cbind(
  `ORG66 short-term response (P2)` = get_NES(org66_p2_snsn38_GSEAres, all_pathways),
  `ORG66 short-term response (P3)` = get_NES(org66_snSN38_GSEAres, all_pathways),
  `ORG66 long-term response (P2)` = get_NES(org66_p2_sndmso_GSEAres, all_pathways),
  `ORG66 long-term response (P3)`= get_NES(org66_SNDMSO_GSEAres, all_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1.5
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette),
  border_color = 'NA', 
  cellwidth = 100,
  cellheight = 20
)
```

### limma::camera

```{r}
### loading in genesets

### Sherman-Baust, 2011
sherman_baust <- read_excel('../data/Literature_genesets/SHERMAN_BAUST_2011.XLSX', sheet='geneset_testing')
sherman_baust_up <- as.data.frame(sherman_baust[sherman_baust$OV90P > 1,c('Symbol')])
sherman_baust_dn <- as.data.frame(sherman_baust[sherman_baust$OV90P < -1,c('Symbol')])

### McCorkle, 2021
mccorkle <- read_excel('../data/Literature_genesets/MCCORKLE_2021.xlsx', sheet='geneset_testing')
mccorkle_up <- as.data.frame(mccorkle[mccorkle$log2FoldChange > 1, c('hgnc_symbol')])
mccorkle_dn <- as.data.frame(mccorkle[mccorkle$log2FoldChange < -1, c('hgnc_symbol')])

### Yang, 2024
yang <- read_excel('../data/Literature_genesets/YANGetal2024.xlsx', sheet='geneset_testing')
yang_up <- as.data.frame(yang[yang$A.128PTX_logFC > 1, c('Gene_name')])
yang_dn <- as.data.frame(yang[yang$A.128PTX_logFC < -1, c('Gene_name')])

### creating geneset list 
paclitaxel_resistance_genesets <- list(sherman_baust_up = sherman_baust_up$Symbol,
                                       sherman_baust_dn = sherman_baust_dn$Symbol,
                                       mccorkle_up = mccorkle_up$hgnc_symbol,
                                       mccorkle_dn = mccorkle_dn$hgnc_symbol,
                                       yang_up = yang_up$Gene_name,
                                       yang_dn = yang_dn$Gene_name)

```

Plate 1:

```{r, fig.height=5, fig.width=5}
### loading "fit"
org38_pac_fit <- readRDS('../output/P1/org38_pac_fit.rds') 
org38_sn38_fit <- readRDS('../output/P1/org38_sn38_fit.rds') 
org49_pac_fit <- readRDS('../output/P1/org49_pac_fit.rds') 
org49_sn38_fit <- readRDS('../output/P1/org49_sn38_fit.rds') 
org66_pac_fit <- readRDS('../output/P1/org66_pac_fit.rds')
org66_sn38_fit <- readRDS('../output/P1/org66_sn38_fit.rds') 

### competitive gene set testing (pre-ranked)
df1 <- limma::cameraPR(statistic = org38_pac_fit$t[,1],
       index = paclitaxel_resistance_genesets,
       use.ranks = TRUE)
df1$Organoid <- 'ORG38'
df1$Geneset <- rownames(df1)

df2 <- limma::cameraPR(statistic = org49_pac_fit$t[,1],
       index = paclitaxel_resistance_genesets,
       use.ranks = TRUE)
df2$Organoid <- 'ORG49'
df2$Geneset <- rownames(df2)

df3 <- limma::cameraPR(statistic = org66_pac_fit$t[,1],
       index = paclitaxel_resistance_genesets,
       use.ranks = TRUE)
df3$Organoid <- 'ORG66'
df3$Geneset <- rownames(df3)

p1df <- rbind(df1, df2)
p1df <- rbind(p1df, df3)
p1df$Timepoint <- 'T1'

ggplot(p1df, aes(x= Geneset, y= FDR, fill= Organoid)) + 
  geom_bar(stat = 'identity', position = 'dodge') + 
  coord_flip() +
  geom_hline(yintercept = 0.05, linetype = 'dashed') +
  theme_minimal()
```

Plate 2:
```{r, fig.height=5, fig.width=5}
### loading "fit"
org38_pac_fit <- readRDS('../output/P2/org38_paPAC_fit.rds') 
org38_sn38_fit <- readRDS('../output/P2/org38_snSN38_fit.rds') 
org49_pac_fit <- readRDS('../output/P2/org49_paPAC_fit.rds') 
org49_sn38_fit <- readRDS('../output/P2/org49_snSN38_fit.rds') 
org66_pac_fit <- readRDS('../output/P2/org66_paPAC_fit.rds')
org66_sn38_fit <- readRDS('../output/P2/org66_snSN38_fit.rds') 

### competitive gene set testing (pre-ranked)
df1 <- limma::cameraPR(statistic = org38_pac_fit$t[,1],
       index = paclitaxel_resistance_genesets,
       use.ranks = TRUE)
df1$Organoid <- 'ORG38'
df1$Geneset <- rownames(df1)

df2 <- limma::cameraPR(statistic = org49_pac_fit$t[,1],
       index = paclitaxel_resistance_genesets,
       use.ranks = TRUE)
df2$Organoid <- 'ORG49'
df2$Geneset <- rownames(df2)

df3 <- limma::cameraPR(statistic = org66_pac_fit$t[,1],
       index = paclitaxel_resistance_genesets,
       use.ranks = TRUE)
df3$Organoid <- 'ORG66'
df3$Geneset <- rownames(df3)

p2df <- rbind(df1, df2)
p2df <- rbind(p2df, df3)
p2df$Timepoint <- 'T2'

ggplot(p2df, aes(x= Geneset, y= FDR, fill= Organoid)) + 
  geom_bar(stat = 'identity', position = 'dodge') + 
  coord_flip() +
  geom_hline(yintercept = 0.05, linetype = 'dashed') +
  theme_minimal()
```

Plate 3:
```{r, fig.height=5, fig.width=5}
### loading "fit"
org38_pac_fit <- readRDS('../output/P3/org38_PPACvsDMSO_fit.rds') 
org38_sn38_fit <- readRDS('../output/P3/org38_SNSN38vsDMSO_fit.rds') 
org49_pac_fit <- readRDS('../output/P3/org49_PPACvsDMSO_fit.rds') 
org49_sn38_fit <- readRDS('../output/P3/org49_SNSN38vsDMSO_fit.rds') 
org66_pac_fit <- readRDS('../output/P3/org66_PPACvsDMSO_fit.rds')
org66_sn38_fit <- readRDS('../output/P3/org66_SNSN38vsDMSO_fit.rds') 

### competitive gene set testing (pre-ranked)
limma::cameraPR(statistic = org38_pac_fit$t[,1],
       index = paclitaxel_resistance_genesets,
       use.ranks = TRUE)
limma::cameraPR(statistic = org49_pac_fit$t[,1],
       index = paclitaxel_resistance_genesets,
       use.ranks = TRUE)
limma::cameraPR(statistic = org66_pac_fit$t[,1],
       index = paclitaxel_resistance_genesets,
       use.ranks = TRUE)

### competitive gene set testing (pre-ranked)
df1 <- limma::cameraPR(statistic = org38_pac_fit$t[,1],
       index = paclitaxel_resistance_genesets,
       use.ranks = TRUE)
df1$Organoid <- 'ORG38'
df1$Geneset <- rownames(df1)

df2 <- limma::cameraPR(statistic = org49_pac_fit$t[,1],
       index = paclitaxel_resistance_genesets,
       use.ranks = TRUE)
df2$Organoid <- 'ORG49'
df2$Geneset <- rownames(df2)

df3 <- limma::cameraPR(statistic = org66_pac_fit$t[,1],
       index = paclitaxel_resistance_genesets,
       use.ranks = TRUE)
df3$Organoid <- 'ORG66'
df3$Geneset <- rownames(df3)

p3df <- rbind(df1, df2)
p3df <- rbind(p3df, df3)
p3df$Timepoint <- 'T3'

ggplot(p3df, aes(x= Geneset, y= FDR, fill= Organoid)) + 
  geom_bar(stat = 'identity', position = 'dodge') + 
  coord_flip() +
  geom_hline(yintercept = 0.05, linetype = 'dashed') +
  theme_minimal()
```
