---
title: "Geneset Enrichment Analysis using MSigDB"
output: html_document
date: "2026-03-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.height=4, fig.width = 6)
```

```{r}
library(tidyverse)
library(RColorBrewer)
library(fgsea)
library(msigdbr)
library(DT)
library(ggpubr)
```

```{r}
create_dt <-  function(x){
  datatable(x,
            extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
}
```

## Genesets

```{r}
#gene_sets_df <- msigdbr(species = 'Homo sapiens', category = 'C2')
gene_sets_df <- msigdbr(species = 'Homo sapiens', category = 'C2', subcollection = 'CP:REACTOME')
head(gene_sets_df)

# Convert to named list format required by fgsea
gene_sets <- gene_sets_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(gene_sets), "gene sets\n"))
```

## Plate 2 {.tabset}

### ORG38 {.tabset}


```{r}
# ### 1. Read in data -----------------------------------------------------------
# df <- read.csv('data/DEgenes/P2/ORG38_PACPAC.csv')
# df <- df %>%
#   rename(gene_symbol = X)
# head(df)
# 
# ### 2. Prepare ranked list of genes -----------------------------------------------
# rankings <- sign(df$logFC)*(-log10(df$P.Value)) # we will use the signed p values from spatial DGE as ranking
# names(rankings) <- df$gene_symbol # genes as names#
# 
# ### plot rankings
# rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking
# plot(rankings)
# 
# ggplot(data.frame(gene_symbol = names(rankings[1:50]), ranks = rankings[1:50]), aes(gene_symbol, ranks)) + 
# 	geom_point() +
# 	theme_classic() + 
# 	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
# 
# ### 3. Run GSEA ---------------------------------------------------------------
# 
# org38_p2_paPAC_GSEAres <- fgsea(pathways = gene_sets, # List of gene sets to check
#                  stats = rankings,
#                  scoreType = 'std', 
#                  minSize = 10,
#                  maxSize = 500,
#                  nproc = 1)
```

### ORG49 {.tabset}

### ORG66 {.tabset}



## Plate 3 {.tabset}

### ORG38 {.tabset}

#### PAC_PAC - PAC_DMSO 

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('data/DEgenes/P3/ORG38_paPAC.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
### FGSEA needs a named vector with the rankings and the genes as names
### We will use  signed p value as ranking: sign(df$log2fc)*(-log10(df$pval)))
# Note that you can also just use the log2FC only - but it will not take into account 
# genes with a large FC but non-significant

rankings <- sign(df$logFC)*(-log10(df$P.Value)) # we will use the signed p values from spatial DGE as ranking
names(rankings) <- df$gene_symbol # genes as names#

### plot rankings
rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking
plot(rankings)

# # Some genes have such low p values that the signed pval is +- inf, we need to change it to the maximum * constant to avoid problems with fgsea
# max_ranking <- max(rankings[is.finite(rankings)])
# min_ranking <- min(rankings[is.finite(rankings)])
# rankings <- replace(rankings, rankings > max_ranking, max_ranking * 10)
# rankings <- replace(rankings, rankings < min_ranking, min_ranking * 10)
# rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking

ggplot(data.frame(gene_symbol = names(rankings[1:50]), ranks = rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
# Easy peasy! Run fgsea with the pathways 
org38_paPAC_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets, # List of gene sets to check
                 stats = rankings,
                 scoreType = 'std', # in this case we have both pos and neg rankings. if only pos or neg, set to 'pos', 'neg'
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) # for parallelisation
)
```

GSEA Results:

```{r}
create_dt(org38_paPAC_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org38_paPAC_GSEAres[order(org38_paPAC_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### SN38_SN38 - SN38_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('data/DEgenes/P3/org38_snSN38.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
### FGSEA needs a named vector with the rankings and the genes as names
### We will use  signed p value as ranking: sign(df$log2fc)*(-log10(df$pval)))
# Note that you can also just use the log2FC only - but it will not take into account 
# genes with a large FC but non-significant

rankings <- sign(df$logFC)*(-log10(df$P.Value)) # we will use the signed p values from spatial DGE as ranking
names(rankings) <- df$gene_symbol # genes as names#

### plot rankings
rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking
plot(rankings)

# # Some genes have such low p values that the signed pval is +- inf, we need to change it to the maximum * constant to avoid problems with fgsea
# max_ranking <- max(rankings[is.finite(rankings)])
# min_ranking <- min(rankings[is.finite(rankings)])
# rankings <- replace(rankings, rankings > max_ranking, max_ranking * 10)
# rankings <- replace(rankings, rankings < min_ranking, min_ranking * 10)
# rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking

ggplot(data.frame(gene_symbol = names(rankings[1:50]), ranks = rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
# Easy peasy! Run fgsea with the pathways 
org38_snSN38_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org38_snSN38_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org38_snSN38_GSEAres[order(org38_snSN38_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### PAC_DMSO - NC_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('data/DEgenes/P3/ORG38_PDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
### FGSEA needs a named vector with the rankings and the genes as names
### We will use  signed p value as ranking: sign(df$log2fc)*(-log10(df$pval)))
# Note that you can also just use the log2FC only - but it will not take into account 
# genes with a large FC but non-significant

rankings <- sign(df$logFC)*(-log10(df$P.Value)) # we will use the signed p values from spatial DGE as ranking
names(rankings) <- df$gene_symbol # genes as names#

### plot rankings
rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking
plot(rankings)

# # Some genes have such low p values that the signed pval is +- inf, we need to change it to the maximum * constant to avoid problems with fgsea
# max_ranking <- max(rankings[is.finite(rankings)])
# min_ranking <- min(rankings[is.finite(rankings)])
# rankings <- replace(rankings, rankings > max_ranking, max_ranking * 10)
# rankings <- replace(rankings, rankings < min_ranking, min_ranking * 10)
# rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking

ggplot(data.frame(gene_symbol = names(rankings[1:50]), ranks = rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
# Easy peasy! Run fgsea with the pathways 
org38_PDMSO_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org38_PDMSO_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org38_PDMSO_GSEAres[order(org38_PDMSO_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### SN38_DMSO - NC_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('data/DEgenes/P3/ORG38_SNDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
rankings <- sign(df$logFC)*(-log10(df$P.Value)) # we will use the signed p values from spatial DGE as ranking
names(rankings) <- df$gene_symbol # genes as names#

### plot rankings
rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking
plot(rankings)

ggplot(data.frame(gene_symbol = names(rankings[1:50]), ranks = rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org38_SNDMSO_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org38_SNDMSO_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org38_SNDMSO_GSEAres[order(org38_SNDMSO_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

### ORG49 {.tabset}

#### PAC_PAC - PAC_DMSO 

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('data/DEgenes/P3/ORG49_paPAC.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
rankings <- sign(df$logFC)*(-log10(df$P.Value))
names(rankings) <- df$gene_symbol

### plot rankings
rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking
plot(rankings)

ggplot(data.frame(gene_symbol = names(rankings[1:50]), ranks = rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org49_paPAC_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org49_paPAC_GSEAres)
```

#### SN38_SN38 - SN38_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('data/DEgenes/P3/org49_snSN38.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
rankings <- sign(df$logFC)*(-log10(df$P.Value))
names(rankings) <- df$gene_symbol

### plot rankings
rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking
plot(rankings)

ggplot(data.frame(gene_symbol = names(rankings[1:50]), ranks = rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org49_snSN38_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org49_snSN38_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org49_snSN38_GSEAres[order(org49_snSN38_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### PAC_DMSO - NC_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('data/DEgenes/P3/ORG49_PDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)


### 2. Prepare ranked list of genes -----------------------------------------------
rankings <- sign(df$logFC)*(-log10(df$P.Value))
names(rankings) <- df$gene_symbol

### plot rankings
rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking
plot(rankings)

ggplot(data.frame(gene_symbol = names(rankings[1:50]), ranks = rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org49_PDMSO_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org49_PDMSO_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org49_PDMSO_GSEAres[order(org49_PDMSO_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)

```

#### SN38_DMSO - NC_DMSO
```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('data/DEgenes/P3/ORG49_SNDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
rankings <- sign(df$logFC)*(-log10(df$P.Value))
names(rankings) <- df$gene_symbol

### plot rankings
rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking
plot(rankings)

ggplot(data.frame(gene_symbol = names(rankings[1:50]), ranks = rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org49_SNDMSO_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org49_SNDMSO_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org49_SNDMSO_GSEAres[order(org49_SNDMSO_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

### ORG66 {.tabset}

#### PAC_PAC - PAC_DMSO 

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('data/DEgenes/P3/ORG66_paPAC.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
rankings <- sign(df$logFC)*(-log10(df$P.Value))
names(rankings) <- df$gene_symbol

### plot rankings
rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking
plot(rankings)

ggplot(data.frame(gene_symbol = names(rankings[1:50]), ranks = rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org66_paPAC_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org66_paPAC_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org66_paPAC_GSEAres[order(org66_paPAC_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### SN38_SN38 - SN38_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('data/DEgenes/P3/org66_snSN38.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
rankings <- sign(df$logFC)*(-log10(df$P.Value))
names(rankings) <- df$gene_symbol

### plot rankings
rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking
plot(rankings)

ggplot(data.frame(gene_symbol = names(rankings[1:50]), ranks = rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org66_snSN38_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org66_snSN38_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org66_snSN38_GSEAres[order(org66_snSN38_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]],nrow=3, ncol=3)
```

#### PAC_DMSO - NC_DMSO

```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('data/DEgenes/P3/ORG66_PDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)


### 2. Prepare ranked list of genes -----------------------------------------------
rankings <- sign(df$logFC)*(-log10(df$P.Value))
names(rankings) <- df$gene_symbol

### plot rankings
rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking
plot(rankings)

ggplot(data.frame(gene_symbol = names(rankings[1:50]), ranks = rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org66_PDMSO_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org66_PDMSO_GSEAres)
```

```{r, fig.height = 9, fig.width=12}
top10 <- org66_PDMSO_GSEAres[order(org66_PDMSO_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], rankings) +
    labs(title = pw)
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

#### SN38_DMSO - NC_DMSO
```{r}
### 1. Read in data -----------------------------------------------------------
df <- read.csv('data/DEgenes/P3/ORG66_SNDMSOvsDMSO.csv')
df <- df %>%
  rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
rankings <- sign(df$logFC)*(-log10(df$P.Value))
names(rankings) <- df$gene_symbol

### plot rankings
rankings <- sort(rankings, decreasing = TRUE) # sort genes by ranking
plot(rankings)

ggplot(data.frame(gene_symbol = names(rankings[1:50]), ranks = rankings[1:50]), aes(gene_symbol, ranks)) + 
	geom_point() +
	theme_classic() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### 3. Run GSEA ---------------------------------------------------------------
org66_SNDMSO_GSEAres <- suppressMessages(
  fgsea(pathways = gene_sets,
                 stats = rankings,
                 scoreType = 'std', 
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1) 
)
```

GSEA Results:

```{r}
create_dt(org66_SNDMSO_GSEAres)
```

```{r, fig.height = 9, fig.width=20}
top10 <- org66_SNDMSO_GSEAres[order(org66_SNDMSO_GSEAres$padj), ][1:10, ]

plots <- lapply(top10$pathway, function(pw) {
  plotEnrichment(gene_sets[[pw]], rankings) +
    labs(title = substr(pw, 10,10000000))
})

ggarrange(plots[[1]], plots[[2]], plots[[3]], plots[[4]], plots[[5]], plots[[6]], plots[[7]], plots[[8]], plots[[9]],nrow=3, ncol=3)
```

### All organoids {.tabset}

#### PAC_PAC - PAC_DMSO 
```{r}
list1 <- org38_paPAC_GSEAres[org38_paPAC_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_paPAC_GSEAres[org49_paPAC_GSEAres$padj<0.05,c('pathway')]
list3 <- org66_paPAC_GSEAres[org66_paPAC_GSEAres$padj<0.05,c('pathway')]

common_strings <- Reduce(intersect, list(list1, list3))
common_strings
```

#### SN38_SN38 - SN38_DMSO 
```{r}
list1 <- org38_snSN38_GSEAres[org38_snSN38_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_snSN38_GSEAres[org49_snSN38_GSEAres$padj<0.05,c('pathway')]
list3 <- org66_snSN38_GSEAres[org66_snSN38_GSEAres$padj<0.05,c('pathway')]

common_strings <- Reduce(intersect, list(list1, list2, list3))
common_strings
```

#### PAC_DMSO - NC_DMSO 
```{r}
list1 <- org38_PDMSO_GSEAres[org38_PDMSO_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_PDMSO_GSEAres[org49_PDMSO_GSEAres$padj<0.05,c('pathway')]
list3 <- org66_PDMSO_GSEAres[org66_PDMSO_GSEAres$padj<0.05,c('pathway')]

common_strings <- Reduce(intersect, list(list1, list2, list3))
common_strings
```

#### SN38_DMSO - NC_DMSO 
```{r}
list1 <- org38_SNDMSO_GSEAres[org38_SNDMSO_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_SNDMSO_GSEAres[org49_SNDMSO_GSEAres$padj<0.05,c('pathway')]
list3 <- org66_SNDMSO_GSEAres[org66_SNDMSO_GSEAres$padj<0.05,c('pathway')]

common_strings <- Reduce(intersect, list(list1,list2, list3))
common_strings
```

### Between Treatments {.tabset}

#### ORG38 Short term response 

```{r}
list1 <- org38_paPAC_GSEAres[org38_paPAC_GSEAres$padj<0.05,c('pathway')]
list2 <- org38_snSN38_GSEAres[org38_snSN38_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1,list2)
common_strings
```

#### ORG38 Long term changes 

```{r}
list1 <- org38_PDMSO_GSEAres[org38_PDMSO_GSEAres$padj<0.05,c('pathway')]
list2 <- org38_SNDMSO_GSEAres[org38_SNDMSO_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1,list2)
common_strings
```

#### ORG49 Short term response 

```{r}
list1 <- org49_paPAC_GSEAres[org49_paPAC_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_snSN38_GSEAres[org49_snSN38_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1,list2)
common_strings
```


#### ORG49 Long term changes 

```{r}
list1 <- org49_PDMSO_GSEAres[org49_PDMSO_GSEAres$padj<0.05,c('pathway')]
list2 <- org49_SNDMSO_GSEAres[org49_SNDMSO_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1,list2)
common_strings
```

#### ORG66 Short term response 

```{r}
list1 <- org66_paPAC_GSEAres[org66_paPAC_GSEAres$padj<0.05,c('pathway')]
list2 <- org66_snSN38_GSEAres[org66_snSN38_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1,list2)
common_strings
```

#### ORG66 Long term changes 

```{r}
list1 <- org66_PDMSO_GSEAres[org66_PDMSO_GSEAres$padj<0.05,c('pathway')]
list2 <- org66_SNDMSO_GSEAres[org66_SNDMSO_GSEAres$padj<0.05,c('pathway')]

common_strings <- intersect(list1,list2)
common_strings
```


