---
title: "ORG49_SN38_all"
output: html_document
date: "2026-06-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(RColorBrewer)
library(fgsea)
library(msigdbr)
library(DT)
library(ggpubr)
library(dplyr)
library(tidyverse)
library(RColorBrewer)
library(ggvenn)
library(gplots)
library(tidyr)
library(ggplot2)

library(clusterProfiler)
library(DOSE)
library(enrichplot)
```

```{r}
create_dt <-  function(x){
  datatable(x,
            extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
}
```

```{r}
### loading genesets
reactome_df <-msigdbr(species = 'Homo sapiens', category = 'C2', subcollection = 'CP:REACTOME')
reactome_df <- reactome_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(reactome_df), "gene sets\n"))

go_df <-msigdbr(species = 'Homo sapiens', category = 'C5')
go_df <- go_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(go_df), "gene sets\n"))

kegg_df <-msigdbr(species = 'Homo sapiens', category = 'C2', subcollection = 'CP:KEGG_MEDICUS')
kegg_df <- kegg_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(kegg_df), "gene sets\n"))

hallmark_df <- msigdbr(species = 'Homo sapiens', category = 'H')
hallmark_df <- hallmark_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(hallmark_df), "gene sets\n"))
```

## Notes

This analysis contains data for ORG49 SN38 treated samples (compared to DMSO) across all MACseq Plates (P1-P3). Differential expression analyses are derived from respected .Rmd/.html files previously published on github: 

[P1 Differential Expression]()
[P2 Differential Expression]()
[P3 Differential Expression]()

Pathway analysis is a ranked analysis using ALL GENES (even statistically non-significant) where rank = logFC x adj.p.val 

Genes are analysed across different databases (KEGG, Reactome, GO, Hallmark) and results should be compared across databases to identify shared signals. 

## Plate 1 {.tabset}

```{r}
### load DE results 
### setting up ORG49 DE data
df <- read.csv('data/DEgenes/P1/ORG49_ncSN38vsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p1_ncSN38 <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p1_ncSN38) <- df$gene_symbol 
```

### GO

```{r}
p1_ncSN38_go <- fgsea(pathways = go_df, 
                 stats = p1_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p1_ncSN38_go)
```

### KEGG

```{r}
p1_ncSN38_kegg <- fgsea(pathways = kegg_df, 
                 stats = p1_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p1_ncSN38_kegg)
```

### Reactome 

```{r}
p1_ncSN38_reactome <- fgsea(pathways = reactome_df, 
                 stats = p1_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p1_ncSN38_reactome)
```

## Plate 2 {.tabset}

```{r}
### ncSN38
### load DE results 
df <- read.csv('data/DEgenes/P2/ORG49_ncSN38vsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p2_ncSN38 <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p2_ncSN38) <- df$gene_symbol 

### snSN38
### load DE results 
df <- read.csv('data/DEgenes/P2/ORG49_snSN38vsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p2_snSN38 <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p2_snSN38) <- as.factor(df$gene_symbol)

### longSN38
### load DE results 
df <- read.csv('data/DEgenes/P2/ORG49_longSN38.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p2_longSN38 <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p2_longSN38) <- df$gene_symbol 

### snDMSO
### load DE results 
df <- read.csv('data/DEgenes/P2/ORG49_snDMSOvsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p2_snDMSO <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p2_snDMSO) <- df$gene_symbol 
```

### GO {.tabset}

#### ncSN38 - ncDMSO

```{r}
p2_ncSN38_go <- fgsea(pathways = go_df, 
                 stats = p2_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p2_ncSN38_go)
```

#### snSN38-snDMSO

```{r}
p2_snSN38_go <- fgsea(pathways = go_df, 
                 stats = p2_snSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p2_snSN38_go)
```

#### longSN38

```{r}
p2_longSN38_go <- fgsea(pathways = go_df, 
                 stats = p2_longSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p2_longSN38_go)
```

#### snDMSO-ncDMSO

```{r}
p2_snDMSO_go <- fgsea(pathways = go_df, 
                 stats = p2_snDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p2_snDMSO_go)
```

### KEGG {.tabset}

#### ncSN38 - ncDMSO

```{r}
p2_ncSN38_kegg <- fgsea(pathways = kegg_df, 
                 stats = p2_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p2_ncSN38_kegg)
```

#### snSN38-snDMSO

```{r}
p2_snSN38_kegg <- fgsea(pathways = kegg_df, 
                 stats = p2_snSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p2_snSN38_kegg)
```

#### longSN38

```{r}
p2_longSN38_kegg <- fgsea(pathways = kegg_df, 
                 stats = p2_longSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p2_longSN38_kegg)
```

#### snDMSO-ncDMSO

```{r}
p2_snDMSO_kegg <- fgsea(pathways = kegg_df, 
                 stats = p2_snDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p2_snDMSO_kegg)
```

### Reactome {.tabset}

#### ncSN38 - ncDMSO

```{r}
p2_ncSN38_reactome <- fgsea(pathways = reactome_df, 
                 stats = p2_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p2_ncSN38_reactome)
```

#### snSN38-snDMSO

```{r}
p2_snSN38_reactome <- fgsea(pathways = reactome_df, 
                 stats = p2_snSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p2_snSN38_reactome)
```

#### longSN38

```{r}
p2_longSN38_reactome <- fgsea(pathways = reactome_df, 
                 stats = p2_longSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p2_longSN38_reactome)
```

#### snDMSO-ncDMSO

```{r}
p2_snDMSO_reactome <- fgsea(pathways = reactome_df, 
                 stats = p2_snDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p2_snDMSO_reactome)
```

## Plate 3 {.tabset}

```{r}
### ncSN38
### load DE results 
df <- read.csv('data/DEgenes/P3/ORG49_ncSN38vsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p3_ncSN38 <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p3_ncSN38) <- df$gene_symbol 

### snSN38
### load DE results 
df <- read.csv('data/DEgenes/P3/ORG49_snSN38vsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p3_snSN38 <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p3_snSN38) <- df$gene_symbol 

### longSN38
### load DE results 
df <- read.csv('data/DEgenes/P3/ORG49_longSN38.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p3_longSN38 <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p3_longSN38) <- df$gene_symbol 

### snDMSO
### load DE results 
df <- read.csv('data/DEgenes/P3/ORG49_snDMSOvsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p3_snDMSO <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p3_snDMSO) <- df$gene_symbol 
```

### GO {.tabset}

#### ncSN38 - ncDMSO

```{r}
p3_ncSN38_go <- fgsea(pathways = go_df, 
                 stats = p3_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p3_ncSN38_go)
```

#### snSN38-snDMSO

```{r}
p3_snSN38_go <- fgsea(pathways = go_df, 
                 stats = p3_snSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p3_snSN38_go)
```

#### longSN38

```{r}
p3_longSN38_go <- fgsea(pathways = go_df, 
                 stats = p3_longSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p3_longSN38_go)
```

#### snDMSO-ncDMSO

```{r}
p3_snDMSO_go <- fgsea(pathways = go_df, 
                 stats = p3_snDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p3_snDMSO_go)
```

### KEGG {.tabset}

#### ncSN38 - ncDMSO

```{r}
p3_ncSN38_kegg <- fgsea(pathways = kegg_df, 
                 stats = p3_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p3_ncSN38_kegg)
```

#### snSN38-snDMSO

```{r}
p3_snSN38_kegg <- fgsea(pathways = kegg_df, 
                 stats = p3_snSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p3_snSN38_kegg)
```

#### longSN38

```{r}
p3_longSN38_kegg <- fgsea(pathways = kegg_df, 
                 stats = p3_longSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p3_longSN38_kegg)
```

#### snDMSO-ncDMSO

```{r}
p3_snDMSO_kegg <- fgsea(pathways = kegg_df, 
                 stats = p3_snDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p3_snDMSO_kegg)
```

### Reactome {.tabset}

#### ncSN38 - ncDMSO

```{r}
p3_ncSN38_reactome <- fgsea(pathways = reactome_df, 
                 stats = p3_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p3_ncSN38_reactome)
```

#### snSN38-snDMSO

```{r}
p3_snSN38_reactome <- fgsea(pathways = reactome_df, 
                 stats = p3_snSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p3_snSN38_reactome)
```

#### longSN38

```{r}
p3_longSN38_reactome <- fgsea(pathways = reactome_df, 
                 stats = p3_longSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p3_longSN38_reactome)
```

#### snDMSO-ncDMSO

```{r}
p3_snDMSO_reactome <- fgsea(pathways = reactome_df, 
                 stats = p3_snDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(p3_snDMSO_reactome)
```
