---
title: "ORG66_PAC_all"
output: html_document
date: "2026-06-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(RColorBrewer)
library(fgsea)
library(msigdbr)
library(DT)
library(ggpubr)
library(dplyr)
library(tidyverse)
library(RColorBrewer)
library(ggvenn)
library(gplots)
library(tidyr)
library(ggplot2)

library(clusterProfiler)
library(DOSE)
library(enrichplot)
library(org.Hs.eg.db)
library(gsubfn)

```

```{r}
create_dt <-  function(x){
  datatable(x,
            extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
}
```

```{r}
### loading genesets
reactome_df <-msigdbr(species = 'Homo sapiens', category = 'C2', subcollection = 'CP:REACTOME')
reactome_df <- reactome_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(reactome_df), "gene sets\n"))

go_df <-msigdbr(species = 'Homo sapiens', category = 'C5')
go_df <- go_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(go_df), "gene sets\n"))

kegg_df <-msigdbr(species = 'Homo sapiens', category = 'C2', subcollection = 'CP:KEGG_MEDICUS')
kegg_df <- kegg_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(kegg_df), "gene sets\n"))

hallmark_df <- msigdbr(species = 'Homo sapiens', category = 'H')
hallmark_df <- hallmark_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(hallmark_df), "gene sets\n"))
```



```{r}
pathway_theme_map <- function(pathway_names){

  case_when(
    grepl("IMMUNE|INTERFERON|INFLAMM|CYTOKINE|ANTIGEN|T_CELL|B_CELL|NK|MHC|TLR|INNATE|ADAPTIVE|COMPLEMENT",
          pathway_names, ignore.case = TRUE) ~ "Immune / Inflammation",
    grepl("SIGNAL|MAPK|ERK|JAK|STAT|PI3K|AKT|WNT|NOTCH|HEDGEHOG|TGF|NF.KB|RAS|RHO|GTPase",
          pathway_names, ignore.case = TRUE) ~ "Signal Transduction",
    grepl("CELL_CYCLE|MITOSIS|MEIOSIS|G1|S PHASE|G2|CHECKPOINT|APC.C|CDK|CYCLIN|REPLICATION|CHROMOSOME_SEGREGATION|CHROMOSOME_CONDENSATION|CHROMOSOME_SEPARATION",
          pathway_names, ignore.case = TRUE) ~ "Cell Cycle / Division",
    grepl("APOPTOSIS|PYROPTOSIS|NECROPTOSIS|FERROPTOSIS|AUTOPHAGY|DNA DAMAGE|STRESS|UNFOLDED PROTEIN",
          pathway_names, ignore.case = TRUE) ~ "Cell Death / Stress Response",
    grepl("METABOLISM|GLYCOLYSIS|TCA|OXIDATIVE PHOSPHORYLATION|FATTY ACID|CHOLESTEROL|LIPID|AMINO ACID|NUCLEOTIDE|METABOLIC|CATABOLIC",
          pathway_names, ignore.case = TRUE) ~ "Metabolism",
    grepl("TRANSLATION|RIBOSOME|SPLICING|RNA|TRANSCRIPTION|RNA POLYMERASE|MRNA|RRNA|TRNA|PROCESSING",
          pathway_names, ignore.case = TRUE) ~ "Gene Expression / RNA Processing",
    grepl("DNA REPAIR|REPLICATION|TELOMERE|CHROMATIN|NUCLEOSOME|EPIGEN|HISTONE|METHYLATION",
          pathway_names, ignore.case = TRUE) ~ "Genome Maintenance / Epigenetics",
    grepl("DEVELOP|DIFFERENTIATION|MORPHOGEN|EMBRYO|AXIS|NEUROGEN|CELL FATE",
          pathway_names, ignore.case = TRUE) ~ "Development / Differentiation",
    grepl("ECM|EXTRACELLULAR|INTEGRIN|ADHESION|COLLAGEN|MIGRATION|CYTOSKELETON|ACTIN|MICROTUBULE",
          pathway_names, ignore.case = TRUE) ~ "ECM / Adhesion / Cytoskeleton",
    grepl("VESICLE|ENDOCYTOSIS|EXOCYTOSIS|GOLGI|ER|LYSOSOME|PEROXISOME|TRANSPORT",
          pathway_names, ignore.case = TRUE) ~ "Transport / Organelle Function",
    grepl("CANCER|CARCINOMA|MELANOMA|LEUKEMIA|TUMOR|ONCOGEN|P53|ONCOGENIC",
          pathway_names, ignore.case = TRUE) ~ "Cancer / Disease",
    TRUE ~ "Other / Unclassified"
  )
}
```

## Notes

This analysis contains data for ORG66 Paclitaxel treated samples (compared to DMSO) across all MACseq Plates (P1-P3). Differential expression analyses are derived from respected .Rmd/.html files previously published on github: 

[P1 Differential Expression]()
[P2 Differential Expression]()
[P3 Differential Expression]()

Pathway analysis is a ranked analysis using ALL GENES (even statistically non-significant) where rank = logFC x adj.p.val 

Genes are analysed across different databases (KEGG, Reactome, GO, Hallmark) and results should be compared across databases to identify shared signals. 

## Plate 1 {.tabset}

```{r}
### load DE results 
### setting up ORG66 DE data
df <- read.csv('data/DEgenes/P1/ORG66_ncPACvsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p1_ncPAC <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p1_ncPAC) <- df$gene_symbol 
```

### GO

```{r}
p1_ncpac_go <- fgsea(pathways = go_df, 
                 stats = p1_ncPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p1_ncpac_go)
```

```{r}
sigPathways <- p1_ncpac_go[padj < 0.05]$pathway
sigRes <- p1_ncpac_go[p1_ncpac_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p1_ncPAC,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p1_ncpac_go_simplified <- p1_ncpac_go[p1_ncpac_go$pathway %in% main_pathways, ]

create_dt(p1_ncpac_go_simplified)

### format dataframe
p1_ncpac_go_simplified$ONTOLOGY <- p1_ncpac_go_simplified$pathway
  
p1_ncpac_go_simplified <- p1_ncpac_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p1_ncpac_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p1_ncpac_go_simplified$theme <- pathway_theme_map(p1_ncpac_go_simplified$pathway)
```

```{r, fig.width=10}
ggplot(p1_ncpac_go_simplified,
       aes(x=label, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
  theme_bw()
```


### KEGG

```{r}
p1_ncpac_kegg <- fgsea(pathways = kegg_df, 
                 stats = p1_ncPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p1_ncpac_kegg)
```

```{r}
sigPathways <- p1_ncpac_kegg[padj < 0.05]$pathway
sigRes <- p1_ncpac_kegg[p1_ncpac_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p1_ncPAC,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p1_ncpac_kegg_simplified <- p1_ncpac_kegg[p1_ncpac_kegg$pathway %in% main_pathways, ]

create_dt(p1_ncpac_kegg_simplified)
```

```{r}
p1_ncpac_kegg_simplified$theme <- pathway_theme_map(p1_ncpac_kegg_simplified$pathway)
```

```{r, fig.width=10}
ggplot(p1_ncpac_kegg_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

### Reactome 

```{r}
p1_ncpac_reactome <- fgsea(pathways = reactome_df, 
                 stats = p1_ncPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p1_ncpac_reactome)
```

```{r}
sigPathways <- p1_ncpac_reactome[padj < 0.05]$pathway
sigRes <- p1_ncpac_reactome[p1_ncpac_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p1_ncPAC,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p1_ncpac_reactome_simplified <- p1_ncpac_reactome[p1_ncpac_reactome$pathway %in% main_pathways, ]

create_dt(p1_ncpac_reactome_simplified)
```

```{r, fig.width=15}
p1_ncpac_reactome_simplified$theme <- pathway_theme_map(p1_ncpac_reactome_simplified$pathway)
p1_ncpac_reactome_simplified <- p1_ncpac_reactome_simplified %>%
  arrange(NES)

ggplot(p1_ncpac_reactome_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```


## Plate 2 {.tabset}

```{r}
### ncPAC
### load DE results 
df <- read.csv('data/DEgenes/P2/ORG66_ncPACvsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p2_ncPAC <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p2_ncPAC) <- df$gene_symbol 

### paPAC
### load DE results 
df <- read.csv('data/DEgenes/P2/ORG66_paPACvsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p2_paPAC <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p2_paPAC) <- as.factor(df$gene_symbol)

### longPAC
### load DE results 
df <- read.csv('data/DEgenes/P2/ORG66_longPAC.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p2_longPAC <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p2_longPAC) <- df$gene_symbol 

### paDMSO
### load DE results 
df <- read.csv('data/DEgenes/P2/ORG66_paDMSOvsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p2_paDMSO <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p2_paDMSO) <- df$gene_symbol 
```

### GO {.tabset}

#### ncPAC - ncDMSO

```{r}
p2_ncpac_go <- fgsea(pathways = go_df, 
                 stats = p2_ncPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_ncpac_go)
```

```{r}
sigPathways <- p2_ncpac_go[padj < 0.05]$pathway
sigRes <- p2_ncpac_go[p2_ncpac_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_ncPAC,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_ncpac_go_simplified <- p2_ncpac_go[p2_ncpac_go$pathway %in% main_pathways, ]

create_dt(p2_ncpac_go_simplified)

### format dataframe
p2_ncpac_go_simplified$ONTOLOGY <- p2_ncpac_go_simplified$pathway
  
p2_ncpac_go_simplified <- p2_ncpac_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p2_ncpac_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p2_ncpac_go_simplified$theme <- pathway_theme_map(p2_ncpac_go_simplified$pathway)
```

```{r, fig.width=10}
ggplot(p2_ncpac_go_simplified,
       aes(x=label, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
  theme_bw()
```

#### paPAC-paDMSO

```{r}
p2_papac_go <- fgsea(pathways = go_df, 
                 stats = p2_paPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_papac_go)
```

```{r}
sigPathways <- p2_papac_go[padj < 0.05]$pathway
sigRes <- p2_papac_go[p2_papac_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_paPAC,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_papac_go_simplified <- p2_papac_go[p2_papac_go$pathway %in% main_pathways, ]

create_dt(p2_papac_go_simplified)

### format dataframe
p2_papac_go_simplified$ONTOLOGY <- p2_papac_go_simplified$pathway
  
p2_papac_go_simplified <- p2_papac_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p2_papac_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p2_papac_go_simplified$theme <- pathway_theme_map(p2_papac_go_simplified$pathway)
```
*No pathways pass multiple testing for simplified analysis*

```{r, fig.width=10}
# ggplot(p2_papac_go_simplified,
#        aes(x=label, y= NES, fill=theme)) +
#   geom_bar(stat='identity') +
#   coord_flip()+
#   facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
#   theme_bw()
```

#### longPAC

```{r}
p2_longPAC_go <- fgsea(pathways = go_df, 
                 stats = p2_longPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_longPAC_go)
```

```{r}
sigPathways <- p2_longPAC_go[padj < 0.05]$pathway
sigRes <- p2_longPAC_go[p2_longPAC_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_longPAC,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_longpac_go_simplified <- p2_longPAC_go[p2_longPAC_go$pathway %in% main_pathways, ]

create_dt(p2_longpac_go_simplified)

### format dataframe
p2_longpac_go_simplified$ONTOLOGY <- p2_longpac_go_simplified$pathway
  
p2_longpac_go_simplified <- p2_longpac_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p2_longpac_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p2_longpac_go_simplified$theme <- pathway_theme_map(p2_longpac_go_simplified$pathway)
```

*No pathways pass multiple testing for simplified analysis*

```{r, fig.width=10}
# ggplot(p2_longpac_go_simplified,
#        aes(x=label, y= NES, fill=theme)) +
#   geom_bar(stat='identity') +
#   coord_flip()+
#   facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
#   theme_bw()
```

#### paDMSO-ncDMSO

```{r}
p2_paDMSO_go <- fgsea(pathways = go_df, 
                 stats = p2_paDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_paDMSO_go)
```

```{r}
sigPathways <- p2_paDMSO_go[padj < 0.05]$pathway
sigRes <- p2_paDMSO_go[p2_paDMSO_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_paDMSO,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_paDMSO_go_simplified <- p2_paDMSO_go[p2_paDMSO_go$pathway %in% main_pathways, ]

create_dt(p2_paDMSO_go_simplified)

### format dataframe
p2_paDMSO_go_simplified$ONTOLOGY <- p2_paDMSO_go_simplified$pathway
  
p2_paDMSO_go_simplified <- p2_paDMSO_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p2_paDMSO_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p2_paDMSO_go_simplified$theme <- pathway_theme_map(p2_paDMSO_go_simplified$pathway)
```
*No pathways pass multiple testing for simplified analysis*

```{r, fig.width=10}
# ggplot(p2_paDMSO_go_simplified,
#        aes(x=label, y= NES, fill=theme)) +
#   geom_bar(stat='identity') +
#   coord_flip()+
#   facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
#   theme_bw()
```

### KEGG {.tabset}

#### ncPAC - ncDMSO

```{r}
p2_ncpac_kegg <- fgsea(pathways = kegg_df, 
                 stats = p2_ncPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_ncpac_kegg)
```

```{r}
sigPathways <- p2_ncpac_kegg[padj < 0.05]$pathway
sigRes <- p2_ncpac_kegg[p2_ncpac_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_ncPAC,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_ncpac_kegg_simplified <- p2_ncpac_kegg[p2_ncpac_kegg$pathway %in% main_pathways, ]

create_dt(p2_ncpac_kegg_simplified)

p2_ncpac_kegg_simplified$theme <- pathway_theme_map(p2_ncpac_kegg_simplified$pathway)
```

```{r, fig.width=10}
ggplot(p2_ncpac_kegg_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### paPAC-paDMSO

```{r}
p2_papac_kegg <- fgsea(pathways = kegg_df, 
                 stats = p2_paPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_papac_kegg)
```

```{r}
sigPathways <- p2_papac_kegg[padj < 0.05]$pathway
sigRes <- p2_papac_kegg[p2_papac_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_paPAC,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_papac_kegg_simplified <- p2_papac_kegg[p2_papac_kegg$pathway %in% main_pathways, ]

create_dt(p2_papac_kegg_simplified)

p2_papac_kegg_simplified$theme <- pathway_theme_map(p2_papac_kegg_simplified$pathway)
```
*No pathways pass multiple testing for simplified analysis*

```{r, fig.width=10, fig.height=3}
# ggplot(p2_papac_kegg_simplified,
#        aes(x=pathway, y= NES, fill=theme)) +
#   geom_bar(stat='identity') +
#   coord_flip()+
#   theme_bw()
```

#### longPAC

```{r}
p2_longPAC_kegg <- fgsea(pathways = kegg_df, 
                 stats = p2_longPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_longPAC_kegg)
```

```{r}
sigPathways <- p2_longPAC_kegg[padj < 0.05]$pathway
sigRes <- p2_longPAC_kegg[p2_longPAC_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_longPAC,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_longPAC_kegg_simplified <- p2_longPAC_kegg[p2_longPAC_kegg$pathway %in% main_pathways, ]

create_dt(p2_longPAC_kegg_simplified)

p2_longPAC_kegg_simplified$theme <- pathway_theme_map(p2_longPAC_kegg_simplified$pathway)
```

*No pathways pass multiple testing for simplified analysis*

```{r, fig.width=10, fig.height=3}
# ggplot(p2_longPAC_kegg_simplified,
#        aes(x=pathway, y= NES, fill=theme)) +
#   geom_bar(stat='identity') +
#   coord_flip()+
#   theme_bw()
```

#### paDMSO-ncDMSO

```{r}
p2_paDMSO_kegg <- fgsea(pathways = kegg_df, 
                 stats = p2_paDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_paDMSO_kegg)
```

```{r}
sigPathways <- p2_paDMSO_kegg[padj < 0.05]$pathway
sigRes <- p2_paDMSO_kegg[p2_paDMSO_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_paDMSO,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_paDMSO_kegg_simplified <- p2_paDMSO_kegg[p2_paDMSO_kegg$pathway %in% main_pathways, ]

create_dt(p2_paDMSO_kegg_simplified)

p2_paDMSO_kegg_simplified$theme <- pathway_theme_map(p2_paDMSO_kegg_simplified$pathway)
```

*No pathways pass multiple testing for simplified analysis*

```{r, fig.width=20}
# ggplot(p2_paDMSO_kegg_simplified,
#        aes(x=pathway, y= NES, fill=theme)) +
#   geom_bar(stat='identity') +
#   coord_flip()+
#   theme_bw()
```

### Reactome {.tabset}

#### ncPAC - ncDMSO

```{r}
p2_ncpac_reactome <- fgsea(pathways = reactome_df, 
                 stats = p2_ncPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_ncpac_reactome)
```

```{r}
sigPathways <- p2_ncpac_reactome[padj < 0.05]$pathway
sigRes <- p2_ncpac_reactome[p2_ncpac_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_ncPAC,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_ncpac_reactome_simplified <- p2_ncpac_reactome[p2_ncpac_reactome$pathway %in% main_pathways, ]

create_dt(p2_ncpac_reactome_simplified)
```

```{r, fig.width=15}
p2_ncpac_reactome_simplified$theme <- pathway_theme_map(p2_ncpac_reactome_simplified$pathway)

ggplot(p2_ncpac_reactome_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### paPAC-paDMSO

```{r}
p2_papac_reactome <- fgsea(pathways = reactome_df, 
                 stats = p2_paPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_papac_reactome)
```

```{r}
sigPathways <- p2_papac_reactome[padj < 0.05]$pathway
sigRes <- p2_papac_reactome[p2_papac_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_paPAC,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_papac_reactome_simplified <- p2_papac_reactome[p2_papac_reactome$pathway %in% main_pathways, ]

create_dt(p2_papac_reactome_simplified)
```
*No pathways pass multiple testing for simplified analysis*

```{r, fig.width=15}
# p2_papac_reactome_simplified$theme <- pathway_theme_map(p2_papac_reactome_simplified$pathway)
# 
# ggplot(p2_papac_reactome_simplified,
#        aes(x=pathway, y= NES, fill=theme)) +
#   geom_bar(stat='identity') +
#   coord_flip()+
#   theme_bw()
```

#### longPAC

```{r}
p2_longPAC_reactome <- fgsea(pathways = reactome_df, 
                 stats = p2_longPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_longPAC_reactome)
```

```{r}
sigPathways <- p2_longPAC_reactome[padj < 0.05]$pathway
sigRes <- p2_longPAC_reactome[p2_longPAC_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_longPAC,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_longPAC_reactome_simplified <- p2_longPAC_reactome[p2_longPAC_reactome$pathway %in% main_pathways, ]

create_dt(p2_longPAC_reactome_simplified)
```

```{r, fig.width=15}
p2_longPAC_reactome_simplified$theme <- pathway_theme_map(p2_longPAC_reactome_simplified$pathway)

ggplot(p2_longPAC_reactome_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### paDMSO-ncDMSO

```{r}
p2_paDMSO_reactome <- fgsea(pathways = reactome_df, 
                 stats = p2_paDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_paDMSO_reactome)
```

```{r}
sigPathways <- p2_paDMSO_reactome[padj < 0.05]$pathway
sigRes <- p2_paDMSO_reactome[p2_paDMSO_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_paDMSO,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_paDMSO_reactome_simplified <- p2_paDMSO_reactome[p2_paDMSO_reactome$pathway %in% main_pathways, ]

create_dt(p2_paDMSO_reactome_simplified)
```
*No pathways pass multiple testing for simplified analysis*

```{r, fig.width=15}
# p2_paDMSO_reactome_simplified$theme <- pathway_theme_map(p2_paDMSO_reactome_simplified$pathway)
# 
# ggplot(p2_paDMSO_reactome_simplified,
#        aes(x=pathway, y= NES, fill=theme)) +
#   geom_bar(stat='identity') +
#   coord_flip()+
#   theme_bw()
```

## Plate 3 {.tabset}

```{r}
### ncPAC
### load DE results 
df <- read.csv('data/DEgenes/P3/ORG66_ncPACvsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p3_ncPAC <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p3_ncPAC) <- df$gene_symbol 

### paPAC
### load DE results 
df <- read.csv('data/DEgenes/P3/ORG66_paPACvsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p3_paPAC <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p3_paPAC) <- df$gene_symbol 

### longPAC
### load DE results 
df <- read.csv('data/DEgenes/P3/ORG66_longPAC.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p3_longPAC <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p3_longPAC) <- df$gene_symbol 

### paDMSO
### load DE results 
df <- read.csv('data/DEgenes/P3/ORG66_paDMSOvsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p3_paDMSO <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p3_paDMSO) <- df$gene_symbol 
```

### GO {.tabset}

#### ncPAC - ncDMSO

```{r}
p3_ncpac_go <- fgsea(pathways = go_df, 
                 stats = p3_ncPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_ncpac_go)
```

```{r}
sigPathways <- p3_ncpac_go[padj < 0.05]$pathway
sigRes <- p3_ncpac_go[p3_ncpac_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_ncPAC,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_ncpac_go_simplified <- p3_ncpac_go[p3_ncpac_go$pathway %in% main_pathways, ]

create_dt(p3_ncpac_go_simplified)

### format dataframe
p3_ncpac_go_simplified$ONTOLOGY <- p3_ncpac_go_simplified$pathway
  
p3_ncpac_go_simplified <- p3_ncpac_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p3_ncpac_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p3_ncpac_go_simplified$theme <- pathway_theme_map(p3_ncpac_go_simplified$pathway)
```

```{r, fig.width=10}
ggplot(p3_ncpac_go_simplified,
       aes(x=label, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
  theme_bw()
```

#### paPAC-paDMSO

```{r}
p3_papac_go <- fgsea(pathways = go_df, 
                 stats = p3_paPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_papac_go)
```

```{r}
sigPathways <- p3_papac_go[padj < 0.05]$pathway
sigRes <- p3_papac_go[p3_papac_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_paPAC,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_papac_go_simplified <- p3_papac_go[p3_papac_go$pathway %in% main_pathways, ]

create_dt(p3_papac_go_simplified)

### format dataframe
p3_papac_go_simplified$ONTOLOGY <- p3_papac_go_simplified$pathway
  
p3_papac_go_simplified <- p3_papac_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p3_papac_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p3_papac_go_simplified$theme <- pathway_theme_map(p3_papac_go_simplified$pathway)
```

```{r, fig.width=10}
ggplot(p3_papac_go_simplified,
       aes(x=label, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
  theme_bw()
```

#### longPAC

```{r}
p3_longPAC_go <- fgsea(pathways = go_df, 
                 stats = p3_longPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_longPAC_go)
```

```{r}
sigPathways <- p3_longPAC_go[padj < 0.05]$pathway
sigRes <- p3_longPAC_go[p3_longPAC_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_longPAC,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_longPAC_go_simplified <- p3_longPAC_go[p3_longPAC_go$pathway %in% main_pathways, ]

create_dt(p3_longPAC_go_simplified)

### format dataframe
p3_longPAC_go_simplified$ONTOLOGY <- p3_longPAC_go_simplified$pathway
  
p3_longPAC_go_simplified <- p3_longPAC_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p3_longPAC_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p3_longPAC_go_simplified$theme <- pathway_theme_map(p3_longPAC_go_simplified$pathway)
```

```{r, fig.width=10}
ggplot(p3_longPAC_go_simplified,
       aes(x=label, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
  theme_bw()
```

#### paDMSO-ncDMSO

```{r}
p3_paDMSO_go <- fgsea(pathways = go_df, 
                 stats = p3_paDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_paDMSO_go)
```

```{r}
sigPathways <- p3_paDMSO_go[padj < 0.05]$pathway
sigRes <- p3_paDMSO_go[p3_paDMSO_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_paDMSO,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_paDMSO_go_simplified <- p3_paDMSO_go[p3_paDMSO_go$pathway %in% main_pathways, ]

create_dt(p3_paDMSO_go_simplified)

### format dataframe
p3_paDMSO_go_simplified$ONTOLOGY <- p3_paDMSO_go_simplified$pathway
  
p3_paDMSO_go_simplified <- p3_paDMSO_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p3_paDMSO_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p3_paDMSO_go_simplified$theme <- pathway_theme_map(p3_paDMSO_go_simplified$pathway)
```

```{r, fig.width=10}
ggplot(p3_paDMSO_go_simplified,
       aes(x=label, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
  theme_bw()
```

### KEGG {.tabset}

#### ncPAC - ncDMSO

```{r}
p3_ncpac_kegg <- fgsea(pathways = kegg_df, 
                 stats = p3_ncPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_ncpac_kegg)
```

```{r}
sigPathways <- p3_ncpac_kegg[padj < 0.05]$pathway
sigRes <- p3_ncpac_kegg[p3_ncpac_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_ncPAC,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_ncpac_kegg_simplified <- p3_ncpac_kegg[p3_ncpac_kegg$pathway %in% main_pathways, ]

create_dt(p3_ncpac_kegg_simplified)

p3_ncpac_kegg_simplified$theme <- pathway_theme_map(p3_ncpac_kegg_simplified$pathway)
```

```{r, fig.width=20}
ggplot(p3_ncpac_kegg_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### paPAC-paDMSO

```{r}
p3_papac_kegg <- fgsea(pathways = kegg_df, 
                 stats = p3_paPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_papac_kegg)
```

```{r}
sigPathways <- p3_papac_kegg[padj < 0.05]$pathway
sigRes <- p3_papac_kegg[p3_papac_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_paPAC,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_papac_kegg_simplified <- p3_papac_kegg[p3_papac_kegg$pathway %in% main_pathways, ]

create_dt(p3_papac_kegg_simplified)

p3_papac_kegg_simplified$theme <- pathway_theme_map(p3_papac_kegg_simplified$pathway)
```

```{r, fig.width=20}
ggplot(p3_papac_kegg_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### longPAC

```{r}
p3_longPAC_kegg <- fgsea(pathways = kegg_df, 
                 stats = p3_longPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_longPAC_kegg)
```

```{r}
sigPathways <- p3_longPAC_kegg[padj < 0.05]$pathway
sigRes <- p3_longPAC_kegg[p3_longPAC_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_longPAC,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_longPAC_kegg_simplified <- p3_longPAC_kegg[p3_longPAC_kegg$pathway %in% main_pathways, ]

create_dt(p3_longPAC_kegg_simplified)

p3_longPAC_kegg_simplified$theme <- pathway_theme_map(p3_longPAC_kegg_simplified$pathway)
```

```{r, fig.width=20}
ggplot(p3_longPAC_kegg_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### paDMSO-ncDMSO

```{r}
p3_paDMSO_kegg <- fgsea(pathways = kegg_df, 
                 stats = p3_paDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_paDMSO_kegg)
```

```{r}
sigPathways <- p3_paDMSO_kegg[padj < 0.05]$pathway
sigRes <- p3_paDMSO_kegg[p3_paDMSO_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_paDMSO,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_paDMSO_kegg_simplified <- p3_paDMSO_kegg[p3_paDMSO_kegg$pathway %in% main_pathways, ]

create_dt(p3_paDMSO_kegg_simplified)

p3_paDMSO_kegg_simplified$theme <- pathway_theme_map(p3_paDMSO_kegg_simplified$pathway)
```

```{r, fig.width=20}
ggplot(p3_paDMSO_kegg_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

### Reactome {.tabset}

#### ncPAC - ncDMSO

```{r}
p3_ncpac_reactome <- fgsea(pathways = reactome_df, 
                 stats = p3_ncPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_ncpac_reactome)
```

```{r}
sigPathways <- p3_ncpac_reactome[padj < 0.05]$pathway
sigRes <- p3_ncpac_reactome[p3_ncpac_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_ncPAC,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_ncpac_reactome_simplified <- p3_ncpac_reactome[p3_ncpac_reactome$pathway %in% main_pathways, ]

create_dt(p3_ncpac_reactome_simplified)
```

```{r, fig.width=15}
p3_ncpac_reactome_simplified$theme <- pathway_theme_map(p3_ncpac_reactome_simplified$pathway)

ggplot(p3_ncpac_reactome_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### paPAC-paDMSO

```{r}
p3_papac_reactome <- fgsea(pathways = reactome_df, 
                 stats = p3_paPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_papac_reactome)
```

```{r}
sigPathways <- p3_papac_reactome[padj < 0.05]$pathway
sigRes <- p3_papac_reactome[p3_papac_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_paPAC,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_papac_reactome_simplified <- p3_papac_reactome[p3_papac_reactome$pathway %in% main_pathways, ]

create_dt(p3_papac_reactome_simplified)
```

```{r, fig.width=15}
p3_papac_reactome_simplified$theme <- pathway_theme_map(p3_papac_reactome_simplified$pathway)

ggplot(p3_papac_reactome_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### longPAC

```{r}
p3_longPAC_reactome <- fgsea(pathways = reactome_df, 
                 stats = p3_longPAC,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_longPAC_reactome)
```

```{r}
sigPathways <- p3_longPAC_reactome[padj < 0.05]$pathway
sigRes <- p3_longPAC_reactome[p3_longPAC_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_longPAC,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_longPAC_reactome_simplified <- p3_longPAC_reactome[p3_longPAC_reactome$pathway %in% main_pathways, ]

create_dt(p3_longPAC_reactome_simplified)
```

```{r, fig.width=15}
p3_longPAC_reactome_simplified$theme <- pathway_theme_map(p3_longPAC_reactome_simplified$pathway)

ggplot(p3_longPAC_reactome_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### paDMSO-ncDMSO

```{r}
p3_paDMSO_reactome <- fgsea(pathways = reactome_df, 
                 stats = p3_paDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_paDMSO_reactome)
```

```{r}
sigPathways <- p3_paDMSO_reactome[padj < 0.05]$pathway
sigRes <- p3_paDMSO_reactome[p3_paDMSO_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_paDMSO,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_paDMSO_reactome_simplified <- p3_paDMSO_reactome[p3_paDMSO_reactome$pathway %in% main_pathways, ]

create_dt(p3_paDMSO_reactome_simplified)
```

```{r, fig.width=15}
p3_paDMSO_reactome_simplified$theme <- pathway_theme_map(p3_paDMSO_reactome_simplified$pathway)

ggplot(p3_paDMSO_reactome_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```
