---
title: "ORG60 - Short term response"
output: html_document
date: "2026-06-09"
---

```{r setup}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.height=8, fig.width = 10)
```

```{r}
### loading required packages
library(tidyverse)
library(Seurat)
library(gtools)
library(tidyseurat)
library(DESeq2)
library(edgeR)
library(DT)
library(dplyr)
library(pheatmap)
library(EnhancedVolcano)
library(RColorBrewer)
library(ggvenn)
library(gplots)
library(clusterProfiler)
library(enrichplot)
library(msigdbr)
library(fgsea)
```

```{r}
### Functions 

source('/Users/dalvinikita/Desktop/macpie/R/filter_genes_by_expression.R')

### buttons for datatables 
create_dt_genes <- function(x){
  colnames <- c('logFC','AveExpr','t','adj.P.Val')
  df <- x %>%
    mutate(across(all_of(colnames),round, digits = 2)) %>%
    mutate(adj.P.Val = format(adj.P.Val, format='e', digits=4)) %>%
    mutate(P.Value = format(P.Value, format='e', digits=4))
  DT::datatable(df,
                extensions = 'Buttons',
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All"))))
}

create_dt <-  function(x){
  datatable(x,
            extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
}


plotRLE <- function(data, main = "", ylab = "Relative Log Expression") {
  # Calculate medians
  col <- brewer.pal(8, 'Set3')
  medians <- apply(data, 1, median, na.rm = TRUE)
  rle_data <- data - medians
  
  boxplot(as.data.frame(rle_data),
          outline = FALSE,
          las = 2,
          ylab = ylab,
          main = main,
          col = col)
  abline(h = 0, col = "red", lty = 2)
}

red_blue_palette <- colorRampPalette(brewer.pal(11, "RdBu"))(100)
```

```{r}
### loading genesets
reactome_df <-msigdbr(species = 'Homo sapiens', category = 'C2', subcollection = 'CP:REACTOME')
reactome_df <- reactome_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
#cat(paste("Loaded", length(reactome_df), "gene sets\n"))

go_df <-msigdbr(species = 'Homo sapiens', category = 'C5')
go_df <- go_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
#cat(paste("Loaded", length(go_df), "gene sets\n"))

kegg_df <-msigdbr(species = 'Homo sapiens', category = 'C2', subcollection = 'CP:KEGG_MEDICUS')
kegg_df <- kegg_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
#cat(paste("Loaded", length(kegg_df), "gene sets\n"))

hallmark_df <- msigdbr(species = 'Homo sapiens', category = 'H')
hallmark_df <- hallmark_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
#cat(paste("Loaded", length(hallmark_df), "gene sets\n"))
```


## Notes

Thresholds used for DE analysis:

• pvalue threshold = 0.05

• log FC threshold = log2(1.2) = 20% change in expression compared to control

```{r, echo=TRUE}
### Threshold Constants
p_value_threshold <- 0.05
```

## Data

```{r}
### Importing MACseq data
### Importing metadata
### Joining data in a Seurat object 

### import metadata
metadata<-read.csv("/Users/dalvinikita/Documents/13_MACseq_PMMSq057/metadata/PMMSq057_metadata.csv",header=T) %>%
  arrange("Barcode") %>%
  mutate(id=gsub("Plate","",Plate_ID)) %>%
  mutate(Treatment_Combination=paste0(PreTreatment,'_',substr(Treatment_1,1,4))) %>%
  mutate(label=paste0(Treatment_Combination,"_",Well_ID)) %>%
  mutate(Concentration_1=as.factor(Concentration_1)) %>%
  mutate(group = paste0(Treatment_Combination,'.',SPLINTR))

row.names(metadata)<-metadata$label

### load in count data 
raw_counts <- Read10X(data.dir = '/Users/dalvinikita/Documents/13_MACseq_PMMSq057/raw/starsolo/')

### filter for the minimum number of reads per gene per sample, as per edgeR
mac <- CreateSeuratObject(counts = raw_counts,
                          project = 'PMMSq057',
                          min.cells = 1,
                          min.features = 1)

mac[["percent.mt"]] <- PercentageFeatureSet(mac, pattern = "^mt-|^MT-")
mac[["percent.ribo"]] <- PercentageFeatureSet(mac, pattern = "^Rp[slp][[:digit:]]|^Rpsa|^RP[SLP][[:digit:]]|^RPSA")

### add location metadata
mac<- mac %>%
  dplyr::inner_join(metadata,by=c(".cell"="Barcode"))

colnames(mac)<- mac$label

mac <- mac[,mac$Organoid=='Org60']
mac <- filter_genes_by_expression(mac,
                                       group_by = "Organoid",
                                       min_counts = 15,
                                       min_samples = 8)

### exclude mito and ribo genes from the counts
mac<-mac[!grepl("^MT-",rownames(mac))]
mac<-mac[!grepl("^RP[SL][[:digit:]]|^RPLP[[:digit:]]|^RPSA",rownames(mac))]

mac_filtered <- mac 

samples <- colnames(mac_filtered)
metadata_filtered <- metadata[samples,]
metadata_filtered <- droplevels(metadata_filtered)
```

A quick look at the metadata:

```{r}
create_dt(metadata_filtered[,c("Well_ID","Barcode","Organoid", "Treatment_Combination", "SPLINTR","PreTreatment", "Treatment_1")])
```

A quick look at the treatment groups across the plate:

```{r}
table(mac_filtered$Treatment_Combination)

table(mac_filtered$Organoid)
```
Checking data order:

```{r, fig.width=10, fig.height=6}
###   Select wells from mac that are not empty 
rawdata <- mac_filtered
sampleinfo <- metadata_filtered[is.na(metadata_filtered$Organoid)!=TRUE | metadata$Treatment_1!="Staurosporine",]
samples <- rownames(sampleinfo)
rawdata <- mac_filtered[,samples]

### Keep only DMSO treated wells
rawdata <- rawdata %>%
  filter(Treatment_1=="DMSO")

### Ensure counts data and sampleinfo are in the same order
order <- colnames(rawdata)
sampleinfo <- sampleinfo[order,]
order_check <- rownames(sampleinfo)==colnames(rawdata)
cat('Instances where metadata sample matches counts matric sample (checking correct order):',"\n", order_check)
```

## Analysis

Differential expression for ORG60 after 48H treatment with either Paclitaxel or SN38.

Factors to include in design matrix: 

• Drug Treatment (SN38, Paclitaxel, or DMSO)

• Concentration (Different for each drug)

Design matrix:


```{r}
### selecting metadata and countdata relevant for ORG 60
### subset metadata to include only ORG60 cells 
columns <- metadata_filtered[metadata_filtered$Organoid=='Org60' & metadata_filtered$Treatment_1!="STAUR",]
sampleinfo <- columns
samples <- c(rownames(sampleinfo))
counts_matrix <- mac_filtered[,samples]
order <- colnames(counts_matrix)
sampleinfo <- sampleinfo[colnames(counts_matrix),]

### factors to include in deisgn matrix: SPLINTR group and treatment, concentration
treatment <- as.factor(sampleinfo$Treatment_1)
concentration <- as.factor(sampleinfo$Concentration_1)

### set up design matrix
### without intercept (means model)
design <- model.matrix(~0+treatment)
rownames(design) <- rownames(sampleinfo)
create_dt(design)
org60_design <- design

PACvsDMSO <- makeContrasts(treatmentPACLITAXEL-treatmentDMSO,levels=colnames(design))
SN38vsDMSO <- makeContrasts(treatmentSN38-treatmentDMSO,levels=colnames(design))
DRUGvsDMSO <- makeContrasts((treatmentPACLITAXEL+treatmentSN38)/2-treatmentDMSO,levels=colnames(design))
```

### Filtering + Normalisation

```{r, fig.height=5, fig.width=5}
### Normalising data
counts<-counts_matrix@assays$RNA$counts
counts <- DGEList(counts = counts) %>%
  calcNormFactors(method = 'TMMwsp')
v <- voom(counts, design, plot = T)
org60_voom_mat <- v
```

RLE of all ORG60 wells before and after normalisation:

```{r, fig.width=16}
par(mfrow=c(1,2))
plotRLE(counts$counts, main='RLE of all wells - raw')
plotRLE(v$E, main='RLE of all wells - normalised')
```

### Differential Expression {.tabset}

#### PACLITAXEL - DMSO 

##### DEGs

Differentially expressed genes between PACLITAXEL and DMSO in ORG60:

```{r}
### following limma workflow
### difference between PAC and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, PACvsDMSO)
fit <- treat(fit, lfc=log(1.2))
org60_PAC_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org60_PACvsDMSO_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')
#write.csv(all_genes, file = 'data/DEgenes/P3/ORG60_PACvsDMSO.csv')

create_dt_genes(org60_PACvsDMSO_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org60_PACvsDMSO_DEgenes)))
up <- paste('• UPREGULATED:',sum(org60_PACvsDMSO_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org60_PACvsDMSO_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

##### Volcano Plot

Volcano Plot of DEGs between Paclitaxel and DMSO with top 10 DEGs labelled: 

```{r, fig.height=8, fig.width=6}
### Assigning different colours for upregulated/downregulates genes
keyvals <- ifelse(
    all_genes$logFC < -1, 'darkred',
      ifelse(all_genes$logFC > 1, 'darkgreen',
        'black'))
  keyvals[is.na(keyvals)] <- 'black'
  names(keyvals)[keyvals == 'darkgreen'] <- 'UP'
  names(keyvals)[keyvals == 'black'] <- 'NO'
  names(keyvals)[keyvals == 'darkred'] <- 'DOWN'
  
### Volcano plot 
volcano_plot <- EnhancedVolcano(all_genes,
                                title="PACLITAXEL VS DMSO",
                                subtitle="ORG60",
                                lab=rownames(all_genes),
                                selectLab = rownames(org60_PACvsDMSO_DEgenes)[1:10],
                                boxedLabels = TRUE,
                                drawConnectors = TRUE,
                                arrowheads = FALSE,
                                pointSize = 3,
                                colAlpha = 1, 
                                x='logFC', 
                                y='adj.P.Val', 
                                pCutoff=0.05,
                                FCcutoff = 1,
                                border="full",
                                legendPosition = "bottom",
                                colCustom = keyvals,
                                ylim = c(0,5))
volcano_plot
```

##### Heatmap

Heatmap of all DEGs between Paclitaxel and DMSO

```{r, fig.height=10, fig.width=10}

### Assigning genes of interest from DEgenes table
### Formatting counts to log(CPM) for heatmap
### subsetting wells of interest
genes <- rownames(org60_PACvsDMSO_DEgenes)
samples_of_interest <- sampleinfo[sampleinfo$Treatment_1=='PACLITAXEL'|sampleinfo$Treatment_1=='DMSO',c('label')]
heatmap_counts <- as.data.frame(v$E[,samples_of_interest])


### Setting up annotation data frame
type_annotation <- data.frame(Drug = as.factor(sampleinfo$Treatment_1),
                              Concentration = as.factor(sampleinfo$Concentration_1))
rownames(type_annotation) <- rownames(sampleinfo)

pheatmap(heatmap_counts[genes,],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         color=rev(red_blue_palette))
     
```


##### Pathway Analysis {.tabset}

```{r}
# ### setting up ORG66 DE data
# df <- read_csv('data/DEgenes/P3/ORG60_PACvsDMSO.csv')
# df <- df %>%
#   dplyr::rename(gene_symbol = `...1`)
# head(df)
# 
# ### calculate rankings 
# org60_pac <- sign(df$logFC)*(-log10(df$P.Value)) 
# names(org60_pac) <- df$gene_symbol 
```

###### GO

```{r}
# org66_pac_go <- fgsea(pathways = go_df, 
#                  stats = org60_pac,
#                  scoreType = 'std',
#                  minSize = 10,
#                  maxSize = 500,
#                  nproc = 1)
# saveRDS(org66_pac_go, file='data/pathwayAnalysis/P3/ORG60_PACvsDMSO_go.rds')

res <- readRDS('data/pathwayAnalysis/P3/ORG60_PACvsDMSO_go.rds')
create_dt(res)
```

###### Reactome

```{r}
# org66_pac_reactome <- fgsea(pathways = reactome_df,
#                  stats = org60_pac,
#                  scoreType = 'std',
#                  minSize = 10,
#                  maxSize = 500,
#                  nproc = 1)
# saveRDS(org66_pac_reactome, file='data/pathwayAnalysis/P3/ORG60_PACvsDMSO_reactome.rds')

res <- readRDS('data/pathwayAnalysis/P3/ORG60_PACvsDMSO_reactome.rds')
create_dt(res)
```

###### KEGG

```{r}
# org66_pac_kegg <- fgsea(pathways = kegg_df,
#                  stats = org60_pac,
#                  scoreType = 'std',
#                  minSize = 10,
#                  maxSize = 500,
#                  nproc = 1)
# saveRDS(org66_pac_kegg, file='data/pathwayAnalysis/P3/ORG60_PACvsDMSO_kegg.rds')

res <- readRDS('data/pathwayAnalysis/P3/ORG60_PACvsDMSO_kegg.rds')
create_dt(res)
```

#### SN38 - DMSO 

##### DEGs

Differentially expressed genes between SN38 and DMSO in ORG60:

```{r}
### following limma workflow
### difference between SN38 and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, SN38vsDMSO)
fit <- treat(fit, lfc=log(1.2))
org60_SN38_fit <- fit
all_genes <- topTreat(fit, number=Inf)
org60_SN38vsDMSO_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')
#write.csv(all_genes, file = 'data/DEgenes/P3/ORG60_SN38vsDMSO.csv')

create_dt_genes(org60_SN38vsDMSO_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org60_SN38vsDMSO_DEgenes)))
up <- paste('• UPREGULATED:',sum(org60_SN38vsDMSO_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org60_SN38vsDMSO_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

##### Volcano Plot 

Volcano plot of DEGs between SN38 and DMSO with top 10 DEGs labelled:

```{r, fig.height=8, fig.width=6}
### Assigning different colours for upregulated/downregulates genes
keyvals <- ifelse(
    all_genes$logFC < -1, 'darkred',
      ifelse(all_genes$logFC > 1, 'darkgreen',
        'black'))
  keyvals[is.na(keyvals)] <- 'black'
  names(keyvals)[keyvals == 'darkgreen'] <- 'UP'
  names(keyvals)[keyvals == 'black'] <- 'NO'
  names(keyvals)[keyvals == 'darkred'] <- 'DOWN'
  
### Volcano plot 
volcano_plot <- EnhancedVolcano(all_genes,
                                title="SN38 VS DMSO",
                                subtitle="ORG60",
                                lab=rownames(all_genes),
                                selectLab = rownames(org60_SN38vsDMSO_DEgenes)[1:10],
                                boxedLabels = TRUE,
                                drawConnectors = TRUE,
                                arrowheads = FALSE,
                                pointSize = 3,
                                colAlpha = 1, 
                                x='logFC', 
                                y='adj.P.Val', 
                                pCutoff=0.05,
                                FCcutoff = 1,
                                border="full",
                                legendPosition = "bottom",
                                colCustom = keyvals)
volcano_plot
```

##### Heatmap {.tabset}

###### Top 100

Heatmap of top 100 DEGs between SN38 and DMSO:

```{r, fig.height=10, fig.width=10}

### Assigning genes of interest from DEgenes table
### Formatting counts to log(CPM) for heatmap
### subsetting wells of interest
genes <- rownames(org60_SN38vsDMSO_DEgenes)
samples_of_interest <- sampleinfo[sampleinfo$Treatment_1=='SN38'|sampleinfo$Treatment_1=='DMSO',c('label')]
heatmap_counts <- as.data.frame(v$E[,samples_of_interest])


### Setting up annotation data frame
type_annotation <- data.frame(Drug = as.factor(sampleinfo$Treatment_1),
                              Concentration = as.factor(sampleinfo$Concentration_1))

rownames(type_annotation) <- rownames(sampleinfo)

pheatmap(heatmap_counts[genes[1:100],],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         color=rev(red_blue_palette))
```

###### Top 500

Heatmap of top 500 DEGs between SN38 and DMSO:

```{r, fig.height=13, fig.width=10}
pheatmap(heatmap_counts[genes[1:500],],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         color=rev(red_blue_palette))
```

##### Pathway Analysis {.tabset}

```{r}
# ### setting up ORG60 DE data
# df <- read_csv('data/DEgenes/P3/ORG60_SN38vsDMSO.csv')
# df <- df %>%
#   dplyr::rename(gene_symbol = `...1`)
# head(df)
# 
# ### calculate rankings 
# org60_sn38 <- sign(df$logFC)*(-log10(df$P.Value)) 
# names(org60_sn38) <- df$gene_symbol 
```

###### GO

```{r}
# org60_sn38_go <- fgsea(pathways = go_df,
#                  stats = org60_sn38,
#                  scoreType = 'std',
#                  minSize = 10,
#                  maxSize = 500,
#                  nproc = 1)
# saveRDS(org60_sn38_go, file='data/pathwayAnalysis/P3/ORG60_SN38vsDMSO_go.rds')

res <- readRDS('data/pathwayAnalysis/P3/ORG60_SN38vsDMSO_go.rds')
create_dt(res)
```

###### Reactome

```{r}
# org60_sn38_reactome <- fgsea(pathways = reactome_df,
#                  stats = org60_sn38,
#                  scoreType = 'std',
#                  minSize = 10,
#                  maxSize = 500,
#                  nproc = 1)
# saveRDS(org60_sn38_reactome, file='data/pathwayAnalysis/P3/ORG60_SN38vsDMSO_reactome.rds')

res <- readRDS('data/pathwayAnalysis/P3/ORG60_SN38vsDMSO_reactome.rds')
create_dt(res)
```

###### KEGG

```{r}
# org60_sn38_kegg <- fgsea(pathways = kegg_df,
#                  stats = org60_sn38,
#                  scoreType = 'std',
#                  minSize = 10,
#                  maxSize = 500,
#                  nproc = 1)
# saveRDS(org60_sn38_kegg, file='data/pathwayAnalysis/P3/ORG60_SN38vsDMSO_kegg.rds')

res <- readRDS('data/pathwayAnalysis/P3/ORG60_SN38vsDMSO_kegg.rds')
create_dt(res)
```

#### DRUG - DMSO

##### DEGs

Differentially expressed genes between (PACLITAXEL+SN38) and DMSO in ORG60:

```{r}
### following limma workflow
### difference between both drugs and DMSO 
fit <- lmFit(v, design)
fit <- contrasts.fit(fit, DRUGvsDMSO)
fit <- treat(fit, lfc=log(1.2))
all_genes <- topTreat(fit, number=Inf)
org60_DRUGvsDMSO_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p')
#write.csv(all_genes, file = 'data/DEgenes/P3/ORG60_DRUGvsDMSO.csv')

create_dt_genes(org60_DRUGvsDMSO_DEgenes)

### Print summary of DE genes table
total <- paste('• TOTAL DE genes:',length(rownames(org60_DRUGvsDMSO_DEgenes)))
up <- paste('• UPREGULATED:',sum(org60_DRUGvsDMSO_DEgenes$logFC>0))
down <- paste('• DOWNREGULATED:',sum(org60_DRUGvsDMSO_DEgenes$logFC<0))
summary=paste (total, up, down, sep="\n")
cat(summary[1])
```

##### Volcano plot

Volcano Plot of DEGs between both drugs and DMSO with top 20 DEGs labelled:

```{r, fig.height=8, fig.width=6}
### Assigning different colours for upregulated/downregulates genes
keyvals <- ifelse(
    all_genes$logFC < -1, 'darkred',
      ifelse(all_genes$logFC > 1, 'darkgreen',
        'black'))
  keyvals[is.na(keyvals)] <- 'black'
  names(keyvals)[keyvals == 'darkgreen'] <- 'UP'
  names(keyvals)[keyvals == 'black'] <- 'NO'
  names(keyvals)[keyvals == 'darkred'] <- 'DOWN'
  
### Volcano plot 
volcano_plot <- EnhancedVolcano(all_genes,
                                title="(PACLITAXEL + SN38) VS DMSO",
                                subtitle="ORG60",
                                lab=rownames(all_genes),
                                selectLab = rownames(org60_DRUGvsDMSO_DEgenes)[1:20],
                                boxedLabels = TRUE,
                                drawConnectors = TRUE,
                                arrowheads = FALSE,
                                pointSize = 3,
                                colAlpha = 1, 
                                x='logFC', 
                                y='adj.P.Val', 
                                pCutoff=0.05,
                                FCcutoff = 1,
                                border="full",
                                legendPosition = "bottom",
                                colCustom = keyvals)
volcano_plot
```

##### Heatmap {.tabset}

###### Top 100

Heatmap of top 100 DEGs between both drugs and DMSO:

```{r, fig.height=12, fig.width=12}

### Assigning genes of interest from DEgenes table
### Formatting counts to log(CPM) for heatmap
### subsetting wells of interest
genes <- rownames(org60_DRUGvsDMSO_DEgenes)
heatmap_counts <- as.data.frame(v$E)


### Setting up annotation data frame
type_annotation <- data.frame(Drug = as.factor(sampleinfo$Treatment_1),
                              Concentration = as.factor(sampleinfo$Concentration_1))
rownames(type_annotation) <- rownames(sampleinfo)

pheatmap(heatmap_counts[genes[1:100],],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         color=rev(red_blue_palette),
         cellwidth = 40,
         cellheight = 6)
     
```

###### Top 500

Heatmap of top 500 DEGs between both drugs and DMSO:

```{r, fig.height=15, fig.width=15}

pheatmap(heatmap_counts[genes[1:500],],
         border_color = 'NA',
         annotation_col = type_annotation,
         treeheight_row = 0,
         color=rev(red_blue_palette),
         cellwidth = 80)
     
```

##### Pathway Analysis {.tabset}

```{r}
# ### setting up ORG66 DE data
# df <- read_csv('data/DEgenes/P3/ORG60_DRUGvsDMSO.csv')
# df <- df %>%
#   dplyr::rename(gene_symbol = `...1`)
# head(df)
# 
# ### calculate rankings 
# org60_drug <- sign(df$logFC)*(-log10(df$P.Value)) 
# names(org60_drug) <- df$gene_symbol 
```

###### GO

```{r}
# org60_drug_go <- fgsea(pathways = go_df,
#                  stats = org60_drug,
#                  scoreType = 'std',
#                  minSize = 10,
#                  maxSize = 500,
#                  nproc = 1)
# saveRDS(org60_drug_go, file='data/pathwayAnalysis/P3/ORG60_DRUGvsDMSO_go.rds')

res <- readRDS('data/pathwayAnalysis/P3/ORG60_DRUGvsDMSO_go.rds')
create_dt(res)
```

###### Reactome

```{r}
# org60_drug_reactome <- fgsea(pathways = reactome_df,
#                  stats = org60_drug,
#                  scoreType = 'std',
#                  minSize = 10,
#                  maxSize = 500,
#                  nproc = 1)
# saveRDS(org60_drug_reactome, file='data/pathwayAnalysis/P3/ORG60_DRUGvsDMSO_reactome.rds')

res <- readRDS('data/pathwayAnalysis/P3/ORG60_DRUGvsDMSO_reactome.rds')
create_dt(res)
```

###### KEGG

```{r}
# org60_drug_kegg <- fgsea(pathways = kegg_df,
#                  stats = org60_drug,
#                  scoreType = 'std',
#                  minSize = 10,
#                  maxSize = 500,
#                  nproc = 1)
# saveRDS(org60_drug_kegg, file='data/pathwayAnalysis/P3/ORG60_DRUGvsDMSO_kegg.rds')

res <- readRDS('data/pathwayAnalysis/P3/ORG60_DRUGvsDMSO_kegg.rds')
create_dt(res)
```
