---
title: "ORG66_SN38_all"
output: html_document
date: "2026-06-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 12, fig.height = 6)
```

```{r}
library(tidyverse)
library(RColorBrewer)
library(fgsea)
library(msigdbr)
library(DT)
library(ggpubr)
library(dplyr)
library(tidyverse)
library(RColorBrewer)
library(ggvenn)
library(gplots)
library(tidyr)
library(ggplot2)

library(clusterProfiler)
library(DOSE)
library(enrichplot)
library(org.Hs.eg.db)
library(gsubfn)

```

```{r}
create_dt <-  function(x){
  datatable(x,
            extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
}
```

```{r}
### loading genesets
reactome_df <-msigdbr(species = 'Homo sapiens', category = 'C2', subcollection = 'CP:REACTOME')
reactome_df <- reactome_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(reactome_df), "gene sets\n"))

go_df <-msigdbr(species = 'Homo sapiens', category = 'C5')
go_df <- go_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(go_df), "gene sets\n"))

kegg_df <-msigdbr(species = 'Homo sapiens', category = 'C2', subcollection = 'CP:KEGG_MEDICUS')
kegg_df <- kegg_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(kegg_df), "gene sets\n"))

hallmark_df <- msigdbr(species = 'Homo sapiens', category = 'H')
hallmark_df <- hallmark_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(hallmark_df), "gene sets\n"))
```



```{r}
pathway_theme_map <- function(pathway_names){

  case_when(
    grepl("IMMUNE|INTERFERON|INFLAMM|CYTOKINE|ANTIGEN|T_CELL|B_CELL|NK|MHC|TLR|INNATE|ADAPTIVE|COMPLEMENT",
          pathway_names, ignore.case = TRUE) ~ "Immune / Inflammation",
    grepl("SIGNAL|MAPK|ERK|JAK|STAT|PI3K|AKT|WNT|NOTCH|HEDGEHOG|TGF|NF.KB|RAS|RHO|GTPase",
          pathway_names, ignore.case = TRUE) ~ "Signal Transduction",
    grepl("CELL_CYCLE|MITOSIS|MEIOSIS|G1|S PHASE|G2|CHECKPOINT|APC.C|CDK|CYCLIN|REPLICATION|CHROMOSOME_SEGREGATION|CHROMOSOME_CONDENSATION|CHROMOSOME_SEPARATION",
          pathway_names, ignore.case = TRUE) ~ "Cell Cycle / Division",
    grepl("APOPTOSIS|PYROPTOSIS|NECROPTOSIS|FERROPTOSIS|AUTOPHAGY|DNA DAMAGE|STRESS|UNFOLDED PROTEIN",
          pathway_names, ignore.case = TRUE) ~ "Cell Death / Stress Response",
    grepl("METABOLISM|GLYCOLYSIS|TCA|OXIDATIVE PHOSPHORYLATION|FATTY ACID|CHOLESTEROL|LIPID|AMINO ACID|NUCLEOTIDE|METABOLIC|CATABOLIC",
          pathway_names, ignore.case = TRUE) ~ "Metabolism",
    grepl("TRANSLATION|RIBOSOME|SPLICING|RNA|TRANSCRIPTION|RNA POLYMERASE|MRNA|RRNA|TRNA|PROCESSING",
          pathway_names, ignore.case = TRUE) ~ "Gene Expression / RNA Processing",
    grepl("DNA REPAIR|REPLICATION|TELOMERE|CHROMATIN|NUCLEOSOME|EPIGEN|HISTONE|METHYLATION",
          pathway_names, ignore.case = TRUE) ~ "Genome Maintenance / Epigenetics",
    grepl("DEVELOP|DIFFERENTIATION|MORPHOGEN|EMBRYO|AXIS|NEUROGEN|CELL FATE",
          pathway_names, ignore.case = TRUE) ~ "Development / Differentiation",
    grepl("ECM|EXTRACELLULAR|INTEGRIN|ADHESION|COLLAGEN|MIGRATION|CYTOSKELETON|ACTIN|MICROTUBULE",
          pathway_names, ignore.case = TRUE) ~ "ECM / Adhesion / Cytoskeleton",
    grepl("VESICLE|ENDOCYTOSIS|EXOCYTOSIS|GOLGI|ER|LYSOSOME|PEROXISOME|TRANSPORT",
          pathway_names, ignore.case = TRUE) ~ "Transport / Organelle Function",
    grepl("CANCER|CARCINOMA|MELANOMA|LEUKEMIA|TUMOR|ONCOGEN|P53|ONCOGENIC",
          pathway_names, ignore.case = TRUE) ~ "Cancer / Disease",
    TRUE ~ "Other / Unclassified"
  )
}
```

## Notes

This analysis contains data for ORG66 SN38 treated samples (compared to DMSO) across all MACseq Plates (P1-P3). Differential expression analyses are derived from respected .Rmd/.html files previously published on github: 

[P1 Differential Expression]()
[P2 Differential Expression]()
[P3 Differential Expression]()

Pathway analysis is a ranked analysis using ALL GENES (even statistically non-significant) where rank = logFC x adj.p.val 

Genes are analysed across different databases (KEGG, Reactome, GO, Hallmark) and results should be compared across databases to identify shared signals. 

## Plate 1 {.tabset}

```{r}
### load DE results 
### setting up ORG66 DE data
df <- read.csv('data/DEgenes/P1/ORG66_ncSN38vsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p1_ncSN38 <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p1_ncSN38) <- df$gene_symbol 
```

### GO

```{r}
p1_ncSN38_go <- fgsea(pathways = go_df, 
                 stats = p1_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p1_ncSN38_go)
```

```{r}
sigPathways <- p1_ncSN38_go[padj < 0.05]$pathway
sigRes <- p1_ncSN38_go[p1_ncSN38_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p1_ncSN38,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p1_ncSN38_go_simplified <- p1_ncSN38_go[p1_ncSN38_go$pathway %in% main_pathways, ]

create_dt(p1_ncSN38_go_simplified)

### format dataframe
p1_ncSN38_go_simplified$ONTOLOGY <- p1_ncSN38_go_simplified$pathway
  
p1_ncSN38_go_simplified <- p1_ncSN38_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p1_ncSN38_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p1_ncSN38_go_simplified$theme <- pathway_theme_map(p1_ncSN38_go_simplified$pathway)
```

```{r}
ggplot(p1_ncSN38_go_simplified,
       aes(x=label, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
  theme_bw()
```


### KEGG

```{r}
p1_ncSN38_kegg <- fgsea(pathways = kegg_df, 
                 stats = p1_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p1_ncSN38_kegg)
```

```{r}
sigPathways <- p1_ncSN38_kegg[padj < 0.05]$pathway
sigRes <- p1_ncSN38_kegg[p1_ncSN38_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p1_ncSN38,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p1_ncSN38_kegg_simplified <- p1_ncSN38_kegg[p1_ncSN38_kegg$pathway %in% main_pathways, ]

create_dt(p1_ncSN38_kegg_simplified)
```

```{r}
p1_ncSN38_kegg_simplified$theme <- pathway_theme_map(p1_ncSN38_kegg_simplified$pathway)
```

```{r, fig.height=3}
ggplot(p1_ncSN38_kegg_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

### Reactome 

```{r}
p1_ncSN38_reactome <- fgsea(pathways = reactome_df, 
                 stats = p1_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p1_ncSN38_reactome)
```

```{r}
sigPathways <- p1_ncSN38_reactome[padj < 0.05]$pathway
sigRes <- p1_ncSN38_reactome[p1_ncSN38_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p1_ncSN38,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p1_ncSN38_reactome_simplified <- p1_ncSN38_reactome[p1_ncSN38_reactome$pathway %in% main_pathways, ]

create_dt(p1_ncSN38_reactome_simplified)
```

```{r}
p1_ncSN38_reactome_simplified$theme <- pathway_theme_map(p1_ncSN38_reactome_simplified$pathway)
p1_ncSN38_reactome_simplified <- p1_ncSN38_reactome_simplified %>%
  arrange(NES)

ggplot(p1_ncSN38_reactome_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```


## Plate 2 {.tabset}

```{r}
### ncSN38
### load DE results 
df <- read.csv('data/DEgenes/P2/ORG66_ncSN38vsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p2_ncSN38 <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p2_ncSN38) <- df$gene_symbol 

### snSN38
### load DE results 
df <- read.csv('data/DEgenes/P2/ORG66_snSN38vsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p2_snSN38 <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p2_snSN38) <- as.factor(df$gene_symbol)

### longSN38
### load DE results 
df <- read.csv('data/DEgenes/P2/ORG66_longSN38.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p2_longSN38 <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p2_longSN38) <- df$gene_symbol 

### snDMSO
### load DE results 
df <- read.csv('data/DEgenes/P2/ORG66_snDMSOvsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p2_snDMSO <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p2_snDMSO) <- df$gene_symbol 
```

### GO {.tabset}

#### ncSN38 - ncDMSO

```{r}
p2_ncSN38_go <- fgsea(pathways = go_df, 
                 stats = p2_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_ncSN38_go)
```

```{r}
sigPathways <- p2_ncSN38_go[padj < 0.05]$pathway
sigRes <- p2_ncSN38_go[p2_ncSN38_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_ncSN38,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_ncSN38_go_simplified <- p2_ncSN38_go[p2_ncSN38_go$pathway %in% main_pathways, ]

create_dt(p2_ncSN38_go_simplified)

### format dataframe
p2_ncSN38_go_simplified$ONTOLOGY <- p2_ncSN38_go_simplified$pathway
  
p2_ncSN38_go_simplified <- p2_ncSN38_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p2_ncSN38_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p2_ncSN38_go_simplified$theme <- pathway_theme_map(p2_ncSN38_go_simplified$pathway)
```

```{r}
ggplot(p2_ncSN38_go_simplified,
       aes(x=label, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
  theme_bw()
```

#### snSN38-snDMSO

```{r}
p2_snSN38_go <- fgsea(pathways = go_df, 
                 stats = p2_snSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_snSN38_go)
```

```{r}
sigPathways <- p2_snSN38_go[padj < 0.05]$pathway
sigRes <- p2_snSN38_go[p2_snSN38_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_snSN38,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_snSN38_go_simplified <- p2_snSN38_go[p2_snSN38_go$pathway %in% main_pathways, ]

create_dt(p2_snSN38_go_simplified)

### format dataframe
p2_snSN38_go_simplified$ONTOLOGY <- p2_snSN38_go_simplified$pathway
  
p2_snSN38_go_simplified <- p2_snSN38_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p2_snSN38_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p2_snSN38_go_simplified$theme <- pathway_theme_map(p2_snSN38_go_simplified$pathway)
```

```{r}
ggplot(p2_snSN38_go_simplified,
       aes(x=label, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
  theme_bw()
```

#### longSN38

```{r}
p2_longSN38_go <- fgsea(pathways = go_df, 
                 stats = p2_longSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_longSN38_go)
```

```{r}
sigPathways <- p2_longSN38_go[padj < 0.05]$pathway
sigRes <- p2_longSN38_go[p2_longSN38_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_longSN38,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_longSN38_go_simplified <- p2_longSN38_go[p2_longSN38_go$pathway %in% main_pathways, ]

create_dt(p2_longSN38_go_simplified)

### format dataframe
p2_longSN38_go_simplified$ONTOLOGY <- p2_longSN38_go_simplified$pathway
  
p2_longSN38_go_simplified <- p2_longSN38_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p2_longSN38_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p2_longSN38_go_simplified$theme <- pathway_theme_map(p2_longSN38_go_simplified$pathway)
```

```{r}
ggplot(p2_longSN38_go_simplified,
       aes(x=label, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
  theme_bw()
```

#### snDMSO-ncDMSO

```{r}
p2_snDMSO_go <- fgsea(pathways = go_df, 
                 stats = p2_snDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_snDMSO_go)
```

```{r}
sigPathways <- p2_snDMSO_go[padj < 0.05]$pathway
sigRes <- p2_snDMSO_go[p2_snDMSO_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_snDMSO,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_snDMSO_go_simplified <- p2_snDMSO_go[p2_snDMSO_go$pathway %in% main_pathways, ]

create_dt(p2_snDMSO_go_simplified)

### format dataframe
p2_snDMSO_go_simplified$ONTOLOGY <- p2_snDMSO_go_simplified$pathway
  
p2_snDMSO_go_simplified <- p2_snDMSO_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p2_snDMSO_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p2_snDMSO_go_simplified$theme <- pathway_theme_map(p2_snDMSO_go_simplified$pathway)
```

```{r }
ggplot(p2_snDMSO_go_simplified,
       aes(x=label, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
  theme_bw()
```

### KEGG {.tabset}

#### ncSN38 - ncDMSO

```{r}
p2_ncSN38_kegg <- fgsea(pathways = kegg_df, 
                 stats = p2_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_ncSN38_kegg)
```

```{r}
sigPathways <- p2_ncSN38_kegg[padj < 0.05]$pathway
sigRes <- p2_ncSN38_kegg[p2_ncSN38_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_ncSN38,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_ncSN38_kegg_simplified <- p2_ncSN38_kegg[p2_ncSN38_kegg$pathway %in% main_pathways, ]

create_dt(p2_ncSN38_kegg_simplified)

p2_ncSN38_kegg_simplified$theme <- pathway_theme_map(p2_ncSN38_kegg_simplified$pathway)
```

```{r, fig.height=3}
ggplot(p2_ncSN38_kegg_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### snSN38-snDMSO

```{r}
p2_snSN38_kegg <- fgsea(pathways = kegg_df, 
                 stats = p2_snSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_snSN38_kegg)
```

```{r}
sigPathways <- p2_snSN38_kegg[padj < 0.05]$pathway
sigRes <- p2_snSN38_kegg[p2_snSN38_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_snSN38,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_snSN38_kegg_simplified <- p2_snSN38_kegg[p2_snSN38_kegg$pathway %in% main_pathways, ]

create_dt(p2_snSN38_kegg_simplified)

p2_snSN38_kegg_simplified$theme <- pathway_theme_map(p2_snSN38_kegg_simplified$pathway)
```

```{r, fig.height=3}
ggplot(p2_snSN38_kegg_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### longSN38

```{r}
p2_longSN38_kegg <- fgsea(pathways = kegg_df, 
                 stats = p2_longSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_longSN38_kegg)
```

```{r}
sigPathways <- p2_longSN38_kegg[padj < 0.05]$pathway
sigRes <- p2_longSN38_kegg[p2_longSN38_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_longSN38,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_longSN38_kegg_simplified <- p2_longSN38_kegg[p2_longSN38_kegg$pathway %in% main_pathways, ]

create_dt(p2_longSN38_kegg_simplified)

p2_longSN38_kegg_simplified$theme <- pathway_theme_map(p2_longSN38_kegg_simplified$pathway)
```

```{r, fig.height=3}
ggplot(p2_longSN38_kegg_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### snDMSO-ncDMSO

```{r}
p2_snDMSO_kegg <- fgsea(pathways = kegg_df, 
                 stats = p2_snDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_snDMSO_kegg)
```

```{r}
sigPathways <- p2_snDMSO_kegg[padj < 0.05]$pathway
sigRes <- p2_snDMSO_kegg[p2_snDMSO_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_snDMSO,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_snDMSO_kegg_simplified <- p2_snDMSO_kegg[p2_snDMSO_kegg$pathway %in% main_pathways, ]

create_dt(p2_snDMSO_kegg_simplified)

p2_snDMSO_kegg_simplified$theme <- pathway_theme_map(p2_snDMSO_kegg_simplified$pathway)
```


```{r, fig.height=3}
ggplot(p2_snDMSO_kegg_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

### Reactome {.tabset}

#### ncSN38 - ncDMSO

```{r}
p2_ncSN38_reactome <- fgsea(pathways = reactome_df, 
                 stats = p2_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_ncSN38_reactome)
```

```{r}
sigPathways <- p2_ncSN38_reactome[padj < 0.05]$pathway
sigRes <- p2_ncSN38_reactome[p2_ncSN38_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_ncSN38,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_ncSN38_reactome_simplified <- p2_ncSN38_reactome[p2_ncSN38_reactome$pathway %in% main_pathways, ]

create_dt(p2_ncSN38_reactome_simplified)
```

```{r}
p2_ncSN38_reactome_simplified$theme <- pathway_theme_map(p2_ncSN38_reactome_simplified$pathway)

ggplot(p2_ncSN38_reactome_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### snSN38-snDMSO

```{r}
p2_snSN38_reactome <- fgsea(pathways = reactome_df, 
                 stats = p2_snSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_snSN38_reactome)
```

```{r}
sigPathways <- p2_snSN38_reactome[padj < 0.05]$pathway
sigRes <- p2_snSN38_reactome[p2_snSN38_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_snSN38,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_snSN38_reactome_simplified <- p2_snSN38_reactome[p2_snSN38_reactome$pathway %in% main_pathways, ]

create_dt(p2_snSN38_reactome_simplified)
```

```{r}
p2_snSN38_reactome_simplified$theme <- pathway_theme_map(p2_snSN38_reactome_simplified$pathway)

ggplot(p2_snSN38_reactome_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### longSN38

```{r}
p2_longSN38_reactome <- fgsea(pathways = reactome_df, 
                 stats = p2_longSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_longSN38_reactome)
```

```{r}
sigPathways <- p2_longSN38_reactome[padj < 0.05]$pathway
sigRes <- p2_longSN38_reactome[p2_longSN38_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_longSN38,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_longSN38_reactome_simplified <- p2_longSN38_reactome[p2_longSN38_reactome$pathway %in% main_pathways, ]

create_dt(p2_longSN38_reactome_simplified)
```

```{r}
p2_longSN38_reactome_simplified$theme <- pathway_theme_map(p2_longSN38_reactome_simplified$pathway)

ggplot(p2_longSN38_reactome_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### snDMSO-ncDMSO

```{r}
p2_snDMSO_reactome <- fgsea(pathways = reactome_df, 
                 stats = p2_snDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p2_snDMSO_reactome)
```

```{r}
sigPathways <- p2_snDMSO_reactome[padj < 0.05]$pathway
sigRes <- p2_snDMSO_reactome[p2_snDMSO_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p2_snDMSO,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p2_snDMSO_reactome_simplified <- p2_snDMSO_reactome[p2_snDMSO_reactome$pathway %in% main_pathways, ]

create_dt(p2_snDMSO_reactome_simplified)
```


```{r}
p2_snDMSO_reactome_simplified$theme <- pathway_theme_map(p2_snDMSO_reactome_simplified$pathway)

ggplot(p2_snDMSO_reactome_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

## Plate 3 {.tabset}

```{r}
### ncSN38
### load DE results 
df <- read.csv('data/DEgenes/P3/ORG66_ncSN38vsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p3_ncSN38 <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p3_ncSN38) <- df$gene_symbol 

### snSN38
### load DE results 
df <- read.csv('data/DEgenes/P3/ORG66_snSN38vsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p3_snSN38 <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p3_snSN38) <- df$gene_symbol 

### longSN38
### load DE results 
df <- read.csv('data/DEgenes/P3/ORG66_longSN38.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p3_longSN38 <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p3_longSN38) <- df$gene_symbol 

### snDMSO
### load DE results 
df <- read.csv('data/DEgenes/P3/ORG66_snDMSOvsDMSO.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### calculate rankings 
p3_snDMSO <- sign(df$logFC)*(-log10(df$P.Value)) 
names(p3_snDMSO) <- df$gene_symbol 
```

### GO {.tabset}

#### ncSN38 - ncDMSO

```{r}
p3_ncSN38_go <- fgsea(pathways = go_df, 
                 stats = p3_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_ncSN38_go)
```

```{r}
sigPathways <- p3_ncSN38_go[padj < 0.05]$pathway
sigRes <- p3_ncSN38_go[p3_ncSN38_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_ncSN38,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_ncSN38_go_simplified <- p3_ncSN38_go[p3_ncSN38_go$pathway %in% main_pathways, ]

create_dt(p3_ncSN38_go_simplified)

### format dataframe
p3_ncSN38_go_simplified$ONTOLOGY <- p3_ncSN38_go_simplified$pathway
  
p3_ncSN38_go_simplified <- p3_ncSN38_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p3_ncSN38_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p3_ncSN38_go_simplified$theme <- pathway_theme_map(p3_ncSN38_go_simplified$pathway)
```

```{r}
ggplot(p3_ncSN38_go_simplified[1:40],
       aes(x=label, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
  theme_bw()
```

#### snSN38-snDMSO

```{r}
p3_snSN38_go <- fgsea(pathways = go_df, 
                 stats = p3_snSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_snSN38_go)
```

```{r}
sigPathways <- p3_snSN38_go[padj < 0.05]$pathway
sigRes <- p3_snSN38_go[p3_snSN38_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_snSN38,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_snSN38_go_simplified <- p3_snSN38_go[p3_snSN38_go$pathway %in% main_pathways, ]

create_dt(p3_snSN38_go_simplified)

### format dataframe
p3_snSN38_go_simplified$ONTOLOGY <- p3_snSN38_go_simplified$pathway
  
p3_snSN38_go_simplified <- p3_snSN38_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p3_snSN38_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p3_snSN38_go_simplified$theme <- pathway_theme_map(p3_snSN38_go_simplified$pathway)
```

```{r}
ggplot(p3_snSN38_go_simplified,
       aes(x=label, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
  theme_bw()
```

#### longSN38

```{r}
p3_longSN38_go <- fgsea(pathways = go_df, 
                 stats = p3_longSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_longSN38_go)
```

```{r}
sigPathways <- p3_longSN38_go[padj < 0.05]$pathway
sigRes <- p3_longSN38_go[p3_longSN38_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_longSN38,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_longSN38_go_simplified <- p3_longSN38_go[p3_longSN38_go$pathway %in% main_pathways, ]

create_dt(p3_longSN38_go_simplified)

### format dataframe
p3_longSN38_go_simplified$ONTOLOGY <- p3_longSN38_go_simplified$pathway
  
p3_longSN38_go_simplified <- p3_longSN38_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p3_longSN38_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p3_longSN38_go_simplified$theme <- pathway_theme_map(p3_longSN38_go_simplified$pathway)
```

```{r}
ggplot(p3_longSN38_go_simplified[1:40],
       aes(x=label, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
  theme_bw()
```

#### snDMSO-ncDMSO

```{r}
p3_snDMSO_go <- fgsea(pathways = go_df, 
                 stats = p3_snDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_snDMSO_go)
```

```{r}
sigPathways <- p3_snDMSO_go[padj < 0.05]$pathway
sigRes <- p3_snDMSO_go[p3_snDMSO_go$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_snDMSO,
                                           pathways = go_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_snDMSO_go_simplified <- p3_snDMSO_go[p3_snDMSO_go$pathway %in% main_pathways, ]

create_dt(p3_snDMSO_go_simplified)

### format dataframe
p3_snDMSO_go_simplified$ONTOLOGY <- p3_snDMSO_go_simplified$pathway
  
p3_snDMSO_go_simplified <- p3_snDMSO_go_simplified %>%
  mutate(ONTOLOGY = gsub('GO', '', p3_snDMSO_go_simplified$ONTOLOGY)) %>%
  mutate(ONTOLOGY = substr(ONTOLOGY, 1,2)) %>%
  mutate(label= gsub("GOBP_|GOMF_|GOCC_|HP_", "", pathway)) %>%
  arrange(padj)

p3_snDMSO_go_simplified$theme <- pathway_theme_map(p3_snDMSO_go_simplified$pathway)
```

```{r, fig.width=15}
ggplot(p3_snDMSO_go_simplified,
       aes(x=label, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  facet_grid(vars(ONTOLOGY), scales = 'free', space='free')+
  theme_bw()
```

### KEGG {.tabset}

#### ncSN38 - ncDMSO

```{r}
p3_ncSN38_kegg <- fgsea(pathways = kegg_df, 
                 stats = p3_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_ncSN38_kegg)
```

```{r}
sigPathways <- p3_ncSN38_kegg[padj < 0.05]$pathway
sigRes <- p3_ncSN38_kegg[p3_ncSN38_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_ncSN38,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_ncSN38_kegg_simplified <- p3_ncSN38_kegg[p3_ncSN38_kegg$pathway %in% main_pathways, ]

create_dt(p3_ncSN38_kegg_simplified)

p3_ncSN38_kegg_simplified$theme <- pathway_theme_map(p3_ncSN38_kegg_simplified$pathway)
```

```{r,fig.height=3}
ggplot(p3_ncSN38_kegg_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### snSN38-snDMSO

```{r}
p3_snSN38_kegg <- fgsea(pathways = kegg_df, 
                 stats = p3_snSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_snSN38_kegg)
```

```{r}
sigPathways <- p3_snSN38_kegg[padj < 0.05]$pathway
sigRes <- p3_snSN38_kegg[p3_snSN38_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_snSN38,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_snSN38_kegg_simplified <- p3_snSN38_kegg[p3_snSN38_kegg$pathway %in% main_pathways, ]

create_dt(p3_snSN38_kegg_simplified)

p3_snSN38_kegg_simplified$theme <- pathway_theme_map(p3_snSN38_kegg_simplified$pathway)
```

```{r,  fig.height=3}
ggplot(p3_snSN38_kegg_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### longSN38

```{r}
p3_longSN38_kegg <- fgsea(pathways = kegg_df, 
                 stats = p3_longSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_longSN38_kegg)
```

```{r}
sigPathways <- p3_longSN38_kegg[padj < 0.05]$pathway
sigRes <- p3_longSN38_kegg[p3_longSN38_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_longSN38,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_longSN38_kegg_simplified <- p3_longSN38_kegg[p3_longSN38_kegg$pathway %in% main_pathways, ]

create_dt(p3_longSN38_kegg_simplified)

p3_longSN38_kegg_simplified$theme <- pathway_theme_map(p3_longSN38_kegg_simplified$pathway)
```

```{r  =3}
ggplot(p3_longSN38_kegg_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### snDMSO-ncDMSO

```{r}
p3_snDMSO_kegg <- fgsea(pathways = kegg_df, 
                 stats = p3_snDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_snDMSO_kegg)
```

```{r}
sigPathways <- p3_snDMSO_kegg[padj < 0.05]$pathway
sigRes <- p3_snDMSO_kegg[p3_snDMSO_kegg$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_snDMSO,
                                           pathways = kegg_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_snDMSO_kegg_simplified <- p3_snDMSO_kegg[p3_snDMSO_kegg$pathway %in% main_pathways, ]

create_dt(p3_snDMSO_kegg_simplified)

p3_snDMSO_kegg_simplified$theme <- pathway_theme_map(p3_snDMSO_kegg_simplified$pathway)
```

```{r,fig.height=3}
ggplot(p3_snDMSO_kegg_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

### Reactome {.tabset}

#### ncSN38 - ncDMSO

```{r}
p3_ncSN38_reactome <- fgsea(pathways = reactome_df, 
                 stats = p3_ncSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_ncSN38_reactome)
```

```{r}
sigPathways <- p3_ncSN38_reactome[padj < 0.05]$pathway
sigRes <- p3_ncSN38_reactome[p3_ncSN38_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_ncSN38,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_ncSN38_reactome_simplified <- p3_ncSN38_reactome[p3_ncSN38_reactome$pathway %in% main_pathways, ]

create_dt(p3_ncSN38_reactome_simplified)
```

```{r,fig.width=20, fig.height=8}
p3_ncSN38_reactome_simplified$theme <- pathway_theme_map(p3_ncSN38_reactome_simplified$pathway)

ggplot(p3_ncSN38_reactome_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### snSN38-snDMSO

```{r}
p3_snSN38_reactome <- fgsea(pathways = reactome_df, 
                 stats = p3_snSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_snSN38_reactome)
```

```{r}
sigPathways <- p3_snSN38_reactome[padj < 0.05]$pathway
sigRes <- p3_snSN38_reactome[p3_snSN38_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_snSN38,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_snSN38_reactome_simplified <- p3_snSN38_reactome[p3_snSN38_reactome$pathway %in% main_pathways, ]

create_dt(p3_snSN38_reactome_simplified)
```

```{r}
p3_snSN38_reactome_simplified$theme <- pathway_theme_map(p3_snSN38_reactome_simplified$pathway)

ggplot(p3_snSN38_reactome_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### longSN38

```{r}
p3_longSN38_reactome <- fgsea(pathways = reactome_df, 
                 stats = p3_longSN38,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_longSN38_reactome)
```

```{r}
sigPathways <- p3_longSN38_reactome[padj < 0.05]$pathway
sigRes <- p3_longSN38_reactome[p3_longSN38_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_longSN38,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_longSN38_reactome_simplified <- p3_longSN38_reactome[p3_longSN38_reactome$pathway %in% main_pathways, ]

create_dt(p3_longSN38_reactome_simplified)
```

```{r, fig.width=20, fig.height=10}
p3_longSN38_reactome_simplified$theme <- pathway_theme_map(p3_longSN38_reactome_simplified$pathway)

ggplot(p3_longSN38_reactome_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```

#### snDMSO-ncDMSO

```{r}
p3_snDMSO_reactome <- fgsea(pathways = reactome_df, 
                 stats = p3_snDMSO,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 0)

create_dt(p3_snDMSO_reactome)
```

```{r}
sigPathways <- p3_snDMSO_reactome[padj < 0.05]$pathway
sigRes <- p3_snDMSO_reactome[p3_snDMSO_reactome$pathway %in% sigPathways, ]

main_pathways <- collapsePathways(sigRes,
                                           stats = p3_snDMSO,
                                           pathways = reactome_df)

main_pathways <- main_pathways[["mainPathways"]]
p3_snDMSO_reactome_simplified <- p3_snDMSO_reactome[p3_snDMSO_reactome$pathway %in% main_pathways, ]

create_dt(p3_snDMSO_reactome_simplified)
```

```{r }
p3_snDMSO_reactome_simplified$theme <- pathway_theme_map(p3_snDMSO_reactome_simplified$pathway)

ggplot(p3_snDMSO_reactome_simplified,
       aes(x=pathway, y= NES, fill=theme)) +
  geom_bar(stat='identity') +
  coord_flip()+
  theme_bw()
```
