---
title: "Plate 1 Pathway Analysis"
output: html_document
date: "2026-05-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.height=4, fig.width = 6)
```

```{r}
library(tidyverse)
library(RColorBrewer)
library(fgsea)
library(msigdbr)
library(DT)
library(ggpubr)
library(dplyr)
library(tidyverse)
library(RColorBrewer)
library(ggvenn)
library(gplots)
library(tidyr)
library(ggplot2)

library(clusterProfiler)
library(DOSE)
library(enrichplot)
```

```{r}
create_dt <-  function(x){
  datatable(x,
            extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
}
```

# Genesets

```{r}

reactome_df <-msigdbr(species = 'Homo sapiens', category = 'C2', subcollection = 'CP:REACTOME')
reactome_df <- reactome_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(reactome_df), "gene sets\n"))

go_df <-msigdbr(species = 'Homo sapiens', category = 'C5')
go_df <- go_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(go_df), "gene sets\n"))

kegg_df <-msigdbr(species = 'Homo sapiens', category = 'C2', subcollection = 'CP:KEGG_MEDICUS')
kegg_df <- kegg_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(kegg_df), "gene sets\n"))

pid_df <- msigdbr(species = 'Homo sapiens', category = 'C2', subcollection = 'CP:PID')
pid_df <- pid_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(pid_df), "gene sets\n"))

hallmark_df <- msigdbr(species = 'Homo sapiens', category = 'H')
hallmark_df <- hallmark_df %>%
  split(x = .$gene_symbol, f = .$gs_name)
cat(paste("Loaded", length(hallmark_df), "gene sets\n"))
```

# Plate 1

```{r}
### setting up ORG38 DE data
### 1. Read in data -----------------------------------------------------------
df <- read.csv('data/DEgenes/VCFG_P1/SA_PMC228 Gorringe Lab  PMMSq018 - Differential expression analysis - org38.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org38_pac_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org38_pac_rankings) <- df$gene_symbol 


### setting up ORG49 DE data
### 1. Read in data -----------------------------------------------------------
df <- read.csv('data/DEgenes/VCFG_P1/SA_PMC228 Gorringe Lab  PMMSq018 - Differential expression analysis - org49.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org49_pac_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org49_pac_rankings) <- df$gene_symbol 


### setting up ORG66 DE data
### 1. Read in data -----------------------------------------------------------
df <- read.csv('data/DEgenes/VCFG_P1/SA_PMC228 Gorringe Lab  PMMSq018 - Differential expression analysis - org66.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org66_pac_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org66_pac_rankings) <- df$gene_symbol 

```

## Reactome {.tabset}

### ORG38 

```{r}
org38_reactome_res <- fgsea(pathways = reactome_df, 
                 stats = org38_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org38_reactome_res)
```


### ORG49

```{r}
org49_reactome_res <- fgsea(pathways = reactome_df, 
                 stats = org49_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org49_reactome_res)
```

### ORG66

```{r}
org66_reactome_res <- fgsea(pathways = reactome_df, 
                 stats = org66_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org66_reactome_res)
```

## GO {.tabset}

### ORG38 

```{r}
org38_GO_res <- fgsea(pathways = go_df, 
                 stats = org38_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org38_GO_res)
```

### ORG49

```{r}
org49_GO_res <- fgsea(pathways = go_df, 
                 stats = org49_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org49_GO_res)
```

### ORG66

```{r}
org66_GO_res <- fgsea(pathways = go_df, 
                 stats = org66_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org66_GO_res)
```

## KEGG {.tabset}

### ORG38 

```{r}
org38_kegg_res <- fgsea(pathways = kegg_df, 
                 stats = org38_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org38_kegg_res)
```

### ORG49 

```{r}
org49_kegg_res <- fgsea(pathways = kegg_df, 
                 stats = org49_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org49_kegg_res)
```

### ORG66

```{r}
org66_kegg_res <- fgsea(pathways = kegg_df, 
                 stats = org66_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org66_kegg_res)
```

## PID {.tabset}

### ORG38 

```{r}
org38_PID_res <- fgsea(pathways = pid_df, 
                 stats = org38_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org38_PID_res)
```

### ORG49 

```{r}
org49_PID_res <- fgsea(pathways = pid_df, 
                 stats = org49_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org49_PID_res)
```

### ORG66

```{r}
org66_PID_res <- fgsea(pathways = pid_df, 
                 stats = org66_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org66_PID_res)
```

## Hallmark {.tabset}

### ORG38 

```{r}
org38_hallmark_res <- fgsea(pathways = hallmark_df, 
                 stats = org38_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org38_hallmark_res)
```

### ORG49 

```{r}
org49_hallmark_res <- fgsea(pathways = hallmark_df, 
                 stats = org49_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org49_hallmark_res)
```

### ORG66 

```{r}
org66_hallmark_res <- fgsea(pathways = hallmark_df, 
                 stats = org66_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org66_hallmark_res)
```

# Plate 2

```{r}
### setting up ORG38 DE data
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P2/ORG38_PACPAC.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org38_pac_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org38_pac_rankings) <- df$gene_symbol 


### setting up ORG49 DE data
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P2/ORG49_PACPAC.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org49_pac_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org49_pac_rankings) <- df$gene_symbol 


### setting up ORG66 DE data
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P2/ORG66_PACPAC.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org66_pac_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org66_pac_rankings) <- df$gene_symbol 

```

## Reactome {.tabset}

### ORG38 

```{r}
org38_reactome_res <- fgsea(pathways = reactome_df, 
                 stats = org38_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org38_reactome_res)
```


### ORG49

```{r}
org49_reactome_res <- fgsea(pathways = reactome_df, 
                 stats = org49_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org49_reactome_res)
```

### ORG66

```{r}
org66_reactome_res <- fgsea(pathways = reactome_df, 
                 stats = org66_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org66_reactome_res)
```

### comparison

```{r}
x <- list(`org38`=org38_reactome_res[org38_reactome_res$padj<0.05,]$pathway,
          `org49`=org49_reactome_res[org49_reactome_res$padj<0.05,]$pathway,
          `org66`=org66_reactome_res[org66_reactome_res$padj<0.05,]$pathway)

x.table <- venn(x[c("org38","org49", "org66")], show.plot = FALSE)
x_attributes <- attributes(x.table)

ggvenn(x, columns = c("org38","org49", "org66"), 
       show_percentage = FALSE,
       text_size = 8)

common_pathways <- x_attributes$intersections$`org38:org66`
```

`r length(common_pathways)` pathways were found to be shared between org38 and org66 : 

  • `r common_pathways`
  
## GO {.tabset}

### ORG38 

```{r}
org38_GO_res <- fgsea(pathways = go_df, 
                 stats = org38_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org38_GO_res)
```

### ORG49

```{r}
org49_GO_res <- fgsea(pathways = go_df, 
                 stats = org49_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org49_GO_res)
```

### ORG66

```{r}
org66_GO_res <- fgsea(pathways = go_df, 
                 stats = org66_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org66_GO_res)
```

### comparison

```{r}
x <- list(`org38`=org38_GO_res[org38_GO_res$padj<0.05,]$pathway,
          `org49`=org49_GO_res[org49_GO_res$padj<0.05,]$pathway,
          `org66`=org66_GO_res[org66_GO_res$padj<0.05,]$pathway)

x.table <- venn(x[c("org38","org49", "org66")], show.plot = FALSE)
x_attributes <- attributes(x.table)

ggvenn(x, columns = c("org38","org49", "org66"), 
       show_percentage = FALSE,
       text_size = 8)

common_pathways <- x_attributes$intersections$`org38:org66`
```

`r length(common_pathways)` pathways were found to be shared between org38 and org66 : 

  • `r common_pathways`
  
## KEGG {.tabset}

### ORG38 

```{r}
org38_kegg_res <- fgsea(pathways = kegg_df, 
                 stats = org38_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org38_kegg_res)
```

### ORG49 

```{r}
org49_kegg_res <- fgsea(pathways = kegg_df, 
                 stats = org49_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org49_kegg_res)
```

### ORG66

```{r}
org66_kegg_res <- fgsea(pathways = kegg_df, 
                 stats = org66_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org66_kegg_res)
```

### comparison

```{r}
x <- list(`org38`=org38_kegg_res[org38_kegg_res$padj<0.05,]$pathway,
          `org49`=org49_kegg_res[org49_kegg_res$padj<0.05,]$pathway,
          `org66`=org66_kegg_res[org66_kegg_res$padj<0.05,]$pathway)

x.table <- venn(x[c("org38","org49", "org66")], show.plot = FALSE)
x_attributes <- attributes(x.table)

ggvenn(x, columns = c("org38","org49", "org66"), 
       show_percentage = FALSE,
       text_size = 8)

common_pathways <- x_attributes$intersections$`org38:org66`
```

`r length(common_pathways)` pathways were found to be shared between org38 and org66 : 

  • `r common_pathways`
  
## PID {.tabset}

### ORG38 

```{r}
org38_PID_res <- fgsea(pathways = pid_df, 
                 stats = org38_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org38_PID_res)
```

### ORG49 

```{r}
org49_PID_res <- fgsea(pathways = pid_df, 
                 stats = org49_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org49_PID_res)
```

### ORG66

```{r}
org66_PID_res <- fgsea(pathways = pid_df, 
                 stats = org66_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org66_PID_res)
```

### comparison

```{r}
x <- list(`org38`=org38_PID_res[org38_PID_res$padj<0.05,]$pathway,
          `org49`=org49_PID_res[org49_PID_res$padj<0.05,]$pathway,
          `org66`=org66_PID_res[org66_PID_res$padj<0.05,]$pathway)

x.table <- venn(x[c("org38","org49", "org66")], show.plot = FALSE)
x_attributes <- attributes(x.table)

ggvenn(x, columns = c("org38","org49", "org66"), 
       show_percentage = FALSE,
       text_size = 8)


```

## Hallmark {.tabset}

### ORG38 

```{r}
org38_hallmark_res <- fgsea(pathways = hallmark_df, 
                 stats = org38_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org38_hallmark_res)
```

### ORG49 

```{r}
org49_hallmark_res <- fgsea(pathways = hallmark_df, 
                 stats = org49_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org49_hallmark_res)
```

### ORG66 

```{r}
org66_hallmark_res <- fgsea(pathways = hallmark_df, 
                 stats = org66_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org66_hallmark_res)
```

### comparison

```{r}
x <- list(`org38`=org38_hallmark_res[org38_hallmark_res$padj<0.05,]$pathway,
          `org49`=org49_hallmark_res[org49_hallmark_res$padj<0.05,]$pathway,
          `org66`=org66_hallmark_res[org66_hallmark_res$padj<0.05,]$pathway)

x.table <- venn(x[c("org38","org49", "org66")], show.plot = FALSE)
x_attributes <- attributes(x.table)

ggvenn(x, columns = c("org38","org49", "org66"), 
       show_percentage = FALSE,
       text_size = 8)

common_pathways <- x_attributes$intersections$`org38:org66`
```

`r length(common_pathways)` pathways were found to be shared between org38 and org66 : 

  • `r common_pathways`
  
# Plate 3

```{r}
### setting up ORG38 DE data
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P3/ORG38_paPAC.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org38_pac_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org38_pac_rankings) <- df$gene_symbol 

### 3. create gene list for visualisation
org38_p3_gene_list <- df$logFC
names(org38_p3_gene_list) <- df$gene_symbol
org38_p3_gene_list<-na.omit(org38_p3_gene_list)
org38_p3_gene_list = sort(org38_p3_gene_list, decreasing = TRUE)

### setting up ORG49 DE data
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P3/ORG49_paPAC.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org49_pac_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org49_pac_rankings) <- df$gene_symbol 


### setting up ORG66 DE data
### 1. Read in data -----------------------------------------------------------
df <- read.csv('../data/DEgenes/P3/ORG66_paPAC.csv')
df <- df %>%
  dplyr::rename(gene_symbol = X)
head(df)

### 2. Prepare ranked list of genes -----------------------------------------------
org66_pac_rankings <- sign(df$logFC)*(-log10(df$P.Value)) 
names(org66_pac_rankings) <- df$gene_symbol 

```

## Reactome {.tabset}

### ORG38 

```{r}
org38_reactome_res <- fgsea(pathways = reactome_df, 
                 stats = org38_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org38_reactome_res)
```


### ORG49

```{r}
org49_reactome_res <- fgsea(pathways = reactome_df, 
                 stats = org49_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org49_reactome_res)
```

### ORG66

```{r}
org66_reactome_p3_res <- fgsea(pathways = reactome_df, 
                 stats = org66_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org66_reactome_p3_res)
```

### comparison

```{r}
x <- list(`org38`=org38_reactome_res[org38_reactome_res$padj<0.05,]$pathway,
          `org49`=org49_reactome_res[org49_reactome_res$padj<0.05,]$pathway,
          `org66`=org66_reactome_res[org66_reactome_res$padj<0.05,]$pathway)

x.table <- venn(x[c("org38","org49", "org66")], show.plot = FALSE)
x_attributes <- attributes(x.table)

ggvenn(x, columns = c("org38","org49", "org66"), 
       show_percentage = FALSE,
       text_size = 8)

common_pathways <- x_attributes$intersections$`org38:org66`
```

`r length(common_pathways)` pathways were found to be shared between org38 and org66 : 

  • `r common_pathways`

## GO {.tabset}

### ORG38 

```{r}
org38_GO_res <- fgsea(pathways = go_df, 
                 stats = org38_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org38_GO_res)
```

### ORG49

```{r}
org49_GO_res <- fgsea(pathways = go_df, 
                 stats = org49_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org49_GO_res)
```

### ORG66

```{r}
org66_p3_GO_res <- fgsea(pathways = go_df, 
                 stats = org66_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org66_p3_GO_res)
```

### comparison

```{r}
x <- list(`org38`=org38_GO_res[org38_GO_res$padj<0.05,]$pathway,
          `org49`=org49_GO_res[org49_GO_res$padj<0.05,]$pathway,
          `org66`=org66_GO_res[org66_GO_res$padj<0.05,]$pathway)

x.table <- venn(x[c("org38","org49", "org66")], show.plot = FALSE)
x_attributes <- attributes(x.table)

ggvenn(x, columns = c("org38","org49", "org66"), 
       show_percentage = FALSE,
       text_size = 8)

common_pathways <- x_attributes$intersections$`org38:org66`
```

`r length(common_pathways)` pathways were found to be shared between org38 and org66 : 

  • `r common_pathways`
  
## KEGG {.tabset}

### ORG38 

```{r}
org38_kegg_res <- fgsea(pathways = kegg_df, 
                 stats = org38_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org38_kegg_res)
```

### ORG49 

```{r}
org49_kegg_res <- fgsea(pathways = kegg_df, 
                 stats = org49_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org49_kegg_res)
```

### ORG66

```{r}
org66_kegg_res <- fgsea(pathways = kegg_df, 
                 stats = org66_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org66_kegg_res)
```

### comparison

```{r}
x <- list(`org38`=org38_kegg_res[org38_kegg_res$padj<0.05,]$pathway,
          `org49`=org49_kegg_res[org49_kegg_res$padj<0.05,]$pathway,
          `org66`=org66_kegg_res[org66_kegg_res$padj<0.05,]$pathway)

x.table <- venn(x[c("org38","org49", "org66")], show.plot = FALSE)
x_attributes <- attributes(x.table)

ggvenn(x, columns = c("org38","org49", "org66"), 
       show_percentage = FALSE,
       text_size = 8)

common_pathways <- x_attributes$intersections$`org38:org66`
```

`r length(common_pathways)` pathways were found to be shared between org38 and org66 : 

  • `r common_pathways`
  
## PID {.tabset}

### ORG38 

```{r}
org38_PID_res <- fgsea(pathways = pid_df, 
                 stats = org38_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org38_PID_res)
```

### ORG49 

```{r}
org49_PID_res <- fgsea(pathways = pid_df, 
                 stats = org49_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org49_PID_res)
```

### ORG66

```{r}
org66_PID_res <- fgsea(pathways = pid_df, 
                 stats = org66_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org66_PID_res)
```

### comparison

```{r}
x <- list(`org38`=org38_PID_res[org38_PID_res$padj<0.05,]$pathway,
          `org49`=org49_PID_res[org49_PID_res$padj<0.05,]$pathway,
          `org66`=org66_PID_res[org66_PID_res$padj<0.05,]$pathway)

x.table <- venn(x[c("org38","org49", "org66")], show.plot = FALSE)
x_attributes <- attributes(x.table)

ggvenn(x, columns = c("org38","org49", "org66"), 
       show_percentage = FALSE,
       text_size = 8)


```


## Hallmark {.tabset}

### ORG38 

```{r}
org38_hallmark_res <- fgsea(pathways = hallmark_df, 
                 stats = org38_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org38_hallmark_res)
```

### ORG49 

```{r}
org49_hallmark_res <- fgsea(pathways = hallmark_df, 
                 stats = org49_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org49_hallmark_res)
```

### ORG66 

```{r}
org66_p3_hallmark_res <- fgsea(pathways = hallmark_df, 
                 stats = org66_pac_rankings,
                 scoreType = 'std',
                 minSize = 10,
                 maxSize = 500,
                 nproc = 1)

create_dt(org66_p3_hallmark_res)
```

### comparison

```{r}
x <- list(`org38`=org38_hallmark_res[org38_hallmark_res$padj<0.05,]$pathway,
          `org49`=org49_hallmark_res[org49_hallmark_res$padj<0.05,]$pathway,
          `org66`=org66_hallmark_res[org66_hallmark_res$padj<0.05,]$pathway)

x.table <- venn(x[c("org38","org49", "org66")], show.plot = FALSE)
x_attributes <- attributes(x.table)

ggvenn(x, columns = c("org38","org49", "org66"), 
       show_percentage = FALSE,
       text_size = 8)

common_pathways <- x_attributes$intersections$`org38:org66`
```

`r length(common_pathways)` pathways were found to be shared between org38 and org66 : 

  • `r common_pathways`
  
```{r}
plotEnrichment(hallmark_df[['HALLMARK_MYC_TARGETS_V1']],
               org38_pac_rankings) + 
  labs(title ='HALLMARK_MYC_TARGETS_V1')

plotEnrichment(hallmark_df[['HALLMARK_MYC_TARGETS_V1']],
               org66_pac_rankings) + 
  labs(title ='HALLMARK_MYC_TARGETS_V1')
```
# NES matrices 

```{r}
### Extract NES - fill missing with 0
get_NES <- function(res, pathways) {
  out <- setNames(rep(0, length(pathways)), pathways)
  match_idx <- match(res$pathway, pathways)
  out[match_idx[!is.na(match_idx)]] <- res$NES[!is.na(match_idx)]
  return(out)
}

get_NES <- function(res, pathways, padj_thresh = 0.05) {
  out <- setNames(rep(0, length(pathways)), pathways)
  sig_res <- res[res$padj < padj_thresh, ]
  match_idx <- match(sig_res$pathway, pathways)
  out[match_idx[!is.na(match_idx)]] <- sig_res$NES[!is.na(match_idx)]
  return(out)
}

col_palette <- colorRampPalette(brewer.pal(11, "RdBu"))(100)
```


```{r, fig.height=8, fig.width=10}
pathways <- c('HALLMARK_TNFA_SIGNALING_VIA_NFKB',
              'HALLMARK_TGF_BETA_SIGNALING',
              'HALLMARK_KRAS_SIGNALING_UP',
              'HALLMARK_INFLAMMATORY_RESPONSE',
              'HALLMARK_IL2_STAT5_SIGNALING',
              'HALLMARK_INTERFERON_GAMMA_RESPONSE',
              'HALLMARK_IL6_JAK_STAT3_SIGNALING',
              'HALLMARK_INTERFERON_ALPHA_RESPONSE',
              'HALLMARK_FATTY_ACID_METABOLISM',
              'HALLMARK_ADIPOGENESIS',
              'HALLMARK_CHOLESTEROL_HOMEOSTASIS')
NES_matrix <- cbind(
  `ORG38 (P3)` = get_NES(org38_hallmark_res, pathways),
  `ORG66 (P3)` = get_NES(org66_hallmark_res, pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^HALLMARK_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette),
  border_color = 'NA', 
  cellwidth = 75,
  cellheight = 15
)
```
```{r, fig.height=15, fig.width=10}
combined_pathways <- union(org38_reactome_res$pathway[org38_reactome_res$padj<0.05], org66_reactome_res$pathway[org66_reactome_res$padj<0.05])

NES_matrix <- cbind(
  `ORG38 (P3)` = get_NES(org38_reactome_res, combined_pathways),
  `ORG66 (P3)` = get_NES(org66_reactome_res, combined_pathways)
)

### Clean pathway names for plotting 
rownames(NES_matrix) <- gsub("^REACTOME_", "", rownames(NES_matrix))
rownames(NES_matrix) <- gsub("_", " ", rownames(NES_matrix))

### Filter to strongest signals
keep <- apply(abs(NES_matrix), 1, max) > 1
NES_matrix <- NES_matrix[keep, ]

### Plot heatmap
pheatmap(
  NES_matrix,
  scale = "none",
  col=rev(col_palette),
  border_color = 'NA', 
  cellwidth = 80,
  cellheight = 10
)
```

```{r}
org38_p2_pac <- read.csv('../data/DEgenes/P2/ORG38_PACPAC.csv')
org38_p3_pac <- read.csv('../data/DEgenes/P3/ORG38_paPAC.csv')

x <- list(`P2`=org38_p2_pac[org38_p2_pac$adj.P.Val < 0.05, c("X")],
          `P3`=org38_p3_pac[org38_p3_pac$adj.P.Val < 0.05, c("X")]
)

x.table <- venn(x[c("P2","P3")], show.plot = FALSE)
x_attributes <- attributes(x.table)

ggvenn(x, columns = c("P2", "P3"), 
       show_percentage = FALSE,
       text_size = 8)

common_genes <- x_attributes$intersections$`P2:P3`
```

```{r}
org66_p2_pac <- read.csv('../data/DEgenes/P2/ORG66_')
org66_p3_pac <- read.csv('../data/DEgenes/P3/ORG66_paPAC.csv')

x <- list(`P2`=org66_p2_pac[org66_p2_pac$adj.P.Val < 0.05, c("X")],
          `P3`=org66_p3_pac[org66_p3_pac$adj.P.Val < 0.05, c("X")]
)

x.table <- venn(x[c("P2","P3")], show.plot = FALSE)
x_attributes <- attributes(x.table)

ggvenn(x, columns = c("P2", "P3"), 
       show_percentage = FALSE,
       text_size = 8)

common_genes <- x_attributes$intersections$`P2:P3`
```

```{r}
### ORG38 pathways 
plotEnrichment(go_df[['GOBP_CYTOKINE_MEDIATED_SIGNALING_PATHWAY']],
               org38_pac_rankings) + 
  labs(title ='GOBP_CYTOKINE_MEDIATED_SIGNALING_PATHWAY')

plotEnrichment(go_df[['GOBP_REGULATION_OF_INFLAMMATORY_RESPONSE']],
               org38_pac_rankings) + 
  labs(title ='GOBP_REGULATION_OF_INFLAMMATORY_RESPONSE')

plotEnrichment(go_df[['GOBP_RESPONSE_TO_TUMOR_NECROSIS_FACTOR']],
               org38_pac_rankings) + 
  labs(title ='GOBP_RESPONSE_TO_TUMOR_NECROSIS_FACTOR')
```

```{r}
plotEnrichment(pid_df[['PID_INTEGRIN1_PATHWAY']],
               org38_pac_rankings) + 
  labs(title ='PID_INTEGRIN1_PATHWAY')

plotEnrichment(pid_df[['PID_SYNDECAN_2_PATHWAY']],
               org38_pac_rankings) + 
  labs(title ='PID_SYNDECAN_PATHWAY')
```

```{r}
plotEnrichment(reactome_df[['REACTOME_METABOLISM_OF_STEROIDS']],
               org66_pac_rankings) + 
  labs(title ='REACTOME_METABOLISM_OF_STEROIDS')

plotEnrichment(reactome_df[['REACTOME_REGULATION_OF_CHOLESTEROL_BIOSYNTHESIS_BY_SREBP_SREBF']],
               org66_pac_rankings) + 
  labs(title ='REGULATION_OF_CHOLESTEROL_BY_SREBP_SREBF')
```

<!-- ## DE IC50 only  -->

<!-- ```{r} -->
<!-- ### Functions  -->

<!-- source('/Users/dalvinikita/Desktop/macpie/R/filter_genes_by_expression.R') -->

<!-- ### buttons for datatables  -->
<!-- create_gene_dt <- function(x){ -->
<!--   colnames <- c('logFC','AveExpr','t','B') -->
<!--   df <- x %>% -->
<!--     mutate(across(all_of(colnames),round, digits = 2)) %>% -->
<!--     mutate(adj.P.Val = format(adj.P.Val, format='e', digits=4)) %>% -->
<!--     mutate(P.Value = format(P.Value, format='e', digits=4)) -->
<!--   DT::datatable(df, -->
<!--                 extensions = 'Buttons', -->
<!--                 options = list(dom = 'Blfrtip', -->
<!--                                buttons = c('copy', 'csv', 'excel', 'pdf', 'print'), -->
<!--                                lengthMenu = list(c(10,25,50,-1), -->
<!--                                                  c(10,25,50,"All")))) -->
<!-- } -->

<!-- create_dt <- function(x){ -->
<!--   DT::datatable(x, -->
<!--                 extensions = 'Buttons', -->
<!--                 options = list(dom = 'Blfrtip', -->
<!--                                buttons = c('copy', 'csv', 'excel', 'pdf', 'print'), -->
<!--                                lengthMenu = list(c(10,25,50,-1), -->
<!--                                                  c(10,25,50,"All")))) -->
<!-- } -->

<!-- plotRLE <- function(data, main = "", ylab = "Relative Log Expression") { -->
<!--   # Calculate medians -->
<!--   col <- brewer.pal(8, 'Set3') -->
<!--   medians <- apply(data, 1, median, na.rm = TRUE) -->
<!--   rle_data <- data - medians -->

<!--   boxplot(as.data.frame(rle_data), -->
<!--           outline = FALSE, -->
<!--           las = 2, -->
<!--           ylab = ylab, -->
<!--           main = main, -->
<!--           col = col) -->
<!--   abline(h = 0, col = "red", lty = 2) -->
<!-- } -->

<!-- red_blue_palette <- colorRampPalette(brewer.pal(11, "RdBu"))(100) -->
<!-- ``` -->

<!-- Thresholds used for DE analysis: -->

<!-- • pvalue threshold = 0.05 -->

<!-- • log FC threshold = log2(1.2) = 20% change in expression compared to control -->

<!-- Differences from previous analyses of MACseq plate 1: -->

<!-- • Utilising new filtering method: count of 10 in at least 80/96 of samples within an ORG line.  -->

<!-- ```{r,echo=TRUE} -->
<!-- ### Threshold Constants -->

<!-- p_value_threshold <- 0.05 -->
<!-- ``` -->

<!-- ## Data -->
<!-- ```{r} -->
<!-- ### Importing MACseq data -->
<!-- ### Importing metadata -->
<!-- ### Joining data in a Seurat object  -->

<!-- ### import metadata -->
<!-- metadata<-read.csv("/Volumes/bioinf/home/ndalvi/MACseq_PMC228/metadata/PMC228.csv",header=T) %>% -->
<!--   arrange("Barcode") %>% -->
<!--   mutate(id=gsub("Plate","",Plate_ID)) %>% -->
<!--   mutate(label=paste0(Treatment_1," ","(",Well_ID,")")) %>% -->
<!--   mutate(interaction_condition=paste0(Drug,".",Concentration,".",SPLINTR)) %>% -->
<!--   mutate(analysis_condition=paste0(Organoid,".",SPLINTR,".",Time)) -->

<!-- ### load in count data  -->
<!-- raw_counts <- Read10X(data.dir = '/Users/dalvinikita/Documents/3_MACseq_PMC228/raw/PMC228/starsolo/SA_PMC228-null.Solo.out/Gene/raw') -->

<!-- row.names(metadata)<-metadata$label -->

<!-- mac <- CreateSeuratObject(counts = raw_counts, -->
<!--                           project = 'PMC228', -->
<!--                           min.cells = 1, -->
<!--                           min.features = 1) -->

<!-- mac[["percent.mt"]] <- PercentageFeatureSet(mac, pattern = "^mt-|^MT-") -->
<!-- mac[["percent.ribo"]] <- PercentageFeatureSet(mac, pattern = "^Rp[slp][[:digit:]]|^Rpsa|^RP[SLP][[:digit:]]|^RPSA") -->

<!-- ### add location metadata -->
<!-- mac<- mac %>% -->
<!--   inner_join(metadata,by=c(".cell"="Barcode")) -->

<!-- mac <- filter_genes_by_expression(mac, -->
<!--                                        group_by = "Organoid", -->
<!--                                        min_counts = 10, -->
<!--                                        min_samples = 80) -->

<!-- ### exclude mito and ribo genes from the counts -->
<!-- mac<-mac[!grepl("^MT-",rownames(mac))] -->
<!-- mac<-mac[!grepl("^RP[SL][[:digit:]]|^RPLP[[:digit:]]|^RPSA",rownames(mac))] -->

<!-- colnames(mac)<- mac$label -->
<!-- ``` -->

<!-- ### ORG38  -->

<!-- #### Design Matrix and Contrasts  -->

<!-- Factors to include in design matrix:  -->

<!-- • SPLINTR/treatment group (drug_label, drug_unlabelled) -->

<!-- • Concentration (IC12.5, IC25, IC50) -->

<!-- • Time (48H VS DMSO) -->

<!-- • Not including ORG as only 1 ORG line  -->

<!-- Design matrix: -->

<!-- ```{r} -->
<!-- ### selecting metadata and countdata relevant for ORG 38 -->
<!-- ### subset metadata to include only ORG38 cells at 48H -->
<!-- columns <- metadata[metadata$Organoid=='ORG38'&metadata$Time=='48H' & metadata$Treatment_1!="STAUR",] -->
<!-- sampleinfo <- columns -->
<!-- columns <- c(columns[,c("label")]) -->

<!-- ### subset counts to include only relevant ORG38 cells  -->
<!-- counts_matrix <- mac[,columns] -->

<!-- ### factors to include in deisgn matrix: treatment/splintr, concentration, time -->
<!-- group <- sampleinfo$interaction_condition -->

<!-- ### set up design matrix -->
<!-- ### without intercept (means model) -->
<!-- design <- model.matrix(~0+group) -->
<!-- rownames(design) <- rownames(sampleinfo) -->
<!-- datatable(design) -->

<!-- ### Specifying contrasts for analysis -->
<!-- pacVSdmso <- makeContrasts(((groupPACLITAXEL.IC50.SPL1+groupPACLITAXEL.IC50.SPL2+groupPACLITAXEL.IC50.UNLABELLED)/3)-((groupDMSO.0.05.SPL1+groupDMSO.0.05.SPL2+groupDMSO.0.05.UNLABELLED)/3), levels=colnames(design)) -->
<!-- ``` -->

<!-- Voom plot:  -->

<!-- ```{r, fig.height=5, fig.width=5} -->
<!-- ### Normalising data -->
<!-- counts<-counts_matrix@assays$RNA$counts -->
<!-- counts <- DGEList(counts = counts) %>% -->
<!--   calcNormFactors(method = 'TMMwsp') -->
<!-- v <- voom(counts, design, plot = T) -->
<!-- org38_voom_matrix <- v -->
<!-- ``` -->

<!-- RLE of all ORG38 wells before and after normalisation: -->

<!-- ```{r, fig.width=16} -->
<!-- par(mfrow=c(1,2)) -->
<!-- plotRLE(counts$counts, main='RLE of all wells - raw') -->
<!-- plotRLE(v$E, main='RLE of all wells - normalised') -->
<!-- ``` -->

<!-- ##### Paclitaxel - DMSO -->

<!-- Differentially expressed genes between Paclitaxel and DMSO in ORG38: -->

<!-- ```{r} -->
<!-- ### following limma trend workflow -->
<!-- ### difference between PAC and DMSO across concentrations -->
<!-- fit <- lmFit(v, design) -->
<!-- fit <- contrasts.fit(fit, pacVSdmso) -->
<!-- fit <- treat(fit, lfc= log2(1.2)) -->
<!-- saveRDS(fit, file='../output/P1/org38_pac_IC50_fit.rds') -->

<!-- all_genes <- topTreat(fit, number=Inf) -->
<!-- org38_pac_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p') -->
<!-- write.csv(org38_pac_DEgenes, file = '../data/DEgenes/P1/ORG38_NCPACvsDMSO_IC50.csv') -->

<!-- datatable(all_genes) -->

<!-- ### Print summary of DE genes table -->
<!-- up <- paste('• UPREGULATED:',sum(org38_pac_DEgenes$logFC > 0)) -->
<!-- down <- paste('• DOWNREGULATED:',sum(org38_pac_DEgenes$logFC< 0)) -->
<!-- total <- paste('• TOTAL DE genes:', length(rownames(org38_pac_DEgenes))) -->
<!-- summary=paste (total, up, down, sep="\n") -->
<!-- cat(summary[1]) -->

<!-- ``` -->

<!-- ```{r, fig.height=10, fig.width=10} -->
<!-- ### Assigning genes of interest from DEgenes table -->
<!-- ### Formatting counts to log(CPM) for heatmap -->
<!-- ### subsetting wells of interest -->
<!-- genes <- rownames(org38_pac_DEgenes) -->
<!-- samples_of_interest <- sampleinfo[sampleinfo$Drug=='PACLITAXEL'|sampleinfo$Drug=='DMSO',c('label')] -->
<!-- heatmap_counts <- as.data.frame(v$E[,samples_of_interest]) -->


<!-- ### Setting up annotation data frame -->
<!-- type_annotation <- data.frame(Drug = as.factor(sampleinfo[,c(17)]), -->
<!--                               Concentration = as.factor(sampleinfo[,c(11)]), -->
<!--                               SPLINTR = as.factor(sampleinfo[,c(16)])) -->
<!-- rownames(type_annotation) <- rownames(sampleinfo) -->

<!-- pheatmap(heatmap_counts[genes[1:50],], -->
<!--          border_color = 'NA', -->
<!--          annotation_col = type_annotation, -->
<!--          treeheight_row = 0, -->
<!--          cellheight = 8, -->
<!--          color=rev(red_blue_palette)) -->

<!-- ``` -->

<!-- ### ORG49  -->

<!-- #### Design Matrix and Contrasts  -->

<!-- Factors to include in design matrix:  -->

<!-- • SPLINTR/treatment group (drug_label, drug_unlabelled) -->

<!-- • Concentration (IC12.5, IC25, IC50) -->

<!-- • Time (48H VS DMSO) -->

<!-- • Not including ORG as only 1 ORG line  -->

<!-- Design matrix: -->

<!-- ```{r} -->
<!-- ### selecting metadata and countdata relevant for ORG49 -->
<!-- ### subset metadata to include only ORG49 cells at 48H -->
<!-- columns <- metadata[metadata$Organoid=='ORG49'&metadata$Time=='48H' & metadata$Treatment_1!="STAUR",] -->
<!-- sampleinfo <- columns -->
<!-- columns <- c(columns[,c("label")]) -->

<!-- ### subset counts to include only relevant ORG49 cells  -->
<!-- counts_matrix <- mac[,columns] -->

<!-- ### factors to include in deisgn matrix: treatment/splintr, concentration, time -->
<!-- group <- sampleinfo$interaction_condition -->

<!-- ### set up design matrix -->
<!-- ### without intercept (means model) -->
<!-- design <- model.matrix(~0+group) -->
<!-- rownames(design) <- rownames(sampleinfo) -->
<!-- datatable(design) -->

<!-- ### Specifying contrasts for analysis -->
<!-- pacVSdmso <- makeContrasts(((groupPACLITAXEL.IC50.SPL1+groupPACLITAXEL.IC50.SPL2+groupPACLITAXEL.IC50.UNLABELLED)/3)-((groupDMSO.0.05.SPL1+groupDMSO.0.05.SPL2+groupDMSO.0.05.UNLABELLED)/3), levels=colnames(design)) -->
<!-- ``` -->

<!-- Voom plot:  -->

<!-- ```{r, fig.height=5, fig.width=5} -->
<!-- ### Normalising data -->
<!-- counts<-counts_matrix@assays$RNA$counts -->
<!-- counts <- DGEList(counts = counts) %>% -->
<!--   calcNormFactors(method = 'TMMwsp') -->
<!-- v <- voom(counts, design, plot = T) -->
<!-- org49_voom_matrix <- v -->
<!-- ``` -->

<!-- RLE of all ORG49 wells before and after normalisation: -->

<!-- ```{r, fig.width=16} -->
<!-- par(mfrow=c(1,2)) -->
<!-- plotRLE(counts$counts, main='RLE of all wells - raw') -->
<!-- plotRLE(v$E, main='RLE of all wells - normalised') -->
<!-- ``` -->

<!-- ##### Paclitaxel - DMSO -->

<!-- Differentially expressed genes between Paclitaxel and DMSO in ORG38: -->

<!-- ```{r} -->
<!-- ### following limma trend workflow -->
<!-- ### difference between PAC and DMSO across concentrations -->
<!-- fit <- lmFit(v, design) -->
<!-- fit <- contrasts.fit(fit, pacVSdmso) -->
<!-- fit <- treat(fit, lfc= log2(1.2)) -->
<!-- saveRDS(fit, file='../output/P1/org49_pac_IC50_fit.rds') -->

<!-- all_genes <- topTreat(fit, number=Inf) -->
<!-- org49_pac_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p') -->
<!-- write.csv(org49_pac_DEgenes, file = '../data/DEgenes/P1/ORG49_NCPACvsDMSO_IC50.csv') -->

<!-- datatable(all_genes) -->

<!-- ### Print summary of DE genes table -->
<!-- up <- paste('• UPREGULATED:',sum(org49_pac_DEgenes$logFC > 0)) -->
<!-- down <- paste('• DOWNREGULATED:',sum(org49_pac_DEgenes$logFC< 0)) -->
<!-- total <- paste('• TOTAL DE genes:', length(rownames(org49_pac_DEgenes))) -->
<!-- summary=paste (total, up, down, sep="\n") -->
<!-- cat(summary[1]) -->

<!-- ``` -->

<!-- ```{r, fig.height=10, fig.width=10} -->
<!-- ### Assigning genes of interest from DEgenes table -->
<!-- ### Formatting counts to log(CPM) for heatmap -->
<!-- ### subsetting wells of interest -->
<!-- genes <- rownames(org49_pac_DEgenes) -->
<!-- samples_of_interest <- sampleinfo[sampleinfo$Drug=='PACLITAXEL'|sampleinfo$Drug=='DMSO',c('label')] -->
<!-- heatmap_counts <- as.data.frame(v$E[,samples_of_interest]) -->


<!-- ### Setting up annotation data frame -->
<!-- type_annotation <- data.frame(Drug = as.factor(sampleinfo[,c(17)]), -->
<!--                               Concentration = as.factor(sampleinfo[,c(11)]), -->
<!--                               SPLINTR = as.factor(sampleinfo[,c(16)])) -->
<!-- rownames(type_annotation) <- rownames(sampleinfo) -->

<!-- pheatmap(heatmap_counts[genes,], -->
<!--          border_color = 'NA', -->
<!--          annotation_col = type_annotation, -->
<!--          treeheight_row = 0, -->
<!--          cellheight = 8, -->
<!--          color=rev(red_blue_palette)) -->

<!-- ``` -->

<!-- ### ORG66  -->

<!-- #### Design Matrix and Contrasts  -->

<!-- Factors to include in design matrix:  -->

<!-- • SPLINTR/treatment group (drug_label, drug_unlabelled) -->

<!-- • Concentration (IC12.5, IC25, IC50) -->

<!-- • Time (48H VS DMSO) -->

<!-- • Not including ORG as only 1 ORG line  -->

<!-- Design matrix: -->

<!-- ```{r} -->
<!-- ### selecting metadata and countdata relevant for ORG66 -->
<!-- ### subset metadata to include only ORG66 cells at 48H -->
<!-- columns <- metadata[metadata$Organoid=='ORG66'&metadata$Time=='48H' & metadata$Treatment_1!="STAUR",] -->
<!-- sampleinfo <- columns -->
<!-- columns <- c(columns[,c("label")]) -->

<!-- ### subset counts to include only relevant ORG66 cells  -->
<!-- counts_matrix <- mac[,columns] -->

<!-- ### factors to include in deisgn matrix: treatment/splintr, concentration, time -->
<!-- group <- sampleinfo$interaction_condition -->

<!-- ### set up design matrix -->
<!-- ### without intercept (means model) -->
<!-- design <- model.matrix(~0+group) -->
<!-- rownames(design) <- rownames(sampleinfo) -->
<!-- datatable(design) -->

<!-- ### Specifying contrasts for analysis -->
<!-- pacVSdmso <- makeContrasts(((groupPACLITAXEL.IC50.SPL1+groupPACLITAXEL.IC50.SPL2)/2)-((groupDMSO.0.05.SPL1+groupDMSO.0.05.SPL2)/2), levels=colnames(design)) -->
<!-- ``` -->

<!-- Voom plot:  -->

<!-- ```{r, fig.height=5, fig.width=5} -->
<!-- ### Normalising data -->
<!-- counts<-counts_matrix@assays$RNA$counts -->
<!-- counts <- DGEList(counts = counts) %>% -->
<!--   calcNormFactors(method = 'TMMwsp') -->
<!-- v <- voom(counts, design, plot = T) -->
<!-- org66_voom_matrix <- v -->
<!-- ``` -->

<!-- RLE of all ORG66 wells before and after normalisation: -->

<!-- ```{r, fig.width=16} -->
<!-- par(mfrow=c(1,2)) -->
<!-- plotRLE(counts$counts, main='RLE of all wells - raw') -->
<!-- plotRLE(v$E, main='RLE of all wells - normalised') -->
<!-- ``` -->

<!-- ##### Paclitaxel - DMSO -->

<!-- Differentially expressed genes between Paclitaxel and DMSO in ORG66: -->

<!-- ```{r} -->
<!-- ### following limma trend workflow -->
<!-- ### difference between PAC and DMSO across concentrations -->
<!-- fit <- lmFit(v, design) -->
<!-- fit <- contrasts.fit(fit, pacVSdmso) -->
<!-- fit <- treat(fit, lfc= log2(1.2)) -->
<!-- saveRDS(fit, file='../output/P1/org66_pac_IC50_fit.rds') -->

<!-- all_genes <- topTreat(fit, number=Inf) -->
<!-- org66_pac_DEgenes <- topTreat(fit, p.value = p_value_threshold, number=Inf, sort.by='p') -->
<!-- write.csv(org66_pac_DEgenes, file = '../data/DEgenes/P1/ORG66_NCPACvsDMSO_IC50.csv') -->

<!-- datatable(all_genes) -->

<!-- ### Print summary of DE genes table -->
<!-- up <- paste('• UPREGULATED:',sum(org66_pac_DEgenes$logFC > 0)) -->
<!-- down <- paste('• DOWNREGULATED:',sum(org66_pac_DEgenes$logFC< 0)) -->
<!-- total <- paste('• TOTAL DE genes:', length(rownames(org66_pac_DEgenes))) -->
<!-- summary=paste (total, up, down, sep="\n") -->
<!-- cat(summary[1]) -->

<!-- ``` -->

<!-- ```{r, fig.height=10, fig.width=10} -->
<!-- ### Assigning genes of interest from DEgenes table -->
<!-- ### Formatting counts to log(CPM) for heatmap -->
<!-- ### subsetting wells of interest -->
<!-- genes <- rownames(org66_pac_DEgenes) -->
<!-- samples_of_interest <- sampleinfo[sampleinfo$Drug=='PACLITAXEL'|sampleinfo$Drug=='DMSO',c('label')] -->
<!-- heatmap_counts <- as.data.frame(v$E[,samples_of_interest]) -->


<!-- ### Setting up annotation data frame -->
<!-- type_annotation <- data.frame(Drug = as.factor(sampleinfo[,c(17)]), -->
<!--                               Concentration = as.factor(sampleinfo[,c(11)]), -->
<!--                               SPLINTR = as.factor(sampleinfo[,c(16)])) -->
<!-- rownames(type_annotation) <- rownames(sampleinfo) -->

<!-- pheatmap(heatmap_counts[genes,], -->
<!--          border_color = 'NA', -->
<!--          annotation_col = type_annotation, -->
<!--          treeheight_row = 0, -->
<!--          cellheight = 8, -->
<!--          color=rev(red_blue_palette)) -->

<!-- ``` -->
