---
title: "SPLINTR Plate 1 QC"
output: html_document
date: "2024-09-10"
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

library(bartools)
library(Seurat)
library(SingleCellExperiment)
library(dplyr)
library(reshape2)
library(tidyr)
library(DT)
library(ggplot2)
library(ggpubr)
```

##Loading data
```{r, include=TRUE}
counts_path <- "/Volumes/bioinf/home/ndalvi/MACseq_PMC228/results/splintr/MGC-SA-7209-full/counts/MGC-SA-7209_S1.counts.tsv"
counts <- readBartabCounts(counts_path)

#changing BC names to make them shorter
counts$barcode <- sub('SPLINTR_mCHERRY_V2_Barcode','BC',counts$barcode)

head(counts)
```

## QC and filter barcodes

To assess quality of data, we can look at the number of barcodes detected per cell and the number of UMIs supporting the most frequent barcode:

```{r}
plotBarcodesPerCell(counts)

plotUmiPerBarcode(counts)
```

In order to find suitable threshold for filtering, we can plot cumulative sum of how many cells would pass each threshold:

```{r}
plotUmiFilterThresholds(counts)
```
We can also filter by removing minor barcodes from cells. A minor barcode is defined as a barcode that has less than half the number of supporting UMIs than the major barcode in the cell: 

```{r}
counts_filtered <- filterBarcodes(counts, umiCountFilter = 25, umiFractionFilter = 0.5)

plotBarcodesPerCell(counts_filtered)
```
As any given well will have multiple barcodes, we can aggregare barcodes per well by concatenating barcodes and UMIs with ';' :

```{r}
counts_agg <- aggregateBarcodes(counts)
head(counts_agg) 
```

## Adding barcode Data to Seurat object

```{r}
#Setting up Seurat object
expression.data <-
  Read10X(data.dir = "/Volumes/bioinf/home/ndalvi/MACseq_PMC228/raw/PMC228/starsolo/SA_PMC228-null.Solo.out/Gene/raw", strip.suffix = T)

sc <- CreateSeuratObject(
  counts = expression.data,
  min.cells = 3,
  min.features = 200,
  project = "PMC228"
)
```

We can import SPLINTR barcode annotation information to the metadata, making sure the barcodes match the respective well/treatments they come from: 

```{r}
#import metadata
metadata<-read.csv("/Volumes/bioinf/home/ndalvi/MACseq_PMC228/metadata/PMC228.csv",header=T) %>%
  arrange("Barcode") %>%
  mutate(id=gsub("Plate","",Plate_ID)) %>%
  mutate(analysis_condition=paste0(Organoid,".",Treatment_1,".",Time))

metadata<- metadata %>%
  left_join(counts_agg,by=c("Barcode"="cellid"))

rownames(metadata) <- metadata$Barcode

datatable(metadata)
```

## Analysis and Visualisation of barcoded MAC-seq data

```{r}
sc <- Seurat::AddMetaData(sc, metadata)

message("Percentage of cells with no barcode detected")
length(which(is.na(sc$barcode))) / ncol(sc) * 100

message("Percentage of cells with a barcode detected")
length(which(!is.na(sc$barcode))) / ncol(sc) * 100
```

Repeating QC plot to visualist # of barcodes per cell:

```{r}
plotBarcodesPerCell(sc, aggregated = T, sep = ",")
```

We can also visualise number of cells per clone, where a clone is defined as "a barcode or unique combination of barcodes".:
```{r}
plotCellsPerGroup(
  sc,
  group = "barcode",
  order = T,
  threshold = 1,
  plot = T,
  label = T,
  sep = ","
)
```


Running Seurat clustering:
```{r}
sc <- NormalizeData(sc)
sc <- ScaleData(sc)
sc <- FindVariableFeatures(sc)
sc <- RunPCA(sc)
sc <- FindNeighbors(sc, dims=1:10)
sc <- FindClusters(sc, resolution = 0.5)
```

Plotting cells in seurat clusters, by organoid:
```{r, fig.height=8, fig.width=20}

p1<- plotCellsInClusters(sc, 
                    group = "Organoid", 
                    factor = "ORG38", 
                    clusters = "seurat_clusters")

p2<- plotCellsInClusters(sc, 
                    group = "Organoid", 
                    factor = "ORG49", 
                    clusters = "seurat_clusters")

p3<- plotCellsInClusters(sc, 
                    group = "Organoid", 
                    factor = "ORG66", 
                    clusters = "seurat_clusters")

ggarrange(p1,p2,p3, ncol=3)


# plotMetrics(sc, group = "barcode", factor = "nCount_RNA", threshold = 2)
# plotCellsPerGroup(sc, group = "barcode", threshold = 1)
# plotCellsInClusters(sc, group = "Organoid", factor = "ORG38", clusters = "seurat_clusters")
```

#Splitting into organoids
```{r}
org_38 <- metadata[metadata$Organoid=='ORG38' 
                   & metadata$Treatment_1 != 'STAUR' 
                   & metadata$Labelled != 'unlabelled'
                   & metadata$Time=='48H',c('Well_ID',"Treatment_1",'Time',"barcode", "bc.umi.count")]


rownames(org_38)<- org_38$Well_ID
treatments<- c(org_38$Well_ID)
org38_counts_combined <- data.frame(barcode=character())

individual_counts <- function(df, well_id){
  subset <- df[df$Well_ID==well_id,c('barcode', 'bc.umi.count')]
  subset <- separate_longer_delim(data=subset, cols=c('barcode','bc.umi.count'),delim = ';')
  subset$bc.umi.count <- as.integer(c(subset$bc.umi.count))
  names(subset)[2]<- well_id
  subset <- drop_na(subset)
  return(subset)
}

for (well in treatments){
  df <- individual_counts(org_38, well)
  org38_counts_combined<- merge(org38_counts_combined, df, all=TRUE)
}

org38_counts_combined[is.na(org38_counts_combined)]<- 0
barcode_data <- DGEList(counts=org38_counts_combined[,2:29])
rownames(barcode_data$counts)<- org38_counts_combined[,1]

#adding samplesheet information to DGElist object
barcode_data$samples$Sample <- org_38$Well_ID
barcode_data$samples$Treatment <- org_38$Treatment_1
barcode_data$samples$Time <- org_38$Time
barcode_data <- barcode_data[rowSums(barcode_data$counts) != 0, ]

plotBarcodeBubble(barcode_data, 
                  proportionCutoff = 5, 
                  labelBarcodes = T,
                  group="Treatment")

plotBarcodeHistogram(barcode_data, topN = 10, alphaLowFreq = 0)
```
